---
title: "Convert to factor"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: forcats

---

#### Function: `as_factor()`

---

**1\. Convert a character variable (Var1) to a factor**

Review the data (d2)

```{r, echo=FALSE}

source("data.R")
d2

```

View the class for Var1

```{r}

class(d2$Var1)

```

Convert Var1 to a factor. 

* Note: In the case of a character variable, the levels will be chosen based on the order in which they appear. In the case of a numeric variable, the levels will be numerically ordered. You can view levels using the `base::levels()` function.

* Note: We are recoding into a new variable using `dplyr::mutate()` and saving over the original variable by naming the new variable the same name as the original. 

```{r}

d2 <- d2 %>% 
  dplyr::mutate(Var1 = forcats::as_factor(Var1))

class(d2$Var1)

levels(d2$Var1)

```

* Note: If you don't want the levels as they are chosen, you can use the function `forcats::fct_relevel()` instead to convert your variable to a factor **and** assign your level order. However, this function will only work if your variable is a character or factor variable.

```{r}

d2 <- d2 %>% 
  dplyr::mutate(Var1 = forcats::fct_relevel(Var1, "a", "b", "c"))

class(d2$Var1)

levels(d2$Var1)

```


### Package: base

---

#### Function: `as.factor()`

---

**1\. Another function to convert a variable (in this case character variable Var1) to a factor is `base::as.factor()**

Review the data (d2)

```{r, echo=FALSE}

source("data.R")
d2

```

View the class for Var1

```{r}

class(d2$Var1)

```

Convert Var1 to a factor. 

* Note: The levels will be chosen alphabetically or numerically. You can view levels using the `base::levels()` function.

* Note: We are recoding into a new variable using `dplyr::mutate()` and saving over the original variable by naming the new variable the same name as the original. 

```{r}

d2 <- d2 %>% 
  dplyr::mutate(Var1 = as.factor(Var1))

class(d2$Var1)

levels(d2$Var1)

```

However, if you want to assign the levels when converting to a factor, you can use the function `base::factor()` instead with the argument *levels*.

```{r, echo=FALSE}

source("data.R")

```


```{r}

d2 <- d2 %>% 
  dplyr::mutate(Var1 = factor(Var1, levels = c("c", "b", "a")))

class(d2$Var1)

levels(d2$Var1)

```

Return to [Data Types](https://github.com/Cghlewis/data-wrangling-functions/wiki/Data-Types)
