---
title: "Convert variable class"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: base

---

#### Function: `as.factor()`

---

**1\. Convert a character variable to a factor (Var1).**

Review the data (d2)

```{r, echo=FALSE}

source("data.R")
d2

```

View the class for Var1

```{r}

class(d2$Var1)

```

Convert Var1 to a factor.

```{r}

d2 <- d2 %>% mutate(Var1 = as.factor(Var1))

class(d2$Var1)

```

---

#### Function: `as.numeric()`

---

**1\. Convert a character variable to numeric (Var1).**

Review the data (d2)

```{r, echo=FALSE}

source("data.R")
d2

```

View the class for Var3

```{r}

class(d2$Var3)

```

Convert Var3 to numeric.

**Note**: In the case of Var3 you will get a warning message that says NAs were introduced in the transformation. In this case, I am okay with that because "X"s were used to denote NAs previously and I want those to be converted to NA.

```{r, warning = FALSE}

d2 <- d2 %>% mutate(Var3 = as.numeric(Var3))

class(d2$Var3)

```


---

#### Function: `as.character()`

---

**1\. Convert a factor to a character variable (Var1).**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

View the class for Var1

```{r}

class(d1$Var1)

```

Convert Var1 to a character.

```{r}

d1 <- d1 %>% mutate(Var1 = as.character(Var1))

class(d1$Var1)

```

### Package: lubridate

---

#### Function: `lubridate::mdy()`

---

**Note**: mdy() is just the format for Var4, you can adjust the format for your character variable (ex: dmy, ymd, ymd_hms)

**1\. Convert a character variable to a date (Var4).**

Review the data (d2)

```{r, echo=FALSE}

source("data.R")
d2

```

View the class for Var4

```{r}

class(d2$Var4)

```

Convert Var4 to a date.

```{r}

d2 <- d2 %>% mutate(Var4 = lubridate::mdy(Var4))

class(d2$Var4)

d2

```
