---
title: "Convert variable class"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: forcats

---

#### Function: `as_factor()`

---

**1\. Convert a variable (Var1) to a factor**

Review the data (d2)

```{r, echo=FALSE}

source("data.R")
d2

```

View the class for Var1

```{r}

class(d2$Var1)

```

Convert Var1 to a factor. 

* Note: In the case of a character variable, the levels will be chosen based on the order in which they appear. In the case of a numeric variable, the levels will be numerically ordered. You can view levels using the `base::levels()` function.

* Note: We are recoding into a new variable using `dplyr::mutate()` and saving over the original variable by naming the new variable the same name as the original. 

```{r}

d2 <- d2 %>% dplyr::mutate(Var1 = forcats::as_factor(Var1))

class(d2$Var1)

levels(d2$Var1)

```

* Note: If you don't want the levels as they are chosen, you can use the function `forcats::fct_relevel()` instead to convert your variable to a factor **and** assign your level order. However, this function will only work if your variable is a character or factor variable.

```{r}

d2 <- d2 %>% dplyr::mutate(Var1 = forcats::fct_relevel(Var1, "a", "b", "c"))

class(d2$Var1)

levels(d2$Var1)

```


### Package: base

---

#### Function: `as.factor()`

---

**1\. Another function to convert a variable (in this case character variable Var1) to a factor is `base::as.factor()**

Review the data (d2)

```{r, echo=FALSE}

source("data.R")
d2

```

View the class for Var1

```{r}

class(d2$Var1)

```

Convert Var1 to a factor. 

* Note: The levels will be chosen alphabetically or numerically. You can view levels using the `base::levels()` function.

* Note: We are recoding into a new variable using `dplyr::mutate()` and saving over the original variable by naming the new variable the same name as the original. 

```{r}

d2 <- d2 %>% dplyr::mutate(Var1 = as.factor(Var1))

class(d2$Var1)

levels(d2$Var1)

```

However, if you want to assign the levels when converting to a factor, you can use the function `base::factor()` instead with the argument *levels*.

```{r, echo=FALSE}

source("data.R")

```


```{r}

d2 <- d2 %>% dplyr::mutate(Var1 = factor(Var1, levels = c("c", "b", "a")))

class(d2$Var1)

levels(d2$Var1)

```

---

#### Function: `as.numeric()`

---

**1\. Convert a character variable (Var1) to numeric**

Review the data (d2)

```{r, echo=FALSE}

source("data.R")
d2

```

View the class for Var3

* Note: Var3 will read in as character because an X was used to denote missing values in the data.

```{r}

class(d2$Var3)

```

Convert Var3 to numeric.

* Note: In the case of Var3 you will get a warning message that says NAs were introduced in the transformation. In this case, I am okay with that because "X"s were used to denote NAs previously and I want those to be converted to NA.

* Note: We are recoding into a new variable using `dplyr::mutate()` and saving over the original variable by naming the new variable the same name as the original. 

```{r, warning = FALSE}

d2 <- d2 %>% dplyr::mutate(Var3 = as.numeric(Var3))

class(d2$Var3)

```

**2\. Convert all character variables to numeric**

Review the data (d4)

```{r, echo=FALSE}

source("data.R")
d4

```

View the class for all variables

* Note: Var2, Var3 and Var4 are read in as character variables because an X was used to denote missing values in the data.

* Note: Another way to have dealt with these columns that have "X" denoting NA, is to have read in the data using a function where you explicitly state what the missing values are. Example: `readr::read_csv("nameoffile.csv", na="X"). If you read in your file this way, the column would have read in as numeric rather than character.

```{r}

str(d4)

```

Convert all character variables to numeric variables

* Note: Using the function `dplyr::across()` we are applying a transformation across multiple columns
* Note: You must wrap *is.numeric*, a predicate function (returns a true/false), in the `tidyselect` selection helper `where()`.
* Note: We are recoding into a new variable using `dplyr::mutate()` and saving over the original variables.

```{r}

d4 <- d4 %>% dplyr::mutate(dplyr::across(where(is.character), as.numeric))

```

View the class for all variables

```{r}

str(d4)

```

You can also call out the exact variables you want to convert

```{r, echo=FALSE}

source("data.R")

```


```{r, include = FALSE}

d4 %>% dplyr::mutate(dplyr::across(c(Var2:Var4), as.numeric))

```

Or in the case of this data frame, since you essentially want all variables to be numeric (Var1 just happens to already be numeric), you could conver *all* variables to numeric using the `tidyselect` selection helper `everything()`.

```{r, echo=FALSE}

source("data.R")

```


```{r, include = FALSE}

d4 %>% dplyr::mutate(dplyr::across(tidyselect::everything(), as.numeric))

```



**3\. Convert a factor variable (Var3) to numeric**

Review the data (d3)

```{r, echo=FALSE}

source("data.R")
d3

```

View the class for Var3

```{r}

class(d3$Var3)

```

Convert Var3 to numeric.

* Note: We are recoding into a new variable using `dplyr::mutate()` and saving over the original variable by naming the new variable the same name as the original. 

* Note: We MUST convert the numeric variable to character before converting to numeric or we will not retain our original values. Instead, `base::as.numeric()` will convert our factor values to their factor levels (3=2, 8=3, 1=1) which is not what we want. See the first example vs the second example.

**Don't do this**

```{r, warning = FALSE}

d3 %>% dplyr::mutate(Var3 = as.numeric(Var3))


```

**Do this**

```{r, warning = FALSE}

d3 <- d3 %>% dplyr::mutate(Var3 = as.numeric(as.character(Var3)))

d3

class(d3$Var3)

```


---

#### Function: `as.character()`

---

**1\. Convert a factor (Var1) to a character variable**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

View the class for Var1

```{r}

class(d1$Var1)

```

Convert Var1 to a character.

* Note: We are recoding into a new variable using `dplyr::mutate()` and saving over the original variable by naming the new variable the same name as the original. 

```{r}

d1 <- d1 %>% dplyr::mutate(Var1 = as.character(Var1))

class(d1$Var1)

```


### Package: lubridate

---

#### Function: `lubridate::mdy()`

---

**Note**: mdy() is just the format for Var4, you can adjust the format for your character variable (ex: dmy, ymd, ymd_hms)

<br>

**1\. Convert a character variable (Var4) to a date**

Review the data (d2)

```{r, echo=FALSE}

source("data.R")
d2

```

View the class for Var4

```{r}

class(d2$Var4)

```

Convert Var4 to a date

* Note: We are recoding into a new variable using `dplyr::mutate()` and saving over the original variable by naming the new variable the same name as the original. 

```{r}

d2 <- d2 %>% dplyr::mutate(Var4 = lubridate::mdy(Var4))

class(d2$Var4)

```
