---
title: "Convert data type using a data dictionary"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: base

---

#### Function: `as.character()`

---

**1\. Recode all ID variables to character**

Review the data (d8)

```{r, echo=FALSE}

source("data.R")
d8

```

View the class for each variable

```{r}

dplyr::glimpse(d8)

```

I see that my ID variables are stored as numeric but I want them to be stored as character. 

Here I am going to bring in my existing data dictionary to help me choose the variables to recode.

Let's view our data dictionary

```{r, echo=FALSE}

source("data.R")
dict

```


I can pull a vector of all the variables that should be character type.

```{r}

var_char <- dict %>%
  dplyr::filter(type == "character") %>%
  dplyr::pull(var_name)

```

And then use this vector to update the type of multiple variables.

```{r}

d8 <- d8 %>%
  dplyr::mutate(across(all_of(var_char), base::as.character))

```

Now we can check our variable types again.

```{r}

dplyr::glimpse(d8)

```


Return to [Data Types](https://github.com/Cghlewis/data-wrangling-functions/wiki/Data Types)
