---
title: "Review data types"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: base

---

#### Function: `class()`

---

**1\. Review the class for each variable**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

View the class for each variable

```{r}

class(d1$Var1)
class(d1$Var2)
class(d1$Var3)
class(d1$Var4)
class(d1$Var5)

```

You could also view the class for each variable using a function like `purrr::map()` to iterate the `base::class()` function over each variable

* Note: `purrr::map()` returns a list.

```{r}

purrr::map(d1, class)

```

---

#### Function: `typeof()`

---

**1\. Review the data type for each variable**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

View the data type for each variable


```{r}

purrr::map(d1, typeof)

```

* Note: You will notice that our date variable (Var4) returns a data type of double (or numeric). This is because internally, R stores dates as numeric values. The class of Var4 remains date though.

* Note: You will also notice that the data type for our factor variable (Var1) is integer even though the values appear to be character. This is because internally, R stores factor values as integers in order to create ordered levels. To learn more about your factor levels, you can use another function called `base::attributes()`.

```{r}

attributes(d1$Var1)

```


### Package: utils

---

#### Function: `str()`

---

**1\. Review the structure of the data frame/tibble**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

View the structure of the data frame/tibble as well as the class of each variable

```{r}

utils::str(d1)

```


### Package: tibble

---

#### Function: `glimpse()`

---

**1\. Review the structure of the data frame/tibble**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

View the structure of the data frame/tibble as well as the class of each variable

```{r}

tibble::glimpse(d1)

```


