---
title: "Reorder values"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: forcats

---

#### Function: `fct_relevel()`

---

**1\. Relevel the factor s_grade to have the following level order: 2, 3, K, 1 (for demonstration purposes)**

Review the data (d2)

```{r, echo=FALSE}

source("data.R")
d2

```

```{r, echo=FALSE}

# Convert s_grade to factor

d2 <- d2 %>% mutate(s_grade = as_factor(s_grade))

```


Review the structure of s_grade

```{r}

str(d2$s_grade)

```

Relevel s_grade

```{r}

d2 <- d2 %>% mutate(s_grade = fct_relevel(s_grade, "2", "3", "K", "1"))

str(d2$s_grade)

```



**2\. Relevel the factor s_grade to be 1,2,3,K**

Review the data (d2)

```{r, echo=FALSE}

source("data.R")
d2

```

Review the structure of s_grade

```{r, echo=FALSE}

d2 <- d2 %>% mutate(s_grade = forcats::as_factor(s_grade))

```

```{r}

str(d2$s_grade)

```

Relevel "K" after the 3rd position. Since the 4th position is last, you could also use the argument *after = Inf* to refer to the last position.

```{r}

d2 <- d2 %>% mutate(s_grade = fct_relevel(s_grade, "K", after = 3))

str(d2$s_grade)

```


**3\. Relevel the factor s_grade to be 1,K,2,3**

Review the data (d2)

```{r, echo=FALSE}

source("data.R")
d2

```

Review the structure of s_grade

```{r, echo=FALSE}

d2 <- d2 %>% mutate(s_grade = forcats::as_factor(s_grade))

```

```{r}

str(d2$s_grade)

```

Relevel "1" to be first. Since I only want "1" to be first and to keep the order the same otherwise, I can just say to relevel "1" and the rest will be implied.

```{r}

d2 <- d2 %>% mutate(s_grade = fct_relevel(s_grade, "1"))

str(d2$s_grade)

```



### Package: labelled

---

#### Function: `to_factor()`

---

In the rare occasion where you have a labelled variable that you want to convert to a factor and then order those levels, you can use `labelled::to_factor()`.

**1\. Order s_grade levels: K,1,2,3**

Review the data (d3)

```{r, echo=FALSE}

source("data.R")
d3

```

Review the structure of s_grade

```{r, echo=FALSE}

d3 <- d3 %>% labelled::set_value_labels (
  s_grade = c( "First"=1, "Second"=2, "Third"=3, "Kindergarten"=0))

```


```{r}

str(d3$s_grade)

```


Convert s_grade to a factor and then order those levels according to their original values using the argument *sort_levels = "v"*.

```{r}

d3v <- d3 %>% mutate(s_grade = labelled::to_factor(s_grade, sort_levels = "v"))

levels(d3v$s_grade)

```

If you want them to be ordered according to labels you can change the argument to *sort_levels = "l"*. 

```{r}

d3l <- d3 %>% mutate(s_grade = labelled::to_factor(s_grade, sort_levels = "l"))

levels(d3l$s_grade)

```

If you want them to be ordered according the order you originally assigned labels you can change the argument to *sort_levels = "n"*. This is also the default when you do not add the "sort_levels" option.

```{r}

d3n <- d3 %>% mutate(s_grade = labelled::to_factor(s_grade, sort_levels = "n"))

levels(d3n$s_grade)

```

