---
title: "Reorder values"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: forcats

---

#### Function: `forcats::fct_relevel()`

---

**1\. Convert s_grade to a factor with the following level order: K, 1, 2, 3**

Review the data (d2)

```{r, echo=FALSE}

source("data.R")
d2

```

Review the structure of s_grade

```{r}

str(d2$s_grade)

```

Convert s_grade to factor and assign levels

```{r}

d2 <- d2 %>% mutate(s_grade = fct_relevel(s_grade, "K", "1", "2", "3"))

str(d2$s_grade)

```


**2\. Relevel s_grade (already a factor) to be 1,2,3,K**

Review the data (d2)

```{r, echo=FALSE}

source("data.R")
d2

```

Review the structure of s_grade

```{r, echo=FALSE}

d2 <- d2 %>% mutate(s_grade = fct_relevel(s_grade, "K", "1", "2", "3"))

```

```{r}

str(d2$s_grade)

```

Relevel s_grade after the 3rd position. Since the 4th position is last, you could also use the argument *after = Inf* to refer to the last position.

```{r}

d2 <- d2 %>% mutate(s_grade = fct_relevel(s_grade, "K", after = 3))

str(d2$s_grade)

```


**3\. Move Var1 after Var2**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Move the variables

```{r}

d1 %>% dplyr::relocate(Var1, .after = Var2)

```

### Package: dplyr

---

#### Function: `dplyr::select()`

---

**1\. Move Var3 to the front**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Move Var3 using the `dplyr::select()` function in addition to the selection helper `everything()`.

```{r}

d1 %>% dplyr::select(Var3, everything())

```


### Package: labelled

---

#### Function: `labelled::to_factor()`

---

In the rare occasion where you have a labelled variable, that you want to convert to a factor and then order those levels, you can use `labelled::to_factor()`.

**1\. Order s_grade levels: K,1,2,3**

Review the data (d2)

```{r, echo=FALSE}

source("data.R")
d3

```

Review the structure of s_grade

```{r, echo=FALSE}

d3 <- d3 %>% labelled::set_value_labels (
  s_grade = c( "First"=1, "Second"=2, "Third"=3, "Kindergarten"=0))

```


```{r}

str(d3$s_grade)

```


Convert s_grade to a factor and then order those levels according to their original values.

```{r}

d3v <- d3 %>% mutate(s_grade = labelled::to_factor(s_grade, sort_levels = "v"))

levels(d3v$s_grade)

```

If you want them to be ordered according to labels you can select *sort_levels = "l"*.

```{r}

d3l <- d3 %>% mutate(s_grade = labelled::to_factor(s_grade, sort_levels = "l"))

levels(d3l$s_grade)

```

If you want them to be ordered according the order you originally assigned labels you can select *sort_levels = "n"*.

```{r}

d3n <- d3 %>% mutate(s_grade = labelled::to_factor(s_grade, sort_levels = "l"))

levels(d3n$s_grade)

```

