---
title: "Reorder variables"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: dplyr

---

#### Function: `relocate()`

---

**1\. Move `Var3` to be ahead of `Var2`**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Move `Var3` in front of `Var2` by using the *.before* argument

```{r}

d1 %>% 
  dplyr::relocate(Var3, .before = Var2)

```


**2\. Move `Var2` and `Var3` to be ahead of `Var1`**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Move `Var2` and `Var3` in front of `Var1`

* Note: No argument is needed because the unmentioned variables will be placed after the mentioned variables.

```{r}

d1 %>% 
  dplyr::relocate(Var2, Var3)


```

You would get the same result listing all variables

```{r}

d1 %>% 
  dplyr::relocate(Var2, Var3, Var1)


```


**3\. Move `Var1` after `Var2`**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Move `Var1` after `Var2` by using the *.after* argument

```{r}

d1 %>% 
  dplyr::relocate(Var1, .after = Var2)

```

---

#### Function: `select()`

---

**1\. Move `Var3` to the front**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Move `Var3` to the front using the `dplyr::select()` function in addition to the `tidyselect` selection helper `everything()`.

```{r}

d1 %>% 
  dplyr::select(Var3, tidyselect::everything())

```


Return to [Reorder](https://github.com/Cghlewis/data-wrangling-functions/wiki/Reorder)