---
title: "Reorder values"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: dplyr

---

#### Function: `arrange()`

---

**1\. Rearrange the variable names in a data dictionary to match the order of the variables in the corresponding dataset**

Review the data (df)

```{r, echo=FALSE}

source("data.R")
df

```

Review the data dictionary

```{r, echo=FALSE}

dd

```

First, pull the variable name order from the existing dataset.

```{r, echo=FALSE}

df_names <- base::names(df)

```

Next, reorder `var_name` according to `df_names` using `dplyr::arrange()` and `base::match()`.

```{r}

dd %>%
  dplyr::arrange(base::match(var_name, df_names))

```

Another option is to use `base::factor()` and assign levels.

```{r eval = FALSE}

dd %>%
  dplyr::arrange(base::factor(var_name,levels = df_names))

```


Return to [Reorder](https://github.com/Cghlewis/data-wrangling-functions/wiki/Reorder)