---
title: "Reorder variables"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: dplyr

---

#### Function: `relocate()`

---

**1\. Move Var3 to be ahead of Var2**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Move Var3 in front of Var2

```{r}

d1 %>% dplyr::relocate(Var3, .before = Var2)

```


**2\. Move Var2 and Var3 to be ahead of Var1**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Move the variables

```{r}

d1 %>% dplyr::relocate(Var2, Var3)


```


**3\. Move Var1 after Var2**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Move the variables

```{r}

d1 %>% dplyr::relocate(Var1, .after = Var2)

```

---

#### Function: `select()`

---

**1\. Move Var3 to the front**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Move Var3 using the `dplyr::select()` function in addition to the selection helper `everything()`.

```{r}

d1 %>% dplyr::select(Var3, everything())

```


