---
title: "Add random id column"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>


### Package: base

---

#### Function: `sample()`

---

**1\. Add a new column (`study_id`) to the data**

Review the data (d5)

```{r, echo = FALSE}

source("data.R")
d5

```


In order to de-identify my data I want to assign my participants a random unique numeric identifier (with a value between 400 and 500). At some point I will remove first and last name from my data and my de-identified data will only contain a study ID. 

I can first create my new variable using the `dplyr::mutate()` function.

Next, in order to assign my random numbers, I can use the `base::sample()` function to sample values between 400 and 500 and I denote that I need 3 of those values (for my 3 cases) in the argument *size*. This function has another argument *replace* which has the default of *FALSE* which means I want all numbers to be unique.

Before I calculate my new variable I need to do one thing. I need to use `base::set.seed()` and choose any number to add as my seed. Setting this seed ensures that if I ever run this code again at a later time, I will get the same randomly generated numbers each time. Very important!

Last, I have added the `dplyr::mutate()` argument *.after* to denote that I want my new variable to be located after the variable "l_name"


```{r}

base::set.seed(3456)

d5 %>% 
  dplyr::mutate(study_id = base::sample(400:500, size = 3), .after = l_name)


```

Instead of typing in the exact *size* argument, I could also use `dplyr::n()` to get the number of cases.

```{r}

base::set.seed(3456)

d5 %>% 
  dplyr::mutate(study_id = base::sample(400:500, size = n()), .after = l_name)


```

**2\. Add `study_id` values for newly added participants**

Review the new data (d6)

```{r, echo = FALSE}

source("data.R")
d6

```

First we need to get a vector of the `study_id` values that we have already used

```{r echo = FALSE}

base::set.seed(3456)

d5 <- d5 %>% 
  dplyr::mutate(study_id = base::sample(400:500, size = n()), .after = l_name)


```

```{r}

used_ids <- d5 %>%
  dplyr::pull(study_id)

```


Now we can create a vector of available `study_id` values

```{r}

new_ids <- tibble::tibble(study_id = 400:500) %>%
  dplyr::filter(!study_id %in% used_ids) %>%
  dplyr::pull(study_id)

```

Now we can randomly assign an ID to our new students that doesn't overlap with the IDs we've already used

```{r}

base::set.seed(1234)

d6 <- d6 %>%
  dplyr::mutate(study_id = base::sample(new_ids, size = n()), .after = l_name)

d6

```

Now we can append the two datasets to get our full sample.

```{r}

dplyr::bind_rows(d5, d6)

```

**3\. Add a new column (`study_id`) to the data but keep the ID the same within groups**

Review the data (d7)

```{r, echo = FALSE}

source("data.R")
d7

```

Here I have duplicate names in my list but I'm not sure which email is correct so I want to assign all duplicates the same ID.

First I will want to standardize capitalization in names. Then I can assign random IDs across groups, and same IDs within groups.

Here I am using `tidyr::nest()` rather than `dplyr::group_by()` because the latter applies rowwise sampling, allowing IDs to repeat across groups. 

```{r}

set.seed(2489)

d7 %>%
  dplyr::mutate(dplyr::across(c(f_name, l_name), stringr::str_to_title)) %>%
  tidyr::nest(data = c(-l_name, -f_name)) %>%
  dplyr::mutate(id = base::sample(200:205, n(), replace = FALSE)) %>%
  tidyr::unnest(data)

```

Similarly we could assign IDs to unique names and then join those IDs to the full file.

```{r}

set.seed(2489)

ids <- d7 %>%
  dplyr::mutate(dplyr::across(c(f_name, l_name), stringr::str_to_title)) %>%
  dplyr::distinct(f_name, l_name) %>%
  dplyr::mutate(id = base::sample(200:205, n()))

d7 %>%
  dplyr::mutate(dplyr::across(c(f_name, l_name), stringr::str_to_title)) %>%  
  dplyr::left_join(ids, by = c("f_name", "l_name"))

```

Return to [Randomize](https://github.com/Cghlewis/data-wrangling-functions/wiki/Randomize)
