---
title: "Add random id column"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>


### Package: base

---

#### Function: `sample()`

---

**1\. Add a new column (study id) to the data**

Review the data (d5)

```{r, echo = FALSE}

source("data.R")
d5

```


In order to de-identify my data I want to assign my participants a random unique numeric identifier (with a value between 400 and 500). At some point I will remove first and last name from my data and my de-identified data will only contain a study ID. 

I can first create my new variable using the `dplyr::mutate()` function.

Next, in order to assign my random numbers, I can use the `base::sample()` function to sample values between 400 and 500 and I denote that I need 3 of those values (for my 3 cases) in the argument *size*. This function has another argument *replace* which has the default of *FALSE* which means I want all numbers to be unique.

Before I calculate my new variable I need to do one thing. I need to use `base::set.seed()` and choose any number to add as my seed. Setting this seed ensures that if I ever run this code again at a later time, I will get the same randomly generated numbers each time. Very important!

Last, I have added the `dplyr::mutate()` argument *.after* to denote that I want my new variable to be located after the variable "l_name"


```{r}

base::set.seed(3456)

d5 %>% 
  dplyr::mutate(study_id = base::sample(400:500, size = 3), .after = l_name)


```

Instead of typing in the exact *size* argument, I could also use the `base::length()` argument and choose first or last name to get the number of cases.

```{r}

set.seed(3456)

d5 %>% 
  dplyr::mutate(study_id = base::sample(400:500, size = base::length(f_name)), .after = l_name)


```


Return to [Randomize](https://github.com/Cghlewis/data-wrangling-functions/wiki/Randomize)
