---
title: "Randomly sample cases"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: dplyr

---

#### Function: `slice_sample()`

---


**1\. Randomly choose 3 teachers per grade level**

Review the data (d1)

```{r, echo = FALSE}

source("data.R")
d1

```

I want to send surveys out to 3 teachers per grade level so I want to randomly choose those teachers from our existing roster.

Let's first see how many teachers we have per grade level.

```{r}

d1 %>%
  janitor::tabyl(grade)
```

Next we can create our random samples. 

First we need to use `base::set.seed()` and choose any number to add as my seed. Setting this seed ensures that if I ever run this code again at a later time, I will get the same random sample each time. Very important!

Next we can use `dplyr::group_by()` to group our data by grade level and then use `dplyr::slice_sample()` to randomly select our cases. We use the argument *n* to say how many cases we want from each group.

```{r}

base::set.seed(1234)

d1 %>%
  dplyr::group_by(grade) %>%
  dplyr::slice_sample(n = 3)

```



Return to [Create New Variables](https://github.com/Cghlewis/data-wrangling-functions/wiki/Create New Variables)
