---
title: "Randomly assign treatment"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: base

---

#### Function: `sample()`

---


**1\. Randomly assign treatment to cases**

Review the data (d1)

```{r, echo = FALSE}

source("data.R")
d1

```

I want to randomly assign a treatment variable to cases. 1 = treatment and 0 = control. I want an even disbursement.

First we need to use `base::set.seed()` and choose any number to add as my seed. Setting this seed ensures that if I ever run this code again at a later time, I will get the same random sample each time. Very important!

Within my `base::sample()` function I can use `base::rep()` to replicate the values of 0 and 1 and assign the number of times to use each value in the *times* argument. I could put the literal number `7` in this argument, or if I wanted R to do the calculation for me, I could use `base::length()` to calculate the number of rows and divide it by 2.

* Note: I am using `dplyr::mutate()` to create my new "treat" variable.

```{r}

set.seed(1234)

d1_new <- d1 %>%
  dplyr::mutate(treat = base::sample(base::rep(c(1,0), times = base::length(id)/2)))

d1_new

```

Let's make sure we have an even number of treatment and control

```{r}

d1_new %>%
  janitor::tabyl(treat)

```


If I cared about grade level and wanted to randomly assign treatment evenly within grade, I could add a `dplyr::group_by()` statement and group by grade before randomly assigning.

Notice now that treatment cannot evenly divided though (each grade level has 7 participants). In this case you can assign different numbers of times to repeat each value. For example, here I chose to repeat 1 three times and 0 four times.

```{r}

set.seed(1234)

d1_new <- d1 %>%
  dplyr::group_by(grade) %>%
  dplyr::mutate(treat = base::sample(base::rep(c(1,0), times = c(3,4))))

d1_new

```

Let's review our treatment now

```{r}

d1_new %>%
  janitor::tabyl(grade, treat)

```

Return to [Randomize](https://github.com/Cghlewis/data-wrangling-functions/wiki/Randomize)
