---
title: "Select columns using a data dictionary"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```


### Package: dplyr

---

#### Function: `select()`

---

**1\. Drop columns using a data dictionary**

Review the data (d2)

```{r, echo=FALSE}

source("data.R")
d2

```

Review our data dictionary.

Oftentimes we develop a data dictionary to include everything that exists in our raw data. It is reasonable to include a column that describes what you plan to do with your variables, such as drop a column.

```{r, echo = FALSE}

dictionary

```

We can now create a character vector of all of the variables we wish to drop using our data dictionary.

* Note: We use `dplyr::filter()` to only keep the variables we wish to drop.

* Note: We use `dplyr::pull()` to extract the one column with the names of the variables in our data dictionary and create a character vector

```{r}

drop_vars <- dictionary %>%
  dplyr::filter(keep == "no") %>%
  dplyr::pull(var_name)

drop_vars

```

We can now use this character vector to select/remove variables from our dataset.

* Note: We use `tidyselect::all_of()` to select variables that are contained in a character vector (an environment variable). 

* Note: We add the `-` to denote that we want to remove variables.

```{r}

d2 %>%
  dplyr::select(-all_of(drop_vars))

```


Return to [Select Variables](https://github.com/Cghlewis/data-wrangling-functions/wiki/Select-Variables)
