---
title: "Remove empty columns"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```


### Package: expss

---

#### Function: `drop_empty_columns()`

---

**1\. Remove empty columns**

Review the data (d4)

```{r, echo=FALSE}

source("data.R")
d4

```

Count the number of columns in the current data using the function `base::ncol()`

```{r}

ncol(d4)

```


Filter out any completely empty column

```{r}

d4 <- d4 %>% expss::drop_empty_columns()

d4

```

Count the number of cases after you filter

```{r}

ncol(d4)

```



### Package: janitor

---

#### Function: `remove_empty()`

---


**1\. Remove empty columns**

Review the data (d4)

```{r, echo=FALSE}

source("data.R")
d4

```

Count the number of columns in the current data using the function `base::ncol()`

```{r}

ncol(d4)

```


Remove any completely empty column

* Note: We add the argument *which* and in this case we select "cols" rather than the other option "rows".

```{r}

d4 <- d4 %>% janitor::remove_empty(which = "cols")

d4

```

Count the number of cases after you filter

```{r}

ncol(d4)

```

Return to [Select Variables](https://github.com/Cghlewis/data-wrangling-functions/wiki/Select-Variables)
