---
title: "Select columns"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```


### Package: dplyr

---

#### Function: `select()`

---

**1\. Drop columns (extra1, extra2, extra3)**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Drop extra1, extra2, extra3

* Note: Note the use of *-* to remove columns.

```{r}

d1 %>% 
  dplyr::select(-c(extra1:extra3))

```

You can also drop columns based on shared column attributes using `tidyselect` selection helpers such as `starts_with()`.

* Note: Other `tidyselect` selection helpers include `contains()`, `ends_with()`, `where()`, `matches()`, and more.

```{r}

d1 %>% 
  dplyr::select(-(tidyselect::starts_with("extra")))

```


**2\. Keep selected columns (id, test_score1, test_score2)**

Review the data (d2)

```{r, echo=FALSE}

source("data.R")
d2

```

Keep id, test_score1 and test_score2

```{r}

d2 %>% 
  dplyr::select(id:test_score2)

```

You can also keep columns based on shared column attributes using `tidyselect` selection helpers such as `contains()`.

* Note: Other `tidyselect` selection helpers include `starts_with()`, `ends_with()`, `where()`, `matches()`, and more.

```{r}

d2 %>% 
  dplyr::select(id, tidyselect::contains("test"))

```

You could also write this with both id and test in the `tidyselect::contains` selection helper. Note the use of `c()` to create a character vector.

```{r}

d2 %>% 
  dplyr::select(tidyselect::contains(c("id", "test")))

```

Another way you can write this is using the selection helper `matches()` which matches based on a regex operator rather than a literal string like `contains()`.

```{r}

d2 %>% 
  dplyr::select(tidyselect::matches("id|test"))

```


**3\. Keep only numeric columns**

Review the data (d2)

```{r, echo=FALSE}

source("data.R")
d2

```

Keep numeric columns by adding the selection helper `where()`.

* Note: You must wrap *is.numeric*, a predicate function (returns a true/false), in the `tidyselect` selection helper `where()`.

```{r}

d2 %>% 
  dplyr::select(where(is.numeric))

```

And if I wanted all numeric columns except `extra2`, I could simply add this to my statement with the `-` operator.

```{r}

d2 %>% 
  dplyr::select(where(is.numeric), -extra2)

```


**4\. Select all columns except extra1**

Review the data (d4)

```{r, echo=FALSE}

source("data.R")
d4

```

Similar to above, I could use the *-* to remove columns

```{r}

d4 %>% 
  dplyr::select(-extra1)

```

Another way I could select all columns except extra1 is to select extra2 through all columns until the last column, using the `tidyselect` selection helper `last_col()`. This is helpful if you don't always know what the last column is but you know you want to keep everything until the end.

```{r}

d4 %>% 
  dplyr::select(extra2:last_col())

```

**5\. Keep selected columns (stress1_reverse, anxiety1_reverse)**

Review the data (d7)

```{r, echo=FALSE}

source("data.R")
d7 %>%
  relocate(stress1, stress1_reverse, anxiety1, anxiety1_reverse)

```

Because there are both reversed and non-reversed versions of these variables, we will need to be careful to not select the non-reversed versions.

Currently this is a small number of variables to select and we could easily just list these variables out in a select statement. But if there were 40 of these variables, we may want to be more strategic about selecting these variables.

We could add 2 selection criteria to select our variables that start with "stress" or "anxiety" and end in "reverse" in order to grab all of our variables of interest.

```{r}

d7 %>%
  dplyr::select(tidyselect::starts_with(c("stress", "anxiety")) & tidyselect::ends_with("reverse"))

```



Return to [Select Variables](https://github.com/Cghlewis/data-wrangling-functions/wiki/Select-Variables)
