---
title: "Select columns within an operation"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```


### Package: dplyr

---

#### Function: `across()`

---

* Note: `dplyr::across()` is a function which allows us to use the `dplyr::select()` semantics within functions like `dplyr::summarise()` or `dplyr::mutate()`

<br>

**1\. Round all mean scores to be 2 digits**

Review the data (d5)

```{r, echo=FALSE}

source("data.R")

d5 %>%
  mutate(across(contains("mean"), ~sprintf("%.3f", .)))


```

We want to round all mean scores to 2 digits.

Since we want to apply a transformation to multiple variables, we can use the `dplyr::across()` function which allows us to use the `dplyr::select()` semantics.

* Note: I am using the `dplyr::mutate()` function to modify our existing variables.

* Note: I am using the `base::round()` function to round our variables with the *digits* argument. **It is very important to note how this function does rounding, which is, it rounds to the even number when rounding at 5**. This is not your normal way of thinking of rounding. When you have a value of 23.5, it will round up to 24. When you have a value of 22.5, it will round down to 22. However, when you round with 1 or more digits, since decimal places are not represented exactly, you may get different results depending on your OS: 1.75 may round to 1.7 or 1.8. See documentation on `round` for more information.

```{r, eval = FALSE}

d5 %>%
  dplyr::mutate(dplyr::across(scale1_mean:scale3_mean,
                round, digits = 2))

```

```{r, echo = FALSE}

d5 %>%
  mutate(across(scale1_mean:scale3_mean,
                round, digits = 2)) %>%
  mutate(across(contains("mean"), ~sprintf("%.2f", .)))

```

Similar to `dplyr::select()` you can also select your variables by using a character vector

```{r, eval = FALSE}

d5 %>%
  dplyr::mutate(dplyr::across(c(scale1_mean, scale2_mean, scale3_mean),
                round, digits = 2))

```

Or using `tidyselect` selection helpers.

* Note: Other `tidyselect` selection helpers include `starts_with()`, `ends_with()`, `where()`, `matches()`, and more.

```{r, eval = FALSE}

d5 %>%
  dplyr::mutate(dplyr::across(contains("mean"),
                round, digits = 2))

```

---

#### Function: `if_any()` or `if_all()`

---

* Note: `dplyr::if_any()` and `dplyr::if_all()` are predicate functions used to select columns in a filtering capacity and return a logical vector. This function is available in version 1.0.5 of `dplyr`. `dplyr::if_any()` returns a true when the statement is true for **any** of the variables. `dplyr::if_all()` returns a true when the statement is true for **all** of the variables.

<br>

**1\. Create a new variable that states if the survey was complete or not**

Review the data (d6)

```{r, echo = FALSE}

d6

```

Create our new complete variable that is 1 if the survey is complete and 0 if it is not.

* Note: I am using the `dplyr::mutate()` function to modify our existing variables.

* Note: I am using `dplyr::case_when()` to recode existing variables into a new variable.


```{r}

d6 %>%
  dplyr::mutate(complete = 
           dplyr::case_when(
             dplyr::if_all(q1:q3, ~ is.na(.x)) ~ 0,
             TRUE ~ 1
           ))

```



Return to [Select Variables](https://github.com/Cghlewis/data-wrangling-functions/wiki/Select-Variables)
