---
title: "Add variable labels"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>


### Package: labelled

---

#### Function: `set_variable_labels()`

---

**1\. Set labels for one or more variables (Var1, Var2, and Var3)**

Review the data (d4)

```{r, echo=FALSE}

source("data.R")
d4

```

Add variable labels

```{r}

d4 <- d4 %>%
  labelled::set_variable_labels (Var1 = "Name", Var2="Interest in Homework", 
                                 Var3 ="Interest in School")

labelled::var_label(d4)

```

**2\. Derive variable labels from variable names (simple example)**

Review the data (d7)

```{r, echo=FALSE}

source("data.R")
d7

```

Set variable labels by adding the `base::names()` function which retrieves the names of objects

```{r}

d7 <- d7 %>% 
  labelled::set_variable_labels(.labels=names(d7))

d7 %>% labelled::var_label()

```

**3\. Derive variable labels from variable names (add sentence case)**

Review the data (d7)

```{r, echo=FALSE}

source("data.R")
d7

```

Set variable labels by adding the `snakecase::to_sentence_case()` function in addition to the `base::names()` function.

```{r}

d7 <- d7 %>% 
  labelled::set_variable_labels(.labels=snakecase::to_sentence_case(names(d7)))


d7 %>% labelled::var_label()


```

**4\. Derive variable labels from variable names (add additional info to labels)**

Review the data (d7)

```{r, echo=FALSE}

source("data.R")
d7

```

Set variable labels by adding the `stringr::str_replace_all()` function in addition to the `base::names()` function

```{r}

d7 <- d7 %>% 
  labelled::set_variable_labels(
    .labels = stringr::str_replace_all(names(.),c("_"=" ", 
                                         "map"="map score",
                                         "math"="in mathematics",
                                         "ela"= "in english language arts",
                                         "20"="19-20 school year")))

d7 %>% labelled::var_label()

```

**5\. Derive variable labels from variable names (add additional info to labels and use sentence case)**

Review the data (d7)

```{r, echo=FALSE}

source("data.R")
d7

```

Set variable labels by adding the `stringr::str_replace_all()` function in addition to the `snakecase::to_sentence_case()` and the `base::names()` function

```{r}

d7 <- d7 %>% 
  labelled::set_variable_labels(
    .labels = snakecase::to_sentence_case(stringr::str_replace_all(names(.),c("_"=" ", 
                                         "map"="map score",
                                         "math"="in mathematics",
                                         "ela"= "in english language arts",
                                         "20"="19-20 school year"))))

d7 %>% labelled::var_label()


```


Return to [Label Data](https://github.com/Cghlewis/data-wrangling-functions/wiki/labelling-data)