---
title: "Add missing value labels"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```


### Package: labelled

**Note**: Using the `labelled` package, adding missing label values will not allow these values to be treated as NA in R. However, they will be labelled as missing when you export to a file type such as .sav and will be treated as missing values in those programs.

---

#### Function: `set_na_values()`

---

**1\.Add a label for one missing value (-999), for one variable (`Var3`).**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Add NA label for -999

```{r}

d1 <- d1 %>%
  labelled::set_na_values(Var3 = -999)

labelled::na_values(d1$Var3)

```

Let's review what this data looks like now.

You can see that -999 has a label of "NA" for `Var3` now.

```{r}
d1

```

You can also see that R recognizes these values as NA.

```{r}

is.na(d1)

```

Be careful though, many functions, however, will not recognize these values as NA. As just one example, both `rowSums()` and `sum()` have the argument *na.rm*. However, these labelled NA values will not be recognized as NA and will be included in your sum, giving you very large negative numbers.

On the flip side, other functions, such as those within the `pointblank` package and even the function `tidyr::replace_na()` both seem to recognize labelled NA as true NA values.

As a rule, test the function first to see how it interacts with labelled missing values. You may need to convert these to true NAs in order to interact with them in your normal capacity.


**2\. Add labels for multiple missing values (-999, 0), for one variable (`Var3`)**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Add NA labels for -999 and 0

```{r}

d1 <- d1 %>%
  labelled::set_na_values(Var3 = c(-999, 0))

labelled::na_values(d1$Var3)

```

If all values -999 to 0 are considered missing you could use the `labelled::set_na_range()` function instead. This is a good alternative if you have more user-defined missing values than are allowed in other statistical programs such as SPSS, which only allow 3 distinct values.

Note that the range will not appear when you use `labelled::na_values()` to review the values. However, if you export the data to a program like SPSS, the ranges will appear as user defined missing values.

```{r}

d1 <- d1 %>% 
  labelled::set_na_range(Var3=c(-999,0))

```


**3\. Add labels to multiple variables (`Var2` and `Var3`) with the same missing value**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Add NA labels for -999 and 0

```{r}

d1 <- d1 %>%
  labelled::set_na_values(
    Var2 = c(-999, 0),
    Var3 = c(-999, 0))

labelled::na_values(d1)

```

If I wanted to just review NA value labels for specific variables, I can select my variables of interest using `dplyr::select()` and then iterate the `labelled::na_values()` function over those variables using the `purrr::map()` function

```{r}

d1 %>%
  dplyr::select(Var2, Var3) %>%
  purrr::map(labelled::na_values)

```

Since both of these variables have the same missing value labels, I could write the function like this.

* Note: The `na_values<-` needs to be in backticks with no spaces. And you also need to have the `labelled` package loaded.

```{r, eval = FALSE}

d1 %>% 
  dplyr::mutate(
  dplyr::across(Var2:Var3, ~(`na_values<-`
                                (., c(-999, 0)))))

```

And again, if all values between -999 and 0 are missing values, I could use `labelled::na_range()` instead.

```{r, eval = FALSE}

d1 %>% 
  dplyr::mutate(
  dplyr::across(Var2:Var3, ~(`na_range<-`
                                (., c(-999, 0)))))

```


Return to [Label Data](https://github.com/Cghlewis/data-wrangling-functions/wiki/Label Data)