---
title: "Add variable labels"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>


### Package: labelled

---

#### Function: `set_variable_labels()`

---

**1\. Set labels for one or more variables (Var1, Var2, and Var3)**

Review the data (d4)

```{r, echo=FALSE}

source("data.R")
d4

```

Add variable labels

```{r}

d4 <- d4 %>%
  labelled::set_variable_labels (Var1 = "Name", Var2="Interest in Homework", 
                                 Var3 ="Interest in School")

labelled::var_label(d4)

```

**2\. Derive variable labels from variable names (simple example)**

Review the data (d7)

```{r, echo=FALSE}

source("data.R")
d7

```

Set variable labels by adding the `base::names()` function which retrieves the names of objects

```{r}

d7 <- d7 %>% 
  labelled::set_variable_labels(.labels=names(d7))

d7 %>% labelled::var_label()

```

**3\. Derive variable labels from variable names (add sentence case)**

Review the data (d7)

```{r, echo=FALSE}

source("data.R")
d7

```

Set variable labels by adding the `snakecase::to_sentence_case()` function in addition to the `base::names()` function.

```{r}

d7 <- d7 %>% 
  labelled::set_variable_labels(.labels=snakecase::to_sentence_case(names(d7)))


d7 %>% labelled::var_label()


```

**4\. Derive variable labels from variable names (add additional info to labels)**

Review the data (d7)

```{r, echo=FALSE}

source("data.R")
d7

```

Set variable labels by adding the `stringr::str_replace_all()` function in addition to the `base::names()` function

```{r}

d7 <- d7 %>% 
  labelled::set_variable_labels(
    .labels = stringr::str_replace_all(names(.),c("_"=" ", 
                                         "map"="map score",
                                         "math"="in mathematics",
                                         "ela"= "in english language arts",
                                         "20"="19-20 school year")))

d7 %>% labelled::var_label()

```

**5\. Derive variable labels from variable names (add additional info to labels and use sentence case)**

Review the data (d7)

```{r, echo=FALSE}

source("data.R")
d7

```

Set variable labels by adding the `stringr::str_replace_all()` function in addition to the `snakecase::to_sentence_case()` and the `base::names()` function

```{r}

d7 <- d7 %>% 
  labelled::set_variable_labels(
    .labels = snakecase::to_sentence_case(stringr::str_replace_all(names(.),c("_"=" ", 
                                         "map"="map score",
                                         "math"="in mathematics",
                                         "ela"= "in english language arts",
                                         "20"="19-20 school year"))))

d7 %>% labelled::var_label()


```


---

#### Function: `copy_labels_from()`

---

**1\. Copy labels from another dataset that has all the same variables**

Review the new data (d5)

```{r}

d5

```

Review the labelled dataset with the same variables (d4)

```{r}

d4

```


View the missing labels for the new data (d5)

```{r}

labelled::var_label(d5)

```

```{r, echo=FALSE}

d4 <- d4 %>%
  labelled::set_variable_labels (Var1 = "Name", Var2="Interest in Homework", 
                                 Var3 ="Interest in School")

```

View the labels for the labelled data (d4)

```{r}

labelled::var_label(d4)

```


Copy labels from d4 to d5

```{r}

d5 <- d5 %>% labelled::copy_labels_from(d4)

d5 %>% labelled::var_label()

```


**2\. Copy labels from another dataset that only has some of the same variables**

`labelled::copy_labels_from()` also works if not all of the same variables exist in both datasets

Review the new dataset d6 that has some of the same variables as d4 but with no labels

```{r}

d6

```

Review the labelled dataset with some of the same variables (d4)

```{r}

d4

```


View the missing labels for d6

```{r}

labelled::var_label(d6)

```


```{r, echo=FALSE}

d4 <- d4 %>%
  labelled::set_variable_labels (Var1 = "Name", Var2="Interest in Homework", 
                                 Var3 ="Interest in School")

```


View the labels for the labelled data (d4)

```{r}

labelled::var_label(d4)

```


Copy labels from d4 to d6

```{r}

d6 <- d6 %>% labelled::copy_labels_from(d4)

d6 %>% labelled::var_label()

```


---

#### Function: `copy_labels()`

---

**Note:** With this function, the dataset you are copying **from** goes first.

<br>

**1\. Similar to `labelled::copy_labels_from`, `labelled::copy_labels` copies labels from one dataset to another**

Review the new data (d5)

```{r, echo=FALSE}

source("data.R")
d5

```

Review the labelled dataset with the same variables (d4)

```{r, echo=FALSE}

d4 <- d4 %>%
  labelled::set_variable_labels (Var1 = "Name", Var2="Interest in Homework", 
                                 Var3 ="Interest in School")
d4

```


View the missing labels for the new data (d5)

```{r}

labelled::var_label(d5)

```

```{r, echo=FALSE}

d4 <- d4 %>%
  labelled::set_variable_labels (Var1 = "Name", Var2="Interest in Homework", 
                                 Var3 ="Interest in School")

```


View the labels for the labelled data (d4)

```{r}

labelled::var_label(d4)

```

Copy labels from d4 to d5

```{r}

d5 <- labelled::copy_labels(d4, d5)

d5 %>% labelled::var_label()

```
