---
title: "Add missing value labels"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```


### Package: labelled

**Note**: Using the `labelled` package, adding missing label values will not allow these values to be treated as NA in R. However, they will be labelled as missing when you export to a file type such as .sav and will be treated as missing values in those programs.

---

#### Function: `set_na_values()`

---

**1\.Add a label for one missing value (-999), for one variable (Var3).**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Add NA label for -999

```{r}

d1 <- d1 %>% labelled::set_na_values(Var3=-999)

labelled::na_values(d1$Var3)

```


**2\. Add labels for multiple missing values (-999, 0), for one variable (Var3)**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Add NA labels for -999 and 0

```{r}

d1 <- d1 %>% 
  labelled::set_na_values(Var3=c(-999,0))

labelled::na_values(d1$Var3)

```

If all values -999 to 0 are considered missing you could use the `labelled::set_na_range()` function instead. This is a good alternative if you have more user-defined missing values than are allowed in other statistical programs such as SPSS, which only allows maybe 4 or 5 distinct values.

Note that the range will not appear when you use `labelled::na_values()` to review the values. However, if you export the data to a program like SPSS, the ranges will appear as user defined missing values.

```{r}

d1 <- d1 %>% 
  labelled::set_na_range(Var3=c(-999,0))

```


**3\. Add labels to multiple variables (Var2 and Var3) with the same missing value**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Add NA labels for -999 and 0

```{r}

d1 <- d1 %>% labelled::set_na_values(Var2 = c(-999,0), 
                                     Var3= c(-999, 0))

labelled::na_values(d1)

```

If I wanted to just review NA value labels for specific variables, I can select my variables of interest using `dplyr::select()` and then iterate the `labelled::na_values()` function over those variables using the `purrr::map()` function

```{r}

d1 %>%
  dplyr::select(Var2, Var3) %>%
  purrr::map(labelled::na_values)

```

Since both of these variables have the same missing value labels, I could write the function like this.

* Note: The na_values<- needs to be in backticks. And you also need to have the `labelled` package loaded.

```{r, eval = FALSE}

d1 %>% 
  dplyr::mutate(
  dplyr::across(Var2:Var3, ~(`na_values<-`
                                (., c(-999, 0)))))

```

Return to [Label Data](https://github.com/Cghlewis/data-wrangling-functions/wiki/Label Data)