---
title: "Add value labels"
output:
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: labelled

---

#### Function: `set_value_labels()`

---

**1\. Add value labels for one or more variables (`Var2` and `Var3`)**

Review the data (d3)

```{r, echo=FALSE}

source("data.R")
d3

```

Add value labels to `Var2` and `Var3`

* Note: Parentheses are *not* required around labels here but I am using them around multi-word labels with spaces. Also using them around all labels will not hurt anything.

```{r}

d3 <- d3 %>%
  labelled::set_value_labels(
    Var2 = c("not interested" = 1, "mildly interested" = 2, "very interested" = 3),
    Var3 = c(none = 1, "reduced-price" = 2, free = 3)
  )

labelled::val_labels(d3)

```

**2\. Create labels, then add labels to variables (`Var2` and `Var3`)**

Review the data (d3)

```{r, echo=FALSE}

source("data.R")
d3

```

Create labels

* Note: Parentheses are *not* required around labels here, although adding them will not hurt.

```{r}

lab <- c(disagree = 1, neutral = 2, agree = 3)

```


Add value labels to `Var2` and `Var3`

```{r}

d3 <- d3 %>%
  labelled::set_value_labels(Var2 = lab, Var3 = lab)

d3 %>%
  labelled::val_labels()

```

---

#### Function: `add_value_labels()`

---

**1\. Update or add one or more value labels to a variable (`Var3`) without replacing/removing the other value labels**

Review the data (d3)

```{r, echo=FALSE}

source("data.R")
d3

```

View current labels for `Var3`

```{r, echo=FALSE}

d3 <- d3 %>%
  labelled::set_value_labels(
    Var3 = c("none" = 1, "reduced-price" = 2, "free" = 3)
  )

d3 %>%
  labelled::val_labels()

```

Update the label for 1 in `Var3`

* Note: Parentheses are *not* required around labels here, although adding them will not hurt.

```{r}

d3 <- d3 %>%
  labelled::add_value_labels(Var3 = c(paid = 1))

d3 %>%
  labelled::val_labels()

```

**2\. Add a new value label to a newly recoded variable (`Var3`) without replacing/removing the other value labels**

Review the data (d3)

```{r, echo=FALSE}

source("data.R")
d3

```

View current labels for `Var3`

```{r, echo=FALSE}

d3 <- d3 %>%
  labelled::set_value_labels(
    Var3 = c("some" = 1, "a lot" = 2, "none" = 3)
  )

d3 %>%
  labelled::val_labels()

```

Recode 1 to 0 for `Var3`

* Note: We used `dplyr::recode()` to recode 3 to 0. We use `janitor::tabyl()` to check the results of the recode.

```{r}

d3 <- d3 %>%
  dplyr::mutate(Var3 = dplyr::recode(Var3, `3` = 0))

d3 %>%
  janitor::tabyl(Var3)

```

Review the current labels after our recode

```{r, echo=FALSE}

d3 %>% 
  labelled::val_labels()

```

We no longer have any cases with a value of 3 since they have been recoded to 0. So our label for 3 is no longer necessary. Also our newly recoded value of 0 has no label.

* Note: Before adding our new value label for 0, we remove the value label for 3 using `labelled::remove_value_labels()`

```{r}

d3 <- d3 %>%
  labelled::remove_value_labels(Var3 = 3) %>%
  labelled::add_value_labels(Var3 = c("none" = 0))

d3 %>% 
  labelled::val_labels()

```


---

#### Function: `labelled()`

---

**1\. Add the same value labels to multiple variables with similar names (`Var1`, `Var2`, and `Var3`)**

Review the data (d3)

```{r, echo=FALSE}

source("data.R")
d3

```

Add the same value labels to multiple variables

* Note: We are modifying existing variables using `dplyr::mutate()`

* Note: We are using `dplyr::across()` to apply a transformation to select columns and adding the `tidyselect` selection helper `starts_with()` to select variables that meet our criteria

* Note: Other `tidyselect` selection helpers include `ends_with()`, `contains()`, `matches()`, and `num_range()`, and more.

* Note: Parentheses are *not* required around labels here, although adding them will not hurt.

```{r}

d3 <- d3 %>%
  dplyr::mutate(across(
    tidyselect::starts_with("Var"),
    ~ labelled::labelled(., labels = c(disagree = 1, neutral = 2, agree = 3))
  ))

d3 %>% 
  labelled::val_labels()

```

**2\. Add the same value labels to all variables**

Review the data (d3)

```{r, echo=FALSE}

source("data.R")
d3

```

Add the same value labels to all variables

* Note: We are modifying existing variables using `dplyr::mutate()`

* Note: We are using `dplyr::across()` to apply a transformation to select columns and adding the `tidyselect` selection helper `everything()` to select all variables

* Note: Other `tidyselect` selection helpers include `ends_with()`, `contains()`, `matches()`, and `num_range()`, and more.

* Note: Parentheses are *not* required around labels here, although adding them will not hurt.

```{r}

d3 <- d3 %>%
  dplyr::mutate(dplyr::across(
    tidyselect::everything(),
    ~ labelled::labelled(., labels = c(disagree = 1, neutral = 2, agree = 3))
  ))

d3 %>%
  labelled::val_labels()

```


Return to [Label Data](https://github.com/Cghlewis/data-wrangling-functions/wiki/Label Data)