---
title: "Add missing value labels"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```


### Package: labelled

**Note**: Using the `labelled` package, adding missing label values will not allow these values to be treated as NA in R. However, they will be labelled as missing when you export to a file type such as .sav and will be treated as NA in those files.

---

#### Function: `set_na_values()`

---

**1\.Add a label for one missing value (-999), for one variable (Var3).**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Add NA label for -999

```{r}

d1 <- d1 %>% labelled::set_na_values(Var3=-999)

labelled::na_values(d1$Var3)

```


**2\. Add labels for multiple missing values (-999, 0), for one variable (Var3)**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Add NA labels for -999 and 0

```{r}

d1 <- d1 %>% labelled::set_na_values(Var3=c(-999,0))

labelled::na_values(d1$Var3)

```


**3\. Add labels to multiple variables (Var2 and Var3) with the same missing value**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Add NA labels for -999 and 0

```{r}

d1 <- d1 %>% labelled::set_na_values(Var2 = -999, Var3= -999)

labelled::na_values(d1)

```

If I wanted to just review NA value labels for specific variables, I can select my variables of interest using `dplyr::select()` and then iterate the `labelled::na_values()` function over those variables using the `purrr::map()` function

```{r}

d1 %>%
  dplyr::select(Var2, Var3) %>%
  purrr::map(labelled::na_values)

```

