---
title: "Add NA value labels"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```


### Package: labelled

**Note**: Using the `labelled` package, adding NA label values will not allow these values to be treated as NA in R. However, they will be labelled as NA when you export to a file type such as .sav and will be treated as NA in those files.

---

#### Function: `labelled::set_na_values()`

---

**1\.Add a label for one variable, one NA value**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Add NA label

```{r}

d1 <- d1 %>% labelled::set_na_values(Var3=-999)

labelled::na_values(d1$Var3)

```


**2\. Add a label for one variable, multiple NA values**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Add NA labels

```{r}

d1 <- d1 %>% labelled::set_na_values(Var3=c(-999,0))

labelled::na_values(d1$Var3)

```


---

#### Function: `labelled::labelled_spss()`

---

**1\. Label NA values for multiple variables that have the same NA values**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Add NA labels

```{r}

d1 <- d1 %>% 
  dplyr::mutate_at(vars(Var2:Var3), 
            ~labelled::labelled_spss(., na_values=-999))

labelled::na_values(d1)

```
