---
title: "Add value labels"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: labelled

---

#### Function: `labelled::set_value_labels()`

---

**1\. Add value labels for one or more variables**

Review the data (d3)

```{r, echo=FALSE}

source("data.R")
d3

```

Add value labels

```{r}

d3 <- d3 %>% labelled::set_value_labels(
  Var2= c("not interested"=1, "mildly interested"=2,"very interested"=3),
                                      Var3=c("none"=1, "reduced-price"=2, "free"=3))

labelled::val_labels(d3)

```

**2\. Create label, then add label to vars**

Review the data (d3)

```{r, echo=FALSE}

source("data.R")
d3

```

Add value labels

```{r}

lab <- c("disagree"=1, "neutral"=2, "agree"=3)

d3 <- d3 %>% labelled::set_value_labels(Var2=lab, Var3=lab)

d3 %>% labelled::val_labels()

```

---

#### Function: `labelled::add_value_labels()`

---

**1\. Update or add one or more value labels without replacing/removing the other value labels**

Review the data (d3)

```{r, echo=FALSE}

source("data.R")
d3

```

View current labels for Var3

```{r, echo=FALSE}

d3 <- d3 %>% labelled::set_value_labels(
  Var3=c("none"=1, "reduced-price"=2, "free"=3))

d3 %>% labelled::val_labels()

```

Update the label for *1* in Var3

```{r}

d3 <- d3 %>% labelled::add_value_labels(Var3=c(paid=1))

d3 %>% labelled::val_labels()

```

---

#### Function: `labelled::labelled()`

---

**1\. Add the same value labels to multiple variables**

Review the data (d3)

```{r, echo=FALSE}

source("data.R")
d3

```

Add value labels

```{r}

d3 <- d3 %>% 
  dplyr::mutate_at(vars(starts_with('Var')), 
            ~labelled::labelled(., labels = c("disagree"=1, "neutral"=2, "agree"=3)))

d3 %>% labelled::val_labels()

```


