---
title: "Add value labels using a data dictionary"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>


### Package: labelled

---

#### Function: `set_value_labels()`

---

**1\. Set labels for one variable, `Var2`, using a long formatted data dictionary**

Review the data (d10)

```{r, echo=FALSE}

source("data.R")

d10

```

In this scenario, we have an existing dataset that provides the labels for our variables. 

Review our data dictionary (labels)

Note that this dataset is not in the usual data dictionary format I would think of (with each row corresponding to one variable). In this case, this would be a specialized long format type of codebook that you may have received or created solely for the purpose of working with value labels.

```{r, echo=FALSE}

source("data.R")

labels

```


Add value labels

* Note: First we need to create a named character vector. We can do this using `dplyr::pull()`. Note that the `label` variable is provided second. This is important because the second variable will become the "names".

```{r}

var_labels <- labels %>%
  dplyr::filter(var == "Var2") %>%
  dplyr::pull(value, label) 

```

```{r}

var_labels

```

We can also do this using `tibble::deframe()`. You will get the same result. 

HOWEVER, do note that the order that you list variables is **different**. In this method, the label (which will be the "names") will need to be provided first. 

```{r, eval = FALSE}

var_labels <- labels %>%
  dplyr::filter(var == "Var2") %>%
  dplyr::select(label, value) %>%
  tibble::deframe()

```

Now we can use this vector to add our variable labels using the argument *.labels*.

```{r}

d10 <- d10 %>%
  labelled::set_value_labels(Var2 = var_labels)

labelled::val_labels(d10)

```

### Package: labelled

---

#### Function: `labelled()`

---

<br>

**1\. Set labels for multiple variables (`Var2` and `Var3`) with the same label values using a long formatted data dictionary**

Review the data (d5)

```{r, echo=FALSE}

source("data.R")

d5

```

This time there are several variables that have the same labels so we want to apply our labels across multiple variables.

Review our data dictionary (labels2)

```{r, echo=FALSE}

source("data.R")

labels2

```

Since labels are repeated in our data dictionary, filter to just one of the variables that have the labels we need.

```{r}

var_labels <- labels2 %>%
  filter(var == "Var2") %>%
  dplyr::pull(value, label) 

```

Or you could remove duplicate labels using `dplyr::distinct()`

```{r, eval = FALSE}

var_labels <- labels2 %>%
  distinct(label, .keep_all = TRUE) %>%
  dplyr::pull(value, label) 

```


```{r}

d5 <- d5 %>% 
  dplyr::mutate(dplyr::across(Var2:Var3, 
            ~labelled::labelled(., labels = var_labels)))

d5 %>% 
  labelled::val_labels()

```

**2\. Set labels for multiple variables (`Var2` and `Var3`) with the same label values using a wide formatted data dictionary**

Review the data (d5)

```{r, echo=FALSE}

source("data.R")

d5

```

Here I have a data dictionary that is formatted in wide format (which is more common in my work). So while I can't necessarily pull the labels from a dictionary that is formatted this way, I can at least pull all variables that have the same labels, and label them at the same time, saving me time, especially if there are many variables with the same labels.

Review our data dictionary (dict3)

```{r, echo=FALSE}

source("data.R")

dict3

```

Pull our all variable names that are recoded as yes = 1 and 2 = no.

```{r}

dict_yesno <- dict3 %>%
  filter(values == "1 = yes| 2 = no") %>%
  pull(var)

```

Let's see what variables were pulled

```{r}

dict_yesno

```


Now we can apply our value labels to those variables using `dplyr::all_of()`.

```{r}

d5 <- d5 %>% 
  dplyr::mutate(dplyr::across(all_of(dict_yesno), 
            ~labelled::labelled(., labels = 
                                  c("yes" = 1,
                                    "no" = 2))))

d5 %>% 
  labelled::val_labels()


```



Return to [Label Data](https://github.com/Cghlewis/data-wrangling-functions/wiki/Label Data)