---
title: "Convert values to labels"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>


### Package: labelled

---

#### Function: `to_factor()`

---

**1\. Convert a labelled numeric variable (Var2 and Var3) to a factor variable (exchange the numeric values for the labels)**

Review the data (d3)

```{r, echo=FALSE}

source("data.R")
d3

```


Add labels

```{r}

d3 <- d3 %>% labelled::set_value_labels(
  Var2= c("not interested"=1, "mildly interested"=2,"very interested"=3),
  Var3=c("none"=1, "reduced-price"=2, "free"=3))

```

Convert to factor

```{r}

d3 %>% labelled::to_factor()


```

**2\. If one of your values is not labelled, turn that value to NA**

Review the data (d3)

```{r, echo=FALSE}

source("data.R")
d3

```


Add value labels to only some values in Var3

```{r}

d3 <- d3 %>% labelled::set_value_labels(
  Var2= c("not interested"=1, "mildly interested"=2,"very interested"=3),
  Var3=c("none"=1, "free"=3))

```

Convert to factor by adding the argument *nolabel_to_na=TRUE* to turn unlabelled values to NA. 

* Note: If you do not add this argument and you leave it as the default *nolabel_to_na=FALSE*, unlabelled values will return as their numeric values rather than NA.

```{r}

d3 %>% labelled::to_factor(nolabel_to_na=TRUE)

```


### Package: rio

---

#### Function: `factorize()`

---

**1\. Convert labelled numeric variables (Var2 and Var3) to factor variables (exchange the numeric values for the labels)**

Review the data (d3)

```{r, echo=FALSE}

source("data.R")
d3

```


Add labels

```{r}

d3 <- d3 %>% labelled::set_value_labels(
  Var2= c("not interested"=1, "mildly interested"=2,"very interested"=3),
  Var3=c("none"=1, "reduced-price"=2, "free"=3))

```

Convert to factor

```{r}

d3 %>% rio::factorize()


```

---

#### Function: `rio::characterize()`

---

**1\. Convert labelled numeric variables (Var2 and Var3) to character variables (exchange the numeric values for the labels)**

Review the data (d3)

```{r, echo=FALSE}

source("data.R")
d3

```


Add labels

```{r}

d3 <- d3 %>% labelled::set_value_labels(
  Var2= c("not interested"=1, "mildly interested"=2,"very interested"=3),
  Var3=c("none"=1, "reduced-price"=2, "free"=3))

```

Convert to character

```{r}

d3 %>% rio::characterize()

```



### Package: haven

---

#### Function: `as_factor()`

---

**1\. Convert labelled numeric variables (Var2 and Var3) to factor variables (exchange the numeric values for the labels)**

Review the data (d3)

```{r, echo=FALSE}

source("data.R")
d3

```


Add labels

```{r}

d3 <- d3 %>% labelled::set_value_labels(
  Var2= c("not interested"=1, "mildly interested"=2,"very interested"=3),
  Var3=c("none"=1, "reduced-price"=2, "free"=3))

```

Convert Var2 and Var3 to factor. 

* Note: The default for `haven::as_factor()` is to assign the labels in place of the number for any labelled value. However, using the *levels* argument, you can choose other options such as pasting together the label and value using the argument *levels = "both"*. You can see all arguments by typing *?as_factor* in the console and choosing the `haven` package.

```{r}

d3 %>% haven::as_factor()


```

You can also specifically convert only labelled variables in your data frame this way (even though `haven::as_factor` will only convert variables with labels anyway) using the functions `dplyr::mutate_if()` as well as `haven::is.labelled()`.

```{r, echo=FALSE}

d3 <- d3 %>% labelled::set_value_labels(
  Var2= c("not interested"=1, "mildly interested"=2,"very interested"=3),
  Var3=c("none"=1, "reduced-price"=2, "free"=3))

```


```{r}

d3 %>% dplyr::mutate_if(haven::is.labelled, haven::as_factor)

```



**2\. Select a specific labelled numeric variable (Var2) to convert to a factor variable (exchange the numeric values for the labels)**

Review the data (d3)

```{r, echo=FALSE}

source("data.R")
d3

```

Add labels

```{r}

d3 <- d3 %>% labelled::set_value_labels(
  Var2= c("not interested"=1, "mildly interested"=2,"very interested"=3))

```


Convert Var2 to a factor by including the `dplyr::mutate()` function to create a new variable (which replaces the old variable because we named it the same).

```{r}

d3 %>% dplyr::mutate(Var2 = haven::as_factor(Var2))


```

