---
title: "Converting values to labels"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>


### Package: labelled

---

#### Function: `labelled::to_factor()`

---

**1\. Convert a labelled numeric variable to a factor variable (exchange the numeric values for the labels)**

Review the data (d3)

```{r, echo=FALSE}

source("data.R")
d3

```


Add labels

```{r}

d3 <- d3 %>% labelled::set_value_labels(
  Var2= c("not interested"=1, "mildly interested"=2,"very interested"=3),
  Var3=c("none"=1, "reduced-price"=2, "free"=3))

```

Convert to factor

```{r}

d3 %>% labelled::to_factor()


```

**2\. If one of your values is not labelled, turn that value to NA**

Review the data (d3)

```{r, echo=FALSE}

source("data.R")
d3

```


Add value labels to only some values in Var3

```{r}

d3 <- d3 %>% labelled::set_value_labels(
  Var2= c("not interested"=1, "mildly interested"=2,"very interested"=3),
  Var3=c("none"=1, "free"=3))

```

Convert to factor

```{r}

d3 %>% labelled::to_factor(nolabel_to_na=TRUE)

```


### Package: rio

---

#### Function: `rio::factorize()`

---

**1\. Convert a labelled numeric variable to a factor variable (exchange the numeric values for the labels)**

Review the data (d3)

```{r, echo=FALSE}

source("data.R")
d3

```


Add labels

```{r}

d3 <- d3 %>% labelled::set_value_labels(
  Var2= c("not interested"=1, "mildly interested"=2,"very interested"=3),
  Var3=c("none"=1, "reduced-price"=2, "free"=3))

```

Convert to factor

```{r}

d3 %>% rio::factorize()


```

---

#### Function: `rio::characterize()`

---

**1\. Convert a labelled numeric variable to a character variable (exchange the numeric values for the labels)**

Review the data (d3)

```{r, echo=FALSE}

source("data.R")
d3

```


Add labels

```{r}

d3 <- d3 %>% labelled::set_value_labels(
  Var2= c("not interested"=1, "mildly interested"=2,"very interested"=3),
  Var3=c("none"=1, "reduced-price"=2, "free"=3))

```

Convert to character

```{r}

d3 %>% rio::characterize()

```



### Package: haven

---

#### Function: `haven::as_factor()`

---

**1\. Convert a labelled numeric variable to a factor variable (exchange the numeric values for the labels)**

Review the data (d3)

```{r, echo=FALSE}

source("data.R")
d3

```


Add labels

```{r}

d3 <- d3 %>% labelled::set_value_labels(
  Var2= c("not interested"=1, "mildly interested"=2,"very interested"=3),
  Var3=c("none"=1, "reduced-price"=2, "free"=3))

```

Convert to factor

```{r}

d3 %>% haven::as_factor()


```
