---
title: "Import and export labelled data"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>


### Package: haven

---

#### Function: `read_sav()`

---

**1\. Read in a labelled SPSS file**

```{r, eval=FALSE}

d <- haven::read_sav("name-of-import.sav")

```


**2\. Read in an SPSS file with NA value labels**

* Note: If the file has user labelled missing values that you want to save, add the option *user_na=TRUE*. Otherwise the labelled NA values will be removed and read in as NA.

```{r, eval=FALSE}

d <- haven::read_sav("file.sav", user_na=TRUE)

```


---

#### Function: `write_sav()`

---

**1\. Export a labelled SPSS file**

```{r, eval=FALSE}

haven::write_sav(d, "name-of-export.sav")

```


### Package: rio

---

#### Function: `characterize()`

---

**1\. Export your variables with the value labels rather than the values**

* Note: You can add `rio::characterize()` when exporting other file types as well such as csv and with using other export functions such as `rio::export()`

```{r, eval=FALSE}

openxlsx::write.xlsx(rio::characterize(d), "name-of-export.xlsx")

```


Return to [Label Data](https://github.com/Cghlewis/data-wrangling-functions/wiki/Label Data)