---
title: "Labelled SPSS"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```


---
  
#### Function: `labelled_spss()`

---

**Note**: If you use this function to add your NA value labels please read this warning first.

**Warning**
  1. If prior to adding your NA values, you added *value labels* to your data and you did **not** use `labelled::labelled_spss()` to add those labels (i.e. you used something like `labelled::set_value_labels()`) then you will lose those value labels for any variable you add NA value labels to using the `labelled::labelled_spss()` function. As far as I can tell, the reason is that your variables will be converting from class *haven_labelled* to *haven_labelled_spss* and the labels do not stick. 
2. If prior to adding your NA values, you added *variable labels* to your data and you did **not** use `labelled::labelled_spss()` to add those labels (i.e. you used something like `labelled::set_variable_labels()`), you will also lose those variable labels for any variable you add NA value labels to. Again, as far as I can tell, the reason is that your variables will be converting from class *haven_labelled* or another class such as *numeric* to *haven_labelled_spss* and the labels do not stick. 

**Solution**
  1. Don't use `labelled::labelled_spss()` if you have value and variable labels. Just add NA values using `labelled::set_na_values()`. The only downside to using that function is that I haven't found a way to assign the same NA value to a ton of variables that share a common attribute (ex: numeric variables). Instead you have to list out each individual variable and assign the NA value individually. But if the number of variables you are working with is small, I would use `labelled::set_na_values()`.

If you need to assign missing values to a large number of variables and want to use `labelled::labelled_spss()`:
  2. You can solve the value label problem by both adding your value labels using `labelled::labelled_spss()`.
3. You can solve the variable label problem by either:
  a) adding your variable labels using `labelled::labelled_spss()`
b) adding your variable labels using another function **after** you have added your value and NA labels using `labelled::labelled_spss()`. Luckily adding your variable labels does not change the variable class so your value and NA labels will remain as long as you run this function after you update value and NA labels.


**1\. Label NA values for multiple variables (Var2 and Var3) that have the same NA value (-999)**
  
Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Add NA labels for -999

* Note: We are modifying existing variables using `dplyr::mutate()`

* Note: We are using `dplyr::across()` to apply a transformation to select columns

```{r}

d1 <- d1 %>% 
  dplyr::mutate(dplyr::across(c(Var2:Var3), 
                              ~labelled::labelled_spss(., na_values=-999)))

labelled::na_values(d1)

```


**2\. Label NA values as -999 for all numeric variables**
  
  Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Add NA labels for -999

* Note: We are modifying existing variables using `dplyr::mutate()`

* Note: We are using `dplyr::across()` to apply a transformation to select columns

* Note: You must wrap *is.numeric*, a predicate function (returns a true/false), in the `tidyselect` selection helper `where()`.

* Note: Other `tidyselect` selection helpers include `starts_with()`, `ends_with()`, `contains()`, `matches()`, and `num_range()`, and more.

```{r}

d1 <- d1 %>% 
  dplyr::mutate(dplyr::across(where(is.numeric), 
                              ~labelled::labelled_spss(., na_values=-999)))

labelled::na_values(d1)

```



---

#### Function: `labelled_spss()`

---

* Note: I don't personally use this function unless I plan to use `labelled::labelled_spss()` to assign the same NA value to many variables at some point. Then I will use this function so that I don't lose my value labels due to compatibility issues. See the *add na value labels* tutorial for more information.

**1\. Set value labels for one or more variables (Var2 and Var3) with the same value labels**

Review the data (d3)

```{r, echo=FALSE}

source("data.R")
d3

```

Add value labels to multiple variables

* Note: We are modifying existing variables using `dplyr::mutate()`

* Note: We are using `dplyr::across()` to apply a transformation to select columns

* Note: Parentheses are *not* required around labels here, although adding them will not hurt.

```{r}

d3 <- d3 %>%
  mutate(across(c(Var2:Var3), ~
  labelled::labelled_spss(., c(disagree=1, neutral=2, agree=3))))

d3 %>% labelled::val_labels()

```

---

#### Function: `labelled_spss()`

---

* Note: I don't personally use this function unless I plan to use `labelled::labelled_spss()` to assign the same NA value to many variables at some point. Then I will use this function so that I don't lose my variable labels due to compatibility issues. See the *add na value labels* tutorial for more information.

**1\. Set variable labels for one variable**

Review the data (d4)

```{r, echo=FALSE}

source("data.R")
d4

```

Add variable labels



```{r}

d4 <- d4 %>% 
  dplyr::mutate(dplyr::across(c(Var2), 
                              ~labelled::labelled_spss(., label= "Interest in homework")))

labelled::var_label(d4)

```



```{r, include=FALSE, echo=FALSE}

# There is no easy way to assign info to multiple different variables using labelled_spss
# You can assign value labels and missing value labels across several of the same variables, but that leaves out the variable labels then

d10 <- d10 %>% 
  dplyr::mutate(dplyr::across(c(Var2), 
                              ~labelled::labelled_spss(., label= "Quality",
                               c("not interested"=1, "mildly interested"=2, 
                                                         "very  interested"=3),
                                                       na_values=-999)))


```


Return to [Label Data](https://github.com/Cghlewis/data-wrangling-functions/wiki/labelling-data)