---
title: "Review labelled data"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>


### Package: labelled

---

#### Function: `look_for()`

---

**1\. Review all labelled information in a data dictionary format**

Let's create some labelled data

```{r, echo=FALSE}

source("data.R")

```


```{r}

d10 <- d10 %>%
  labelled::set_variable_labels (Var1 = "Name", 
                                 Var2="Interest in Homework", 
                                 Var3 ="Lunch Status") %>%
  labelled::set_value_labels(Var2= c("not interested"=1, "mildly interested"=2, 
                                     "very  interested"=3), Var3=c("none"=1, 
                                    "reduced-price"=2, "free"=3)) %>%
  labelled::set_na_values(Var2=-999, Var3=-999)

```


View the labelled data (d10)

```{r, echo=FALSE}

d10

```

Review labelled information

* Note: The argument *details=TRUE* is added to get na_values for variables

* Note: `knitr::kable()` function used to make output more readable but we lose the value labels. If you run the `labelled::look_for()` function without adding `knitr::kable()` you will see the value labels.

* Note: The function `labelled::generate_dictionary()` is another function equivalent to `labelled::look_for()`. You can use either function.

```{r}

knitr::kable(labelled::look_for(d10, details = TRUE) %>% 
               dplyr::select(variable, label, class, value_labels, na_values))

```

```{r, echo=FALSE, include=FALSE}

dictionary <- labelled::generate_dictionary(d10, details = TRUE) %>%
  select(c(variable, label, value_labels, class, na_values))

```

---

#### Function: `var_label()`

---

**1\. Review just the variable labels**

View the data (d10)

```{r, echo=FALSE}

source("data.R")

d10 <- d10 %>%
  labelled::set_variable_labels (Var1 = "Name", 
                                 Var2="Interest in Homework", 
                                 Var3 ="Lunch Status") %>%
  labelled::set_value_labels(Var2= c("not interested"=1, "mildly interested"=2, 
                                     "very  interested"=3), Var3=c("none"=1, 
                                    "reduced-price"=2, "free"=3)) %>%
  labelled::set_na_values(Var2=-999, Var3=-999)


d10
```

Review variable labels

```{r}

d10 %>% labelled::var_label()

```

---

#### Function: `val_labels()`

---

**1\. Review just the value labels**

View the data (d10)

```{r, echo=FALSE}

source("data.R")

d10 <- d10 %>%
  labelled::set_variable_labels (Var1 = "Name", 
                                 Var2="Interest in Homework", 
                                 Var3 ="Lunch Status") %>%
  labelled::set_value_labels(Var2= c("not interested"=1, "mildly interested"=2, 
                                     "very  interested"=3), Var3=c("none"=1, 
                                    "reduced-price"=2, "free"=3)) %>%
  labelled::set_na_values(Var2=-999, Var3=-999)


d10
```

Review value labels

```{r}

d10 %>% labelled::val_labels()

```

---

#### Function: `na_values()`

---

**1\. Review just the NA value labels**

View the data (d10)

```{r, echo=FALSE}

source("data.R")

d10 <- d10 %>%
  labelled::set_variable_labels (Var1 = "Name", 
                                 Var2="Interest in Homework", 
                                 Var3 ="Lunch Status") %>%
  labelled::set_value_labels(Var2= c("not interested"=1, "mildly interested"=2, 
                                     "very  interested"=3), Var3=c("none"=1, 
                                    "reduced-price"=2, "free"=3)) %>%
  labelled::set_na_values(Var2=-999, Var3=-999)


d10
```

Review NA value labels

```{r}

d10 %>% labelled::na_values()

```


### Package: sjPlot

---

#### Function: `view_df()`

---

**1\. Review all labelled information in a data dictionary format**

View the data (d10)

```{r, echo=FALSE}

source("data.R")

d10 <- d10 %>%
  labelled::set_variable_labels (Var1 = "Name", 
                                 Var2="Interest in Homework", 
                                 Var3 ="Lunch Status") %>%
  labelled::set_value_labels(Var2= c("not interested"=1, "mildly interested"=2, 
                                     "very  interested"=3), Var3=c("none"=1, 
                                    "reduced-price"=2, "free"=3)) %>%
  labelled::set_na_values(Var2=-999, Var3=-999)


d10
```

Review labelled information

* Note: You will notice that we do not get NA labels here like we do using the `labelled::look_for()` function. We also do not get the variable class. Instead we can ask for the variable type by adding the argument *show.type=TRUE*, which shows us that Var2 for instance is numeric but does not let us know that the class is haven_labelled_spss.

```{r}

d10 %>% sjPlot::view_df(show.type=TRUE)

```

