---
title: "Review labelled data"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>


### Package: labelled

---

#### Function: `labelled::look_for()`

---

**1\. Review all labelled information in a data dictionary format**

View the labelled data (d10)

```{r, echo=FALSE}

source("data.R")

d10 <- d10 %>%
  labelled::set_variable_labels (Var1 = "Name", 
                                 Var2="Interest in Homework", 
                                 Var3 ="Lunch Status") %>%
  labelled::set_value_labels(Var2= c("not interested"=1, "mildly interested"=2, 
                                     "very  interested"=3), Var3=c("none"=1, 
                                    "reduced-price"=2, "free"=3)) %>%
  labelled::set_na_values(Var2=-999, Var3=-999)


d10
```

Review labelled information

```{r}

knitr::kable(labelled::look_for(d10, details = TRUE) %>% 
               select (variable, label, class, value_labels, na_values))

```

---

#### Function: `labelled::var_label()`

---

**1\. Review just variable labels**

View the data (d10)

```{r, echo=FALSE}

source("data.R")

d10 <- d10 %>%
  labelled::set_variable_labels (Var1 = "Name", 
                                 Var2="Interest in Homework", 
                                 Var3 ="Lunch Status") %>%
  labelled::set_value_labels(Var2= c("not interested"=1, "mildly interested"=2, 
                                     "very  interested"=3), Var3=c("none"=1, 
                                    "reduced-price"=2, "free"=3)) %>%
  labelled::set_na_values(Var2=-999, Var3=-999)


d10
```

Review variable labels

```{r}

d10 %>% labelled::var_label()

```

---

#### Function: `labelled::val_labels()`

---

**1\. Review just value labels**

View the data (d10)

```{r, echo=FALSE}

source("data.R")

d10 <- d10 %>%
  labelled::set_variable_labels (Var1 = "Name", 
                                 Var2="Interest in Homework", 
                                 Var3 ="Lunch Status") %>%
  labelled::set_value_labels(Var2= c("not interested"=1, "mildly interested"=2, 
                                     "very  interested"=3), Var3=c("none"=1, 
                                    "reduced-price"=2, "free"=3)) %>%
  labelled::set_na_values(Var2=-999, Var3=-999)


d10
```

Review value labels

```{r}

d10 %>% labelled::val_labels()

```

---

#### Function: `labelled::na_values()`

---

**1\. Review just NA value labels**

View the data (d10)

```{r, echo=FALSE}

source("data.R")

d10 <- d10 %>%
  labelled::set_variable_labels (Var1 = "Name", 
                                 Var2="Interest in Homework", 
                                 Var3 ="Lunch Status") %>%
  labelled::set_value_labels(Var2= c("not interested"=1, "mildly interested"=2, 
                                     "very  interested"=3), Var3=c("none"=1, 
                                    "reduced-price"=2, "free"=3)) %>%
  labelled::set_na_values(Var2=-999, Var3=-999)


d10
```

Review NA value labels

```{r}

d10 %>% labelled::na_values()

```

