---
title: "Delete files"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)

library(fs)
library(here)
library(tidyverse)
```

<br>

### Package: fs

---

#### Function: `file_delete()`

---

**1\. Delete an observation file ("summary_obs.xlsx") from a folder ("obs_2022-03-15").**


Let's review the current files in both folders.

Files in "obs_2022-03-15"

```{r}

base::basename(fs::dir_ls(here::here("file-system", "obs_2022-03-15")))

```

Now let's delete "summary_obs.xlsx"

```{r}
    
fs::file_delete(path = here::here("file-system", "obs_2022-03-15", "summary_obs.xlsx"))

```

Now let's view our files in the folder.

```{r}

base::basename(fs::dir_ls(here::here("file-system", "obs_2022-03-15")))

```

**2\. Delete ALL files from a folder ("obs_2022-03-15").**

If we wanted to delete ALL files in the folder "obs_2022-03-15", we could grab all our file names and pass that in our path.

```{r}

files <- base::basename(fs::dir_ls(here::here("file-system", "obs_2022-03-15")))
    
fs::file_delete(path = here::here("file-system", "obs_2022-03-15", files))

```

Now let's view our files in the folder.

They are all gone.

```{r}

base::basename(fs::dir_ls(here::here("file-system", "obs_2022-03-15")))

```


Another option is to *delete the entire folder* by simply not listing any specific files. 

```{r, eval = FALSE}
    
fs::file_delete(path = here::here("file-system", "obs_2022-03-15"))

```

Return to [File System](https://github.com/Cghlewis/data-wrangling-functions/wiki/File System)