---
title: "Rename files"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)

library(fs)
library(here)
library(tidyverse)
```

<br>

### Package: fs

---

#### Function: `file_move()`

---


**1\. Rename one file ("summary_obs.xlsx") in the folder "obs_2022-03-15".**


Let's review the current files

```{r}

base::basename(fs::dir_ls(here::here("file-system", "obs_2022-03-15")))

```

Rename "summary_obs.xlsx" to "summary_tch_obs.xlsx"

```{r}

fs::file_move(path = here::here("file-system", "obs_2022-03-15", "summary_obs.xlsx"),
              new_path = here::here("file-system", "obs_2022-03-15", "summary_tch_obs.xlsx"))

```

Let's review the files now

```{r}

base::basename(fs::dir_ls(here::here("file-system", "obs_2022-03-15")))

```

**2\. Rename multiple files in the folder "obs_2022-03-15".**

Let's review the current files

```{r, echo = FALSE}

fs::file_move(here::here("file-system", "obs_2022-03-15", "summary_tch_obs.xlsx"),
              here::here("file-system", "obs_2022-03-15", "summary_obs.xlsx"))

```

```{r}

base::basename(fs::dir_ls(here::here("file-system", "obs_2022-03-15")))

```

Rename obs_a, obs_b, and obs_c to obs1, obs2, obs3.

(01) First let's grab the files we want to rename and put them in an object "files_old".

I could definitely write out all of the file names and put them into a vector, like this.

```{r, eval = FALSE}

files_old <- c(here::here("file-system", "obs_2022-03-15","obs_a.xlsx"),
               here::here("file-system", "obs_2022-03-15","obs_b.xlsx"),
               here::here("file-system", "obs_2022-03-15","obs_c.xlsx"))

```

But if I had tons of files, this would take forever. We've already seen that `fs::dir_ls()` grabs the names of all of our files in our directory so we can use this function to more efficiently grab our existing files.

Since there is a "summary_obs.xlsx" file that I do not want to rename, I am going to add the `fs::dir_ls()` argument *glob* here which allows you to grab files/folders using a globbing pattern. That pattern is passed to `base::grep()` to filter paths. The "?" in globbing matches one character.

```{r}

files_old <- fs::dir_ls(here("file-system", "obs_2022-03-15"), glob = "*obs_?.xlsx")

```

You could instead use the *regexp* argument and use a regular expression pattern to grab the files you want. This pattern is also passed to `base::grep()` to filter paths. The "[a-z]" in regex matches one alphabetic character.

```{r}

files_old <- fs::dir_ls(here("file-system", "obs_2022-03-15"), regexp = "*obs_[a-z].xlsx")

```

If I wanted to grab all files, I could simply write the code like this.

```{r, eval = FALSE}

files_old <- fs::dir_ls(here::here("file-system", "obs_2022-03-15"))

```


(02) Now let's create the new names for our files and put them in an object "files_new".

Again, I could write out the new names for all of our files but if we are renaming many files, we should find a more efficient way to rename.

* Note: I am using `base::paste0()` to paste the word "obs" with the file number and ".xslx".
* Note: I am using `base::length()` to create numbers for the number of files that exist in the folder.

```{r}

files_new <- here::here("file-system", "obs_2022-03-15",(base::paste0( "obs", 1:length(files_old), ".xlsx")))

```


(03) Now let's rename the files

```{r}

fs::file_move(path = files_old, new_path = files_new)


```


Let's view the new names.

```{r}

base::basename(fs::dir_ls(here::here("file-system", "obs_2022-03-15")))

```

If I didn't want to create new names from scratch but simply wanted to replace a part of the existing name, I could consider something like this instead, using `stringr::str_replace()` to replace any "_letter" with the file number.

```{r, eval = FALSE}

files_old <- fs::dir_ls(here::here("file-system", "obs_2022-03-15"), glob='obs_*')

file_number <- as.character(1:length(files_old))
  
files_new <- stringr::str_replace(files_old, "_[a-z]", file_number)
  
fs::file_move(path = files_old, new_path = files_new)

```

Return to [File System](https://github.com/Cghlewis/data-wrangling-functions/wiki/File System)