---
title: "Import multiple files and combine"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)

library(tidyverse)

```


### Package: purrr

---

#### Function: `map_dfr()`

---

<br>

**1a\. Read in all Excel files from the "attendance" folder and bind them together. Note that an asterisk was used to denote NAs and we need to notify R of that.**

Review the files in our directory

* Note: I am currently working in an R project. I used the `here::here()` function which implicitly sets my directory to the top level (root directory) of my current project. Then any subsequent folders can be listed in descending order. In this case my folder "attendance" is within another folder called "import-data" which is located in the top level of my project directory. You could also just list your entire file path.

```{r, eval = FALSE}

fs::dir_ls(path = here::here("import-data", "attendance"))

```

```{r, echo = FALSE}

d <- fs::dir_ls(path = here::here("import-data", "attendance"))

stringr::str_replace_all(
  d, "C:/Users/Crystal/Desktop/desktop/R/Working R Stuff/data-wrangling-functions/", "C:/Users/Me/project/") 

```

We see that there are 4 Excel files in the directory and one csv file. We only want to import the Excel files.

We can read in all of our files using `fs::dir_ls()` and bind them together using `purrr::map_dfr()`

* Note: By adding the `dir_ls()` argument *glob = "\*.xlsx"* you are saying you only want to read in the names of all Excel files with the extension ".xlsx". In this case, while there are 5 files in the directory, only 4 of them were Excel files, one file was csv format.

* Note: While basic `purrr::map()` always returns a list, `purrr::map_dfr()` combines and returns tibbles (data frames) by stacking them on top of each other (using `dplyr::bind_rows`). The main arguments you include are *.x*, a list or atomic vector (files to read), and *.f*, which in this scenario is a function.

* Note: Below I have not added a "." for the *.x* argument to denote that I am referring to the character vector we created in the previous step. In this instance, it is implied.

* Note: For this example I am using `readxl::read_excel()` for the *.f* argument. If you wanted to read in other file types, you would simply change the function to the one required to read in your file type.

* Note: If not all files you read in have the same variables, the appended file will include all variables and whatever data is available for those variables.

* Note: If files contain the same variable names but the variable classes aren't the same for a variable (ex: `V1` = character in one file and `V1` = numeric in another file) then datasets cannot be combined. In the case of `readxl::read_excel()`, you could add the *col_types* argument to make that variable type the same across all files. If you are unsure what types of variables exist in each dataset, you can use `janitor::compare_df_cols()` to see an output of each data frames' variables and variable types. By adding the argument *return = "mismatch"* the output will only show any variables that are different types across data frames.

* Note: In the function `readxl::read_excel()` I added the argument *na=* because I know the files I am reading in use "*" to denote missing values.

```{r}

d <- fs::dir_ls(path = here::here("import-data", "attendance"), glob ="*.xlsx") %>%
  purrr::map_dfr(readxl::read_excel, na="*")

head(d, n = c(10,5))

```

* Note: If you did not want to use `here::here()` you could just list the full file path.

```{r, eval=FALSE}

d <- fs::dir_ls(path = "C:/Users/Name/project/import-data/attendance", 
                glob ="*.xlsx") %>%
  purrr::map_dfr(readxl::read_excel, na="*")

```

* Update 2024-06-13: `purrr::map_dfr()` is now superseded. You can still use it but the preferred method of binding files now is using `purrr::list_rbind()` from `purrr` version 1.0.2

```{r eval = FALSE}

d <-
  fs::dir_ls(path = here::here("import-data", "attendance"),
             glob = "*.xlsx") %>%
  purrr::map(readxl::read_excel, na = "*") %>%
  purrr::list_rbind()

```

**1b\. Add an identifier to differentiate which rows came from which files, by adding the argument ".id" and have it equal to whatever you want that new variable to be called. This will give you the full file name as the row identifier to differentiate which rows came from which file.**

Add an identifier using the argument *.id=* and naming the variable whatever you want, in this case "source"

```{r}

d <- fs::dir_ls(path = here::here("import-data", "attendance"), glob ="*.xlsx") %>%
  map_dfr(readxl::read_excel, na="*", .id = "source")

```

```{r, echo=FALSE}

d <- d %>% 
  mutate(source = stringr::str_replace_all(
  source, "C:/Users/Crystal/Desktop/desktop/R/Working R Stuff/data-wrangling-functions/", "C:/Users/Me/project/")) 

head(d, n=c(10,5))

```

* Update 2024-06-13: `purrr::map_dfr()` is now superseded. You can still use it but the preferred method of binding files is using `purrr::list_rbind()` from `purrr` version 1.0.2, and then adding the argument *names_to* rather than *.id*

```{r eval = FALSE}

d <-
  fs::dir_ls(path = here::here("import-data", "attendance"),
             glob = "*.xlsx") %>%
  purrr::map(readxl::read_excel, na = "*") %>%
  purrr::list_rbind(names_to = "source")

```

**2\. Read in all Excel files except one from the directory and bind them together.**

Review the files in our "attendance" folder again

```{r}

base::dir(path = here::here("import-data", "attendance"))

```

Read in all files except "school-5_attendance.xlsx"

* Note: You do not have to write the entire file name "school-5_attendance.xlsx" because `stringr::str_subset()` includes an argument to detect a pattern.

```{r, eval = FALSE}

d <- fs::dir_ls(path = here::here("import-data", "attendance"), 
                glob ="*.xlsx") %>%
  stringr::str_subset(., "school-5_attendance", negate = TRUE) %>%
  purrr::map_dfr(readxl::read_excel, .id = "source")

```

If you wanted to exclude more than one file, you can add a "|" between the file names.

```{r, eval = FALSE}

d <- fs::dir_ls(path = here::here("import-data", "attendance"), 
                glob ="*.xlsx") %>%
  stringr::str_subset(., "school-4_attendance|school-5_attendance", negate = TRUE) %>%
  purrr::map_dfr(readxl::read_excel, .id = "source")


```

### Package: readr

---

#### Function: `read_csv()`

---


**1\. Read in a vector of CSV file names and bind them.**

Review the files in our "csv_data" folder

```{r}

base::dir(path = here::here("import-data", "csv_data"))

```

**Note:** The *file* argument for `read_csv()` allows you to provide "Either a path to a file, a connection, or literal data (either a single string or a raw vector). Since it allows vectors, you directly provide paths to multiple files as a vector in the *file* argument. `read_excel()` does not allows this.

If you want, you can add an identifier using the argument *id=* and name the variable whatever you want, in this case "source"

```{r, eval = FALSE}

readr::read_csv(c("csv_data/school-2_attendance.csv",
           "csv_data/school-3_attendance.csv"), id = "source")

```

```{r, echo=FALSE}

read_csv(c("csv_data/school-2_attendance.csv",
           "csv_data/school-3_attendance.csv"), id = "source") %>%
  filter(!is.na(YEAR))

```


Return to [Import Files](https://github.com/Cghlewis/data-wrangling-functions/wiki/Import Files)