---
title: "Import from a url"
output: 
  html_document:
    anchor_sections: FALSE
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

### Package: readxl

---
  
#### Function: `readxl::read_excel()`
  
---
  
**1\. Read in an Excel file**

A url directly does not work as a path for the `readr::read_excel()` function. Therefore reading an Excel file directly from a url is a two part process.

First, you need to download the file to a temporary directory on your machine. From your temporary file you can read the file into R. 

You can use the `curl` package to download the file. 
First assign the url to an object.
Then assign your temporary directory to an object.
Then use these arguments in the `curl::curl_download()` function.

```{r}

url <- "https://apps.dese.mo.gov/MCDS/FileDownloadWebHandler.ashx?filename=4d81bf0a-32f5Building%20Level%20Per%20Pupil%20Expenditures%202021.xlsx"

tmp <- tempfile()

curl::curl_download(url, tmp)

```

You will now have the file downloaded in a temporary directory.

Next you can read in the Excel file same as you would as reading an Excel file from any directory on your computer.

Type *?read_excel* in the console for available arguments

```{r}

d <- readxl::read_excel(tmp)

```

### Package: readr

---
  
#### Function: `readr::read_csv()`
  
---
  
**1\. Read in a csv file**

Unlike readxl::read_excel(), you can actually read in a url directly with readr::read_csv().

Type *?read_csv* in the console for available arguments

```{r, eval=FALSE}

d <- readr::read_csv("https://stacks.stanford.edu/file/druid:db586ns4974/district%20gaps%20%28pooled%20year%2C%20grade%20and%20sub%29_v1_1.csv")

```
