---
title: "Import sheets as separate data frames"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)

library(tidyverse)

```


### Package: purrr

---

#### Function: `map()`

---

**1\. Read in all sheets of an Excel file as separate data frames.**

View the sheets in our workbook

```{r}

path <- (here::here("import-data", "data", "all-school_attendance.xlsx"))

path %>% 
readxl::excel_sheets()

```

There are several ways to read in each sheet as a data frame. 

First you can read in one sheet at a time using `readxl::read_excel` and adding the argument *sheet=*, which allows you to call the name of the sheet OR the position.

```{r, eval=FALSE}

df1 <- readxl::read_excel(path, sheet = "Sheet1")

df2 <- readxl::read_excel(path, sheet = "Sheet2")

df3 <- readxl::read_excel(path, sheet = "Sheet3")

```

Second, you can read grab all of the Excel sheets using `readxl::excel_sheets()` and then iterate our function (`readxl::read_excel()`) over each sheet using `purrr::map()` to create a list of data frames.

We can then use the `$` to select each sheet from the list by name.

* Note: Using `purrr::set_names()` we are creating a named list (giving each sheet a name)

```{r, eval=FALSE}

data <- path %>% 
  readxl::excel_sheets() %>% 
  purrr::set_names() %>% 
  purrr::map(readxl::read_excel, path = path)

df1 <- data$Sheet1

df2 <- data$Sheet2

df3 <- data$Sheet3

```

But if we have many sheets, the method above can be time consuming. A better option may be to use `base::list2env()`again.

```{r, echo = FALSE}

rm(list = ls())

```

```{r}

# Create our list of data frames
my_file_list <- here::here("import-data", "data", "all-school_attendance.xlsx") %>% 
  readxl::excel_sheets() %>% 
  purrr::set_names() %>% 
  purrr::map(readxl::read_excel, path = here::here("import-data", "data", "all-school_attendance.xlsx"))

# Create environment variables from our list
list2env(my_file_list, envir=.GlobalEnv)

# Review all objects in environment
ls()

```


Return to [Import Files](https://github.com/Cghlewis/data-wrangling-functions/wiki/Import Files)