---
title: "Import sheets and bind into one data frame"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)

library(tidyverse)

```


### Package: purrr

---

#### Function: `map_dfr()`

---

**1\. Read in all sheets of an Excel file and bind them into one large data frame.**

View the sheets in our workbook

```{r}

path <- (here::here("import-data", "data", "all-school_attendance.xlsx"))

path %>% 
readxl::excel_sheets()

```

We first use `readxl::excel_sheets()` to grab the sheet names and put them into a character vector.
Then we set the names of the sheets in the vector using `purrr::set_names()`. 
Last we read the sheets in using `readxl::read_excel()` and bind them using `purrr::map_dfr()`.
We add the `purrr::map_dfr()` argument *.id* to add a new variable that we call `source` so we know which sheet each row comes from.

```{r}

data <- path %>% 
  readxl::excel_sheets() %>% 
  purrr::set_names() %>% 
  purrr::map_dfr(readxl::read_excel, path = path, .id = "source")

head(data)


```

* Update 2024-06-13: `purrr::map_dfr()` is now superseded. You can still use it but the preferred method of binding files now is using `purrr::list_rbind()` from `purrr` version 1.0.2. Similar to `dplyr::bind_rows()`, used in `purrr::map_dfr()`, `purrr::list_rbind()` matches by variable names and requires variables to be the same types. It also allows some variables to be missing from some sheets (filling with NA as needed).

```{r}

data <- path %>% 
  readxl::excel_sheets() %>% 
  purrr::set_names() %>%
  purrr::map(readxl::read_excel, path = path) %>%
  purrr::list_rbind(names_to = "sheet_name")

head(data)

```


Return to [Import Files](https://github.com/Cghlewis/data-wrangling-functions/wiki/Import Files)