---
title: "Collapse categories"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: dplyr

---

#### Function: `if_else()`

---

**Note:** `dplyr::if_else()` will evaluate in order, so proceed from most specific to most general. The function will not write over a previous evaluation even if something meets a new criteria. 

**Note:** `dplyr::if_else()` will not evaluate NAs as TRUE or FALSE and therefore will return NAs. If you want to specifically evaluate NAs you will need to call out NAs.

<br>

**1\. Collapse all "Google Meet" in character var (online_platform) into one category**

Review the data (d6)

```{r, echo=FALSE}

source("data.R")
d6

```

Collapse all spellings of Google Meet into one uniform category

* Note: We are recoding into a new variable using `dplyr::mutate()` and naming the new variable a different name than the original. It keeps both the new and old versions of the variable. 

* Note: Using `stringr::str_detect()` we can detect different spellings of google meet. I enter a pattern ("google|meet") that means to look for any value in online_platform that contains the word "google" OR "meet" and recategorize those to "Google Meet"

```{r}

d6 %>%  
  dplyr::mutate(online_platform_new=dplyr::if_else(
  stringr::str_detect(online_platform, "google|meet"), "Google Meet", online_platform))

```

We can also do this using `%in%` instead of `stringr::str_detect()`

```{r}

d6 %>%  
  dplyr::mutate(online_platform_new=dplyr::if_else(
  online_platform %in% c("google","meet", "Google Meet"), "Google Meet", online_platform))

```


**2\. Collapse all "Google Meet" in character var (online_platform) into one category and address case sensitivity**

Review the data (d12)

```{r, echo=FALSE}

source("data.R")
d12

```

Collapse all spellings of Google Meet into one uniform category


* Note: Using `stringr::str_detect()` we can detect different spellings of google meet. It is important to note that this function is case sensitive. As we saw above in the data, the "google|meet" pattern we used before will no longer catch our upper case "Google". So there are a number of ways we can deal with this. The first way we will try is by adding stringr::regex() and adding the argument *ignore_case = TRUE*.

```{r}

d12 %>%  
  dplyr::mutate(online_platform_new=dplyr::if_else(
  stringr::str_detect(online_platform, stringr::regex("google|meet", ignore_case = TRUE)), "Google Meet", online_platform))

```

We can also do this by adding (?i) to the pattern, which means match case insensitive. 

```{r}

d12 %>%  
  dplyr::mutate(online_platform_new=dplyr::if_else(
  stringr::str_detect(online_platform, "(?i)google|meet"), "Google Meet", online_platform))

```


**3\. Collapse a numeric variable (Var2) into discrete character categories**

Review the data (d8)

```{r, echo=FALSE}

source("data.R")
d8

```

Collapse Var2 into low, medium, high categories.

* Note: We are using a nested if-else statement here to meet multiple criteria and multiple outputs

* Note: We use the & operator to say if a value meets both conditions

* Note: We cannot use `dplyr::between` because our left side is not >=

```{r}

d8 %>%  
  dplyr::mutate(Var2_new = dplyr::if_else(Var2 >14, "high",
                                        dplyr::if_else((Var2 >6 & Var2 <=14), "medium", "low")))

```

* Note: Because `dplyr::if_else()` evaluates in order, we don't actually have to add the right side of our "medium" logic because all of the highs have been assigned and will not be written over by the mediums

```{r}

d8 %>%  
  dplyr::mutate(Var2_new = dplyr::if_else(Var2 >14, "high",
                                        dplyr::if_else((Var2 >6), "medium", "low")))

```


**4\. Collapse a numeric variable (Var2) into discrete numeric categories**

Review the data (d10)

```{r, echo=FALSE}

source("data.R")
d10

```

Create an indicator variable. 1 = 8 or lower, 0 = 9 or higher

*Note: NA will remain as NA because if_else does not evaluate it as TRUE or FALSE.

```{r}

d10 %>%  
  dplyr::mutate(Var2_new = dplyr::if_else(Var2 <= 8, 1, 0))

```

*Note: If you want to evaluate NA, you will need to specifically add it

```{r}

d10 %>%  
  dplyr::mutate(Var2_new = dplyr::if_else(Var2 <= 8 | is.na(Var2), 1, 0))

```

You could also specifically add NA this way as well

```{r}

d10 %>%  
  dplyr::mutate(Var2_new = dplyr::if_else(Var2 >8 & !is.na(Var2), 0, 1))

```

---

#### Function: `case_when()`

---

**Note:** `dplyr::case_when()` will evaluate in order, so proceed from most specific to most general. The function will not write over a previous evaluation even if something meets a new criteria. 

<br>

**1\. Collapse all "Google Meet" into one category**

Review the data (d6)

```{r, echo=FALSE}

source("data.R")
d6

```

Collapse all spellings of Google Meet into one uniform category

* Note: We are recoding into a new variable using `dplyr::mutate()` and naming the new variable a different name than the original. It keeps both the new and old versions of the variable. 

* Note: Using `stringr::str_detect()` we can detect different spellings of google meet

* Note: You can also combine the pattern "google|meet" like we did above in `dplyr::if_else()` rather than separate the patterns out.

```{r}

d6 %>%  
  dplyr::mutate(online_platform_new = dplyr::case_when(
  stringr::str_detect(online_platform,"google") ~ "Google Meet", 
  stringr::str_detect(online_platform,"meet") ~ "Google Meet", 
  TRUE ~ online_platform))

```

* Note: If you do not add *TRUE ~ online_platform*, you will get NA for all categories that are not specifically mentioned in the statement.

```{r}

d6 %>%  
  dplyr::mutate(online_platform_new = dplyr::case_when(
  stringr::str_detect(online_platform,"google") ~ "Google Meet", 
  stringr::str_detect(online_platform,"meet") ~ "Google Meet"))

```

* Note: And if you don't care about the other platforms, you can always collapse the remaining categories into another name, such as "other". Notice here that our NA value will be recoded into "Other". 

```{r}

d6 %>%  
  dplyr::mutate(online_platform_new = dplyr::case_when(
  stringr::str_detect(online_platform,"google") ~ "Google Meet", 
  stringr::str_detect(online_platform,"meet") ~ "Google Meet", 
  TRUE ~ "Other"))

```

* Note: If you want to keep NA as NA, you would need to specifically call out NA. For `dplyr::case_when()` you need to call out the specific type of NA. In this case *NA_character_*.

```{r}

d6 %>%  
  dplyr::mutate(online_platform_new = dplyr::case_when(
  is.na(online_platform) ~ NA_character_,
  stringr::str_detect(online_platform,"google") ~ "Google Meet", 
  stringr::str_detect(online_platform,"meet") ~ "Google Meet", 
  TRUE ~ "Other"))

```


**2\. Collapse all "Google Meet" into one category and clarify that "google" means Google Hangouts**

Review the data (d9)

```{r, echo=FALSE}

source("data.R")
d9

```

Collapse all spellings of Google Meet into one uniform category and clarify that google means Google Hangouts

* Note: Using `stringr::str_detect()` we can detect different spellings of google meet

* Note: Here I am adding the `base::tolower()` function to deal with the varying capitalization. You can also use the solutions we showed above in `dplyr::if_else()` where we add either (?i) to the pattern or add `stringr::regex()` around the pattern with the argument *ignore_case = TRUE*.

* Note: Note that I am not putting "google" at the top because if I did, all instances of "google" would be recoded to google hangouts.

```{r}

d9 %>%  
  dplyr::mutate(online_platform=tolower(online_platform)) %>%
  dplyr::mutate(online_platform_new = dplyr::case_when(
  stringr::str_detect(online_platform,"met") ~ "google meet", 
  stringr::str_detect(online_platform,"meet") ~ "google meet", 
  stringr::str_detect(online_platform, "google") ~ "google hangouts",
  stringr::str_detect(online_platform, "hangouts") ~ "google hangouts",
  TRUE ~ online_platform))

```

* Note: See how the result would be different if I put "google" at the top. The ordering matters. Once a statement is evaluated, it will not be rewritten.

```{r}

d9 %>%  
  dplyr::mutate(online_platform=tolower(online_platform)) %>%
  dplyr::mutate(online_platform_new = dplyr::case_when(
  stringr::str_detect(online_platform, "google") ~ "google hangouts",
  stringr::str_detect(online_platform,"met") ~ "google meet", 
  stringr::str_detect(online_platform,"meet") ~ "google meet",
  stringr::str_detect(online_platform, "hangouts") ~ "google hangouts",
  TRUE ~ online_platform))

```


---

#### Function: `recode()`

---


**1\. Collapse all "Google Meet" into one category**

Review the data (d6)

```{r, echo=FALSE}

source("data.R")
d6

```

Collapse all spellings of Google Meet into one uniform category


```{r}

d6 %>%  
  dplyr::mutate(online_platform_new =
                  dplyr::recode(
  online_platform, google = "Google Meet", meet = "Google Meet"))

```

Return to [Recode](https://github.com/Cghlewis/data-wrangling-functions/wiki/Recode)