---
title: "Recode values for multiple variables"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: dplyr

---

#### Function: `case_when()`

---

**1\. Recode across multiple character variables**

Review the data (d15)

```{r, echo=FALSE}

source("data.R")
d15

```

In this example, missing values were entered as character values in a numeric column. So now these columns are considered character.

Recode "MR" to -98 and "UI" to -99 for Var1, Var2, Var3. And then convert the character variable to numeric.

* Note: Unlike `dplyr::recode()` where we did not have to put quotes around the old character values, we now have to put quotes around the old and new character values.

* Note: I use `dplyr::across()` to select multiple variables to apply my function to.

* Note: The . is used here to say apply this function to all of the variables I selected above.

* Note: TRUE means, if a value was not evaluated in my arguments above, replace with the value I give. In this case, I am saying just replace with the existing value.


```{r}

d15 <- d15 %>%
  dplyr::mutate(dplyr::across(Var1:Var3,
                       ~ case_when(
                       . == "MR" ~ "-98",
                       . == "UI" ~ "-99",
                       TRUE ~ .)))

d15

```

While my character values have now be recoded to numeric values, my variable is still considered character so I would need one additional step to convert it to numeric (using the `base::as.numeric()` function). 

However, before I do that, I want to be certain I have no other character values left in my variables. If I do still have other character values and I try to use `base::as.numeric()` those values will be converted to NA.

I can check that there are no more instances of MR or UI using `dplyr::filter()` in conjunction with the `dplyr::if_any()` predicate function to select columns where the condition applies to any of the columns. This function is available in version 1.0.5 of `dplyr`.

* Note: We use the `tidyselect` selection helper `everything()` to refer to all variables. I could also have selected variables Var1:Var3.

* .x is used to say, for all of the variables I have selected

```{r}

d15 %>% 
dplyr::filter(dplyr::if_any(everything(), ~ .x %in% c('MR', 'UI')))

```

If I wanted to be more general, I could also just check for *any* alphabetical character left in my variables using `stringr::str_detect()` and including the regex pattern *[a-zA-Z]*

```{r}

d15 %>% 
dplyr::filter(dplyr::if_any(everything(), ~ stringr::str_detect(., pattern = "[a-zA-Z]")))

```

Now that I know my variable has all character values removed, I can convert them to numeric.

* Note: I am using `dplyr::across()` to apply a function to multiple variables. 

```{r}

d15 %>%
  dplyr::mutate(dplyr::across(Var1:Var3, as.numeric))

```

**2\. Recode across multiple numeric variables**

Review the data (d15)

```{r, echo=FALSE}

source("data.R")
d16

```

In this situation, -80 means the item was not provided to a student. If a student has a -80 for all items in a scale, we want to assign a -80 for the sum and mean of that scale to denote that this scale was not provided (to differentiate between a value of NA meaning the data are actually missing).

The two scales are "sc" and "tr" and each are made up of 2 items.

* Note: I am using `dplyr::across()` to apply a function to multiple variables. 

* Note: I am using the `dplyr::if_all()` predicate function to select columns where the condition applies to all columns. This function is available in version 1.0.5 of `dplyr`.


```{r}

d16 %>%
  dplyr::mutate(dplyr::across(sc_sum:sc_mean,
                 ~dplyr::case_when(
                    dplyr::if_all(sc_1:sc_2, ~ (. == -80)) ~ -80,
                    TRUE ~ .)),
  dplyr::mutate(dplyr::across(tr_sum:tr_mean,
                  ~dplyr::case_when(
                    dplyr::if_all(tr_1:tr_2, ~ (. == -80)) ~ -80,
                    TRUE ~ .))))
    
```


Return to [Recode](https://github.com/Cghlewis/data-wrangling-functions/wiki/Recode)