---
title: "Recode NA"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: tidyr

---

#### Function: `replace_na()`

---

**1\. Replace NA values for one variable (Var3)**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Recode NA to 0

* Note: We are recoding back into the same variable using `dplyr::mutate()`. However, we could have recoded into a new variable by changing the name of *Var3*.

```{r}

d1 %>% 
  dplyr::mutate(Var3 = tidyr::replace_na(Var3, replace = 0))

```


**2\. Replace NA for multiple variables (Var2 and Var3) with the same NA value**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Recode NA for Var2 and Var3 to 0

* Note: We are recoding back into the same variable using `dplyr::mutate()`. 
* Note: We are using `dplyr::across()` to apply a transformation to multiple columns.

```{r}

d1 %>% 
  dplyr::mutate(dplyr::across(c(Var2:Var3), tidyr::replace_na, 0))

```


**3\. Replace NA for multiple variables of the same class with the same NA value (Var2 and Var3)**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Recode NA for Var2 and Var3 to -999

* Note: We are recoding back into the same variable using `dplyr::mutate()`. 
* Note: We are using `dplyr::across()` to apply a transformation to multiple columns.
* Note: You must wrap *is.numeric*, a predicate function (returns a true/false), in the `tidyselect` selection helper `where()`.

```{r}

d1 %>% 
  dplyr::mutate(dplyr::across(where(is.numeric), tidyr::replace_na, -999))

```


**4\. Replace NA Var2 and Var3 with different values**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Replace NA with -999 for Var2 and replace NA with 0 for Var3.

Here we can use `base::list()` to list the variables we want to transform. Because we are not replacing NAs in a vector, we do not have to use `dplyr::mutate()` to do our transformation.

```{r}

d1 %>% 
  tidyr::replace_na(list(Var2 = -999, Var3 = 0))

```

### Package: dplyr

---

#### Function: `na_if()`

---

**1\. Replace a value for the entire dataset with NA**

Review the data (d3)

```{r, echo=FALSE}

source("data.R")
d3

```

Replace -999 for the entire dataset with NA

```{r}

d3 %>% 
  dplyr::na_if(-999)

```


**2\. Replace a value for select variables (Var2 and Var3) with NA**

Review the data (d3)

```{r, echo=FALSE}

source("data.R")
d3

```

Replace -999 for for Var2 and Var3 with NA

* Note: We are recoding back into the same variable using `dplyr::mutate()`. 
* Note: We are using `dplyr::across()` to apply a transformation to multiple columns.

```{r}

d3 %>% 
  dplyr::mutate(dplyr::across(c(Var2:Var3), dplyr::na_if, -999))

```

**3\. Replace a value for multiple variables of the same class (in this case numeric variables) with NA**

Review the data (d3)

```{r, echo=FALSE}

source("data.R")
d3

```

Replace -999 for numeric variables with NA

* Note: We are recoding back into the same variable using `dplyr::mutate()`.
* Note: We are using `dplyr::across()` to apply a transformation to multiple columns.
* Note: You must wrap *is.numeric*, a predicate function (returns a true/false), in the `tidyselect` selection helper `where()`.


```{r}

d3 %>% 
  dplyr::mutate(dplyr::across(where(is.numeric), dplyr::na_if, -999))

```

Return to [Recode](https://github.com/Cghlewis/data-wrangling-functions/wiki/Recode)