---
title: "Recode NA"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: tidyr

---

#### Function: `replace_na()`

---

**1\. Replace NA values for one variable (`Var2`)**

Review the data (d18)

```{r, echo=FALSE}

source("data.R")
d18

```

Recode NA to 0


```{r}

d18 %>% 
  dplyr::mutate(Var2 = tidyr::replace_na(Var2, replace = 0))

```


**2\. Replace NA for multiple variables (`Var1` and `Var2`) with the same NA value**

Review the data (d18)

```{r, echo=FALSE}

source("data.R")
d18

```

Recode NA for `Var1` and `Var2` to 0

* Note: We are using `dplyr::across()` to apply a transformation to multiple columns.

```{r}

d18 %>%
  dplyr::mutate(dplyr::across(Var1:Var2, ~ tidyr::replace_na(., 0)))

```

**3\. Replace NA for multiple variables (`Var2` and `Var3`) with the same NA value using your project data dictionary**

Review the data (d18)

```{r, echo=FALSE}

source("data.R")
d18

```

It's possible that you may have built a data dictionary before collecting data and you may have added a column to your dictionary that helps you identify which columns need to be recoded in a particular way.

Let's review our very simple data dictionary.

```{r}

dict

```

Here we see our data dictionary has a column called "var_type" which tells us which variables are binary (or coded 0 and 1).

Oftentimes when you download "select all" questions from a survey platform, each option is downloaded with a 1 (if selected) and NA (if not selected). There are times when this format makes sense (say for visualizations). However, for analysis purposes we probably want those variables to show 1s (Yes) and 0s (No) instead of 1s and NAs.

Instead of listing out all of the variables we want to recode, we can pull a vector of variable names using our dictionary.

```{r}

vars <- dict %>%
  dplyr::filter(var_type == "binary") %>%
  dplyr::select(var_name) %>%
  dplyr::pull()

```

We can now use this vector in our `tidyr::replace_na()` function.

* Note: We use `tidyselect::all_of()` to denote that we are selecting variable names from an external vector ("vars").

```{r}

d18 %>%
  dplyr::mutate(dplyr::across(tidyselect::all_of(vars), ~tidyr::replace_na(., 0)))

```


**4\. Replace NA for multiple variables of the same class with the same NA value (`Var2` and `Var3`)**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Recode NA for `Var2` and `Var3` to -999

* Note: You must wrap *is.numeric*, a predicate function (returns a true/false), in the `tidyselect` selection helper `where()`.

```{r}

d1 %>% 
  dplyr::mutate(dplyr::across(where(is.numeric), ~tidyr::replace_na(., -999)))

```


**5\. Replace NA `Var2` and `Var3` with different values**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Replace NA with -999 for `Var2` and replace NA with 0 for `Var3`.

Here we can use `base::list()` to list the variables we want to transform. Because we are not replacing NAs in a vector, we do not have to use `dplyr::mutate()` to do our transformation.

```{r}

d1 %>% 
  tidyr::replace_na(list(Var2 = -999, Var3 = 0))

```

### Package: dplyr

---

#### Function: `na_if()`

---

**1\. Replace a value for the entire dataset with NA**

Review the data (d3)

```{r, echo=FALSE}

source("data.R")
d3

```

Replace -999 for the entire dataset with NA

```{r}

d3 %>% 
  dplyr::na_if(-999)

```


**2\. Replace a value for select variables (`Var2` and `Var3`) with NA**

Review the data (d3)

```{r, echo=FALSE}

source("data.R")
d3

```

Replace -999 for for `Var2` and `Var3` with NA

```{r}

d3 %>% 
  dplyr::mutate(dplyr::across(Var2:Var3, ~dplyr::na_if(., -999)))

```

**3\. Replace a value for multiple variables of the same class (in this case numeric variables) with NA**

Review the data (d3)

```{r, echo=FALSE}

source("data.R")
d3

```

Replace -999 for numeric variables with NA


```{r}

d3 %>%
  dplyr::mutate(dplyr::across(where(is.numeric), ~ dplyr::na_if(.,-999)))

```

Return to [Recode](https://github.com/Cghlewis/data-wrangling-functions/wiki/Recode)