---
title: "Recode values"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: dplyr

---

#### Function: `recode()`

---

**Note:** The `dplyr::recode()` formula is: old value=new value, this is opposite of `dplyr::rename()`

<br>

**1\. Recode character values into different character values (for gender)**

Review the data (d4)

```{r, echo=FALSE}

source("data.R")
d4

```

Recode gender

* Note: We did not have to put quotes around the old values because `dplyr::recode()` replaces character or factor values by their name.

* Note: Parentheses *are* required around the new value when recoding into a character/factor variable.

* Note: We are recoding back into the same variable using `dplyr::mutate()`. However, we could have recoded into a new variable by changing the name of *gender*.

```{r}

d4 %>% dplyr::mutate(gender=dplyr::recode(gender, m="male", f="female", n="nonbinary"))

```


**2\. Recode a character variable (gender) into a numeric variable**

Review the data (d4)

```{r, echo=FALSE}

source("data.R")
d4

```

Recode gender

* Note: We did not have to put quotes around the old values because `dplyr::recode()` replaces character or factor values by their name.

* Note: Since we are recoding into a numeric variable, no quotes are necessary for the numeric new value.

* Note: We are recoding back into the same variable using `dplyr::mutate()`. However, we could have recoded into a new variable by changing the name of *gender*.

```{r}

d4 %>% dplyr::mutate(gender=dplyr::recode(gender, m=1, f=2, n=3))

```


**3\. Recode just one value in a variable (Var2)**

Review the data (d5)

```{r, echo=FALSE}

source("data.R")
d5

```

Recode `16` to `4` in Var2

Note: We *did* have to put quotes around the old values because `dplyr::recode()` replaces numeric values by either their name or their position. If a number with no quotes/backticks is given, it will assume it is a position.

Note: Since we are recoding into a numeric variable, no quotes are necessary are the numeric new value.

Note: Notice here that we are only recoding one value in the variable. There is a *default* option that you can set all not recoded values to, and if no default is supplied and the replacement values are the same variable class as the original values (ex: numeric and numeric), then unmatched values are unchanged. *However*, if the replacement value is a new variable class (ex: numeric to character) then all other values will be recoded to NA.


```{r}

d5 %>% dplyr::mutate(Var2=dplyr::recode(Var2, `16`=4))

```

**4\. Recode one value in a numeric haven labelled variable (Var3)**

Review the data (d5)

```{r, echo=FALSE}

source("data.R")
d5

```

View the variable labels for Var3

* Note: You'll see in this example there are three values and two value labels. In this instance, 2s are supposed to also be yes, but they do not contain a value label.

```{r, echo=FALSE}

d5 <- d5 %>% labelled::set_value_labels(Var3 = c(no = 0, yes = 1))

```


```{r}

labelled::val_labels(d5$Var3)

```


Recode `2` to `1` in Var3

Note: We *did* have to put quotes around the old values because `dplyr::recode()` replaces numeric values by either their name or their position. If a number with no quotes/backticks is given, it will assume it is a position.

Note: Since we are recoding into a numeric variable, no quotes are necessary are the numeric new value.

Note: Notice here that we are only recoding one value in the variable. There is a *.default* option that you can set all not recoded values to, and if no default is supplied and the replacement values are the same variable class as the original values (ex: numeric and numeric), then unmatched values are unchanged. *However*, if the replacement value is a new variable class (ex: numeric to character) then all other values will be recoded to NA.

```{r}

d5 <- d5 %>% dplyr::mutate(Var3 = dplyr::recode(Var3, `2`=1))

d5

```

View the variable labels for Var3

Note: The value labels are retained. The default for the `dplyr::recode()` option *.keep_value_labels* is TRUE. However, if you do not want to keep the value labels you can set the option to FALSE.


```{r}

labelled::val_labels(d5$Var3)

```

**4\. Recode multiple values in a numeric haven labelled variable (Var3)**

Review the data (d13)

```{r, echo=FALSE}

source("data.R")
d13

```

View the variable labels for Var3

* Note: You'll see in this example there are three values and two value labels. In this instance, 2s are supposed to also be yes, but they do not contain a value label.

```{r, echo=FALSE}

d13 <- d13 %>% labelled::set_value_labels(Var3 = c(no = 5, yes = 6))

```

```{r}

d13 %>%
  dplyr::select(Var3) %>%
  labelled::val_labels()

```

Recode `5` to `0` and `6` to `1` in Var3

Note: We *did* have to put quotes around the old values because `dplyr::recode()` replaces numeric values by either their name or their position. If a number with no quotes/backticks is given, it will assume it is a position.

Note: Since we are recoding into a numeric variable, no quotes are necessary are the numeric new value.

```{r}

d13 <- d13 %>% dplyr::mutate(Var3 = dplyr::recode(Var3, `5`=0, `6` = 1))

d13

```

View the variable labels for Var3

Note: We left the default *.keep_value_labels* is TRUE so our value labels are still funky.

```{r}

d13 %>%
  dplyr::select(Var3) %>%
  labelled::val_labels()

```


In this instance I don't want to keep the original value labels. I don't necessarily need to set the *.keep_value_labels* to FALSE though. I can just add new value labels and they will overwrite the previous labels.

```{r}

d13 <- d13 %>% labelled::set_value_labels(Var3 = c(no = 0, yes = 1))

d13 %>%
  dplyr::select(Var3) %>%
  labelled::val_labels()

```

---

#### Function: `recode_factor()`

---

**Note:** The `dplyr::recode_factor()` formula is: old value=new value, this is opposite of `dplyr::rename()`

**Note:** The function `dplyr::recode_factor` will convert your variable to factor **and** recode it.

<br>

**1\. Convert a character variable (lunch) into a factor and recode it**

Review the data (d4)

```{r, echo=FALSE}

source("data.R")
d4

```

Recode lunch status

* Note: We did not have to put quotes around the old values because `dplyr::recode_factor()` replaces character or factor values by their name.

* Note: Parentheses *are* required around the new value when recoding into a character/factor variable.

* Note: We are recoding back into the same variable using `dplyr::mutate()`. However, we could have recoded into a new variable by changing the name of *lunch*.

```{r}

d4 <- d4 %>% dplyr::mutate(lunch=dplyr::recode_factor(lunch, p="paid", r="reduced", f="free"))

d4

```

Review the factor levels

```{r}

levels(d4$lunch)

```

**2\. Recode a numeric variable (lunch) into a factor**

Review the data (d7)

```{r, echo=FALSE}

source("data.R")
d7

```

Recode lunch status

Note: We *did* have to put quotes around the old values because `dplyr::recode_factor()` replaces numeric values by either their name or their position. If a number with no quotes/backticks is given, it will assume it is a position.

* Note: Parentheses are required around new value when recoding into a character/factor variable.

* Note: We are recoding back into the same variable using `dplyr::mutate()`. However, we could have recoded into a new variable by changing the name of *lunch*.

```{r}

d7 <- d7 %>% dplyr::mutate(lunch=dplyr::recode_factor(
  lunch, "1"="paid", "2"="reduced", "3"="free"))

d7

```

Review the factor levels

```{r}

levels(d7$lunch)

```


### Package: forcats

---

#### Function: `fct_recode()`

---

**Note:** The `forcats::fct_recode()` formula is: new value=old value, this is opposite of `dplyr::recode()` but the same as `dplyr::rename()`

**Note:** The function `forcats::fct_recode` will convert your variable to factor **and** recode it.

<br>

**1\. Recode a numeric variable (lunch) into a factor**

Review the data (d7)

```{r, echo=FALSE}

source("data.R")
d7

```

Recode lunch status

* Note: `forcats::fct_recode` will only recode factor or character variables. If your variable is numeric, you will need to convert it to character or factor before using recode factor. As you see below, I converted our numeric variable to character in order to use the function.

* Note: Completely counterintuitive to what we've seen with `dplyr::recode()`, 1) you *do* have to put quotes around your character old value, 2) if you new values are character, you do *not* have to put quotes around them, 3) however, if your new values are numeric, you *do* need quotes around them.

* Note: We are recoding back into the same variable using `dplyr::mutate()`. However, we could have recoded into a new variable by changing the name of *lunch*.

```{r}

d7 <- d7 %>% dplyr::mutate(lunch=forcats::fct_recode(as.character(lunch), paid="1", reduced="2", free="3"))

d7

```

Review the factor levels

```{r}

levels(d7$lunch)

```

Return to [Recode](https://github.com/Cghlewis/data-wrangling-functions/wiki/Recode)