---
title: "Recode dates"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: dplyr

---

#### Function: `case_when()`

---

**1\. Recode all missing survey dates to the midpoint of the date range for id 3 and 6**

Review the data (d17)

```{r, echo=FALSE}

source("data.R")
d17 <- d17 %>% 
  mutate(svy_date = as.Date(svy_date))

d17

```



But first we need to find the midpoint of the `svy_date` variable. We can do that by getting the min and max (using `base::range()`) and then taking the mean using `base::mean()` of that range. We need to make sure to add the argument *na_rm = TRUE* to calculate the range despite having NA values. We can save the value in a vector to be used later.

* Note: Notice that the class of our "svy_date" variable is date (we can see above). If it was character we would need to convert it to date before performing calculations.

```{r}

date_mid <- base::mean(base::range(d17$svy_date, na.rm = TRUE))

date_mid

```
Now we can use this vector in our recoding statement. We can use `dplyr::case_when()` to recode the NA values. 

* Note: I am using `dplyr::mutate()` to create a new `svy_date` variable which will replace the existing variable.

* Note: TRUE means, if a value was not evaluated in my arguments above, replace with the value I give. In this case, I am saying just replace with the existing value.

```{r}

d17_new <- d17 %>%
  dplyr::mutate(svy_date = dplyr::case_when(
    id %in% c(3,6) ~ date_mid,
    TRUE ~ svy_date
  ))

d17_new

```

Awesome. This looks like it worked great. But wait, let's do one last check. Let's make a table of the values.

```{r}

d17_new %>%
  janitor::tabyl(svy_date)

```

Well that doesn't look right at all. For some reason this `janitor::tabyl()` function is not recognizing that my new "2022-10-07" values should be the same as the existing "2022-10-07" values. So I decide to try out a few date formats in my recode function until I find one that forces R to recognize them as the same value. I land on `lubridate::ymd()`.

```{r}

d17_new <- d17 %>%
  dplyr::mutate(svy_date = dplyr::case_when(
    id %in% c(3,6) ~ lubridate::ymd(date_mid),
    TRUE ~ svy_date
  ))

d17_new

```

```{r}

d17_new %>%
  janitor::tabyl(svy_date)

```


Return to [Recode](https://github.com/Cghlewis/data-wrangling-functions/wiki/Recode)