---
title: "Recode values using a wide formatted data dictionary"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: dplyr

---

#### Function: `case_when()`

---

<br>


**1\. Recode values using a wide formatted data dictionary**

Review the data (d24)

```{r, echo=FALSE}

source("data.R")
d24

```

Review our current wide formatted data dictionary (dict_wide)

```{r, echo = FALSE}

dict_wide

```

Select all variables with the same recoding logic (-99 -> 0)

```{r}

recode_multi_select <- dict_wide %>%
  dplyr::filter(transformation == "-99 -> 0") %>%
  dplyr::pull(var_name)

recode_multi_select

```

Now we can use that vector of variable names to recode these variables ll at once.

```{r}

d24 %>%
  dplyr::mutate(dplyr::across(
    dplyr::all_of(recode_multi_select),
                ~dplyr::case_when(
                  . == -99 ~ 0,
                  TRUE ~ .
                )))

```


Go to [Recode using long formatted data dictionary](https://cghlewis.github.io/data-wrangling-functions/recode-values/recode-values-dictionary.html)

Return to [Recode](https://github.com/Cghlewis/data-wrangling-functions/wiki/Recode)