---
title: "Recode values using a long formatted data dictionary"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: dplyr

---

#### Function: `recode()`

---

**Note:** The `dplyr::recode()` formula is: old value=new value, this is opposite of `dplyr::rename()`.

<br>


**1\. Recode values using a long formatted data dictionary**

Review the data (d23)

```{r, echo=FALSE}

source("data.R")
d23

```

Review our current long formatted data dictionary (dict_long) that contains the old values of our variable `org_name` (old_values) and the new values that we want instead (new_values).

Note that this dataset is not in the usual data dictionary format I would think of (with each row corresponding to one variable). In this case, this would be a specialized long format type of codebook that you may have received or created solely for the purpose of recoding values.

```{r, echo = FALSE}

dict_long

```

In order to use this information to recode our values, we first need to pull a named vector that includes any value that is recoded.

```{r}

dict_long <- dict_long %>%
  dplyr::filter(!is.na(new_value)) %>%
  tibble::deframe()

```

We can now use the named vector to recode our values. We can use `!!!` to splice a our vector into the argument of a quoting expression.

```{r}

d23 %>% 
  dplyr::mutate(org_name = dplyr::recode(org_name, !!!dict_long))

```

Go to [Recode using wide formatted data dictionary](https://cghlewis.github.io/data-wrangling-functions/recode-values/recode-values-dictionary-wide.html)

Return to [Recode](https://github.com/Cghlewis/data-wrangling-functions/wiki/Recode)