---
title: "Recode factors"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: dplyr

---

#### Function: `recode_factor()`

---

**Note:** The `dplyr::recode_factor()` formula is: old value=new value, this is opposite of `dplyr::rename()`

**Note:** The function `dplyr::recode_factor()` will convert your variable to factor **and** recode it.

<br>

**1\. Convert a character variable (lunch) into a factor and recode it**

Review the data (d4)

```{r, echo=FALSE}

source("data.R")
d4

```

Recode lunch status

* Note: We did not have to put quotes around the old values because `dplyr::recode_factor()` replaces character or factor values by their name.

* Note: Quotes *are* required around the new value when recoding into a character/factor variable.

* Note: We are recoding back into the same variable using `dplyr::mutate()`. However, we could have recoded into a new variable by changing the name of *lunch*.

```{r}

d4 <- d4 %>%
  dplyr::mutate(lunch = dplyr::recode_factor(lunch, p = "paid", r = "reduced", f = "free"))

d4

```

Review the factor levels

```{r}

levels(d4$lunch)

```

**2\. Recode a numeric variable (lunch) into a factor**

Review the data (d7)

```{r, echo=FALSE}

source("data.R")
d7

```

Recode lunch status

Note: We *did* have to put back ticks around the old values because `dplyr::recode_factor()` replaces numeric values by either their name or their position. If a number with no quotes/backticks is given, it will assume it is a position.

```{r}

d7 <- d7 %>%
  dplyr::mutate(lunch = dplyr::recode_factor(
    lunch,
    "1" = "paid",
    "2" = "reduced",
    "3" = "free"
  ))

d7

```

Review the factor levels

```{r}

levels(d7$lunch)

```

<br>


### Package: forcats

---

#### Function: `fct_recode()`

---

**Note:** The `forcats::fct_recode()` formula is: new value = old value, this is opposite of `dplyr::recode()` but the same as `dplyr::rename()`

**Note:** The function `forcats::fct_recode` will convert your variable to factor **and** recode it.

<br>

**1\. Recode a numeric variable (`lunch`) into a factor**

Review the data (d7)

```{r, echo=FALSE}

source("data.R")
d7

```

Recode lunch status

* Note: `forcats::fct_recode` will only recode factor or character variables. If your variable is numeric, you will need to convert it to character or factor before using recode factor. As you see below, I converted our numeric variable to character in order to use the function.

* Note: Counterintuitive to `dplyr::recode()`, 1) you *do* have to put quotes around your character old value, 2) if you new values are character, you do *not* have to put quotes around them, 3) however, if your new values are numeric, you *do* need quotes around them.

```{r}

d7 <- d7 %>%
  dplyr::mutate(lunch = forcats::fct_recode(
    as.character(lunch),
    paid = "1",
    reduced = "2",
    free = "3"
  ))

d7

```

Review the factor levels

```{r}

base::levels(d7$lunch)

```


Return to [Recode](https://github.com/Cghlewis/data-wrangling-functions/wiki/Recode)