---
title: "Recode values using a named vector"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: dplyr

---

#### Function: `recode()`

---

**Note:** The `dplyr::recode()` formula is: old value=new value, this is opposite of `dplyr::rename()`.

<br>


**1\. Recode values using a long formatted data dictionary**

Review the data 23

```{r, echo=FALSE}

source("data.R")
d23

```

Review our current long formatted data dictionary that contains the old values of our variable `org_name` (old_values) and the new values that we want instead (new_values)

```{r, echo = FALSE}


dict_org <- tribble(~old_value, ~new_value,
                    "Org Name A", "New Org Name A",
                    "Org Name B", "New Org Name B",
                    "Org Name C", "New Org Name C",
                    "Org Name D", NA_character_,
                    "Org Name E", "Org Name D",
                    "Org Name F", NA_character_)

dict_org

```

In order to use this information to recode our values, we first need to pull a named vector.

```{r}

dict_org <- dict_org %>%
  dplyr::filter(!is.na(new_value)) %>%
  tibble::deframe()

```

We can now use the named vector to recode our values. We can use `!!!` !!! to splice a our vector into the argument of a quoting expression.

```{r}

d23 %>% 
  dplyr::mutate(org_name = dplyr::recode(org_name, !!!dict_org))

```


Return to [Recode](https://github.com/Cghlewis/data-wrangling-functions/wiki/Recode)