---
title: "Recode NA"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: tidyr

---

#### Function: `tidyr::replace_na()`

---

**1\. Replace NA for one variable**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Check the values

```{r}

d1 %>% janitor::tabyl(Var3, show_na=TRUE)

```

Recode NA

```{r}

d1 <- d1 %>% dplyr::mutate(Var3 = tidyr::replace_na(Var3, 0))

```

Check that the recode worked

```{r}

d1 %>% janitor::tabyl(Var3, show_na=TRUE)

```

**2\. Replace NA for multiple variables with the same NA value**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Check the values

```{r}

lapply(d1, table, useNA="always")

```

Recode NA for Var2 and Var3

```{r}

d1 <- d1 %>% dplyr::mutate_at(vars(Var2:Var3), tidyr::replace_na , 0)

```

Check the values after the recode

```{r}

lapply(d1, table, useNA="always") 

```

**3\. Replace NA for multiple variables of the same class with the same NA value**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Check the values

```{r}

lapply(d1, table, useNA="always") 

```

Recode NA for Var2 and Var3

```{r}

d1 <- d1 %>% dplyr::mutate_if(is.numeric , tidyr::replace_na, replace = -999)

```

Check the values after the recode

```{r}

lapply(d1, table, useNA="always") 

```


### Package: dplyr

---

#### Function: `dplyr::na_if()`

---

**1\. Replace a value for the entire dataset with NA**

Review the data (d3)

```{r, echo=FALSE}

source("data.R")
d3

```

Replace -999 for the entire dataset with NA

```{r}

d3 %>% dplyr::na_if(-999)

```


**2\. Replace a value for select variables with NA**

```{r, echo=FALSE}

source("data.R")
d3

```

Replace -999 for for Var2 and Var3 with NA

```{r}

d3 %>% dplyr::mutate_at(vars(Var2:Var3), dplyr::na_if, -999) 

```

**3\. Replace a value for multiple variables of the same class with NA**

```{r, echo=FALSE}

source("data.R")
d3

```

Replace -999 for numeric variables with NA

```{r}

d3 %>% dplyr::mutate(across(where(is.numeric), dplyr::na_if, -999))

```

You can also apply `dplyr::na_if` with `dplyr::mutate_if` or `dplyr::mutate_at`

```{r, echo=FALSE}

source("data.R")

d3

```

```{r}

d3 %>% mutate_if(is.numeric, na_if, -999)

```


