---
title: "Recode values for multiple variables"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: dplyr

---

#### Function: `case_when()`

---

**Note:** `dplyr::case_when()` will evaluate in order, so proceed from most specific to most general. The function will not write over a previous evaluation even if something meets a new criteria. 

<br>

**1\. Recode across multiple character variables (all variables)**

Review the data (d19)

```{r, echo=FALSE}

source("data.R")
d19

```

When this data was entered it appears that both lower and upper case letters were used. 

We want to recode all "y" to "Y" and all "n" to "N" for Var1, Var2 and Var3.

* Note: Unlike `dplyr::recode()` where we did not have to put quotes around the old character values, we now have to put quotes around the old and new character values.

* Note: I use `dplyr::across()` to select multiple variables to apply my function to.

* Note: We use the `tidyselect` selection helper `everything()` to refer to all variables. I could also have selected variables Var1:Var3.

* Note: The . is used here to say apply this function to all of the variables I selected above.

* Note: TRUE means, if a value was not evaluated in my arguments above, replace with the value I give. In this case, I am saying just replace with the existing value.


```{r}

d19 %>%
  dplyr::mutate(dplyr::across(everything(),
                                 ~case_when(
                                   . == "y" ~ "Y",
                                   . == "n" ~ "N",
                                   TRUE ~ .
                                 )))

```

A more sophisticated way to do this might be to bypass the `dplyr::case_when()` statement all together though and use something like `stringr::str_to_upper()` though.

```{r}

d19 %>%
  dplyr::mutate(dplyr::across(everything(), ~stringr::str_to_upper(.)))

```

**2\. Recode across multiple character variables (Var1:Var3**

Review the data (d15)

```{r, echo=FALSE}

source("data.R")
d15

```

In this example, missing values were entered as character values in a numeric column. So now these columns are considered character.

Recode "MR" to -98 and "UI" to -99 for Var1, Var2, Var3. 

Notice that I am putting -98 and -99 in quotes. That is because this variable is currently considered a character variable. So we will recode first, and then convert the variable to numeric.


```{r}

d15 <- d15 %>%
  dplyr::mutate(dplyr::across(Var1:Var3,
                       ~ case_when(
                       . == "MR" ~ "-98",
                       . == "UI" ~ "-99",
                       TRUE ~ .)))

d15

```

Remember I still need to do one additional step to convert this variable to numeric (using the `base::as.numeric()` function). 

However, before I do that, I want to be certain I have no other character values left in my variables. If I do still have other character values and I try to use `base::as.numeric()` those values will be converted to NA.

I can check that there are no more instances of MR or UI using `dplyr::filter()` in conjunction with the `dplyr::if_any()` predicate function to select columns where the condition applies to any of the columns. This function is available in version 1.0.5 of `dplyr`.

* .x (or just .) is used to say, for all of the variables I have selected

```{r}

d15 %>% 
  dplyr::filter(dplyr::if_any(everything(), ~ .x %in% c('MR', 'UI')))

```

If I wanted to be more general, I could also just check for *any* alphabetical character left in my variables using `stringr::str_detect()` and including the regex pattern *[a-zA-Z]*

```{r}

d15 %>% 
  dplyr::filter(dplyr::if_any(everything(), ~ stringr::str_detect(., pattern = "[a-zA-Z]")))

```

Now that I know my variable has all character values removed, I can convert them to numeric.

```{r}

d15 %>%
  dplyr::mutate(dplyr::across(Var1:Var3, as.numeric))

```

**3\. Recode across multiple numeric variables (sum and means variables)**

Review the data (d15)

```{r, echo=FALSE}

source("data.R")
d16

```

In this situation, -80 means the item was not provided to a student. If a student has a -80 for all items in a scale, we want to assign a -80 for the sum and mean of that scale to denote that this scale was not provided (to differentiate between a value of NA meaning the data are actually missing).

The two scales are "sc" and "tr" and each are made up of 2 items.

* Note: I am using the `dplyr::if_all()` predicate function to select columns where the condition applies to all columns. This function is available in version 1.0.5 of `dplyr`.


```{r}

d16 %>%
  dplyr::mutate(dplyr::across(sc_sum:sc_mean,
                 ~dplyr::case_when(
                    dplyr::if_all(sc_1:sc_2, ~ . == -80) ~ -80,
                    TRUE ~ .)),
  dplyr::mutate(dplyr::across(tr_sum:tr_mean,
                  ~dplyr::case_when(
                    dplyr::if_all(tr_1:tr_2, ~ . == -80) ~ -80,
                    TRUE ~ .))))
    
```


Return to [Recode](https://github.com/Cghlewis/data-wrangling-functions/wiki/Recode)