---
title: "Recode using a packaged dictionary"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: gendercoder

---

#### Function: `recode_gender()`

---

1. There may be times when you would rather recode open-ended gender responses based on an existing dictionary. In this case, you can use an existing package that has a dictionary built in. In this example, we will recode gender using the `gendercoder` package.

To see the full dictionary used for recodes, see [GenderDictionary_List.csv](https://github.com/ropenscilabs/gendercoder/blob/master/inst/extdata/GenderDictionary_List.csv)

Review the data (d11)

```{r, echo=FALSE}

source("data.R")
d11

```

Recode open-response gender into standardized gender categories based on a built-in dictionary. 

* Note: In this package you can choose between a broader recoding of gender, using the argument *dictionary = manylevels_en* (the default) or a narrower recoding using the argument *dictionary = fewlevels_en*.

* Note: You can also choose to retain any values that are not found using the argument *retain_unmatched = TRUE*.

* Note: We are recoding into a new variable using `dplyr::mutate()` and naming the new variable a different name than the original. It keeps both the new and old versions of the variable. 

```{r}

d11 %>%  
  dplyr::mutate(gender_new = gendercoder::recode_gender(gender, retain_unmatched = TRUE))

```


```{r, echo = FALSE, eval = FALSE, include = FALSE}

# Not working
# If we wanted to use the fewer levels dictionary, we could # change the *dictionary* argument and get this output # instead.

d11 %>%  
  dplyr::mutate(gender_new =
                  gendercoder::recode_gender(gender,
                                             dictionary =
                                               fewlevels_en,
                                            retain_unmatched
                                            = TRUE))

```

Return [Recode](https://github.com/Cghlewis/data-wrangling-functions/wiki/Recode)
