---
title: "Recode values"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: dplyr

---

#### Function: `recode()`

---

**Note:** The `dplyr::recode()` formula is: old value=new value, this is opposite of `dplyr::rename()`

<br>

**1\. Recode character values into different character values (for gender)**

Review the data (d4)

```{r, echo=FALSE}

source("data.R")
d4

```

Recode gender

* Note: Parentheses are required around the new value when recoding into a character variable.

* Note: We are recoding back into the same variable using `dplyr::mutate()`. However, we could have recoded into a new variable by changing the name of *gender*.

```{r}

d4 %>% dplyr::mutate(gender=dplyr::recode(gender, m="male", f="female", n="nonbinary"))

```

**2\. Recode a character variable (lunch) into a factor**

Review the data (d4)

```{r, echo=FALSE}

source("data.R")
d4

```

Recode lunch status

* Note: Parentheses are required around new value when recoding into a character/factor variable.

* Note: We are recoding back into the same variable using `dplyr::mutate()`. However, we could have recoded into a new variable by changing the name of *lunch*.

```{r}

d4 <- d4 %>% dplyr::mutate(lunch=dplyr::recode_factor(lunch, p="paid", r="reduced", f="free"))

d4

```

Review the factor levels

```{r}

levels(d4$lunch)

```

**3\. Recode a character variable (gender) into a numeric variable**

Review the data (d4)

```{r, echo=FALSE}

source("data.R")
d4

```

Recode gender

* Note: We are recoding back into the same variable using `dplyr::mutate()`. However, we could have recoded into a new variable by changing the name of *gender*.

```{r}

d4 %>% dplyr::mutate(gender=dplyr::recode(gender, m=1, f=2, n=3))

```


**4\. Recode just one value in a variable (Var2)**

Review the data (d5)

```{r, echo=FALSE}

source("data.R")
d5

```

Recode `16` to `4` in Var2

* Note:  The `dplyr::recode()` function considers the old value as a *name* and therefore, in order to use a number as a name or a character value, you have to surround it with backticks

```{r}

d5 %>% dplyr::mutate(Var2=dplyr::recode(Var2, `16`=4))

```

