---
title: "Recode values"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: dplyr

---

#### Function: `recode()`

---

**Note:** The `dplyr::recode()` formula is: old value=new value, this is opposite of `dplyr::rename()`.

<br>


**1\. Recode character values into different character values (for `gender`)**

Review the data (d4)

```{r, echo=FALSE}

source("data.R")
d4

```

Recode `gender`

* Note: We did not have to put quotes around the old character values because `dplyr::recode()` replaces character or factor values by their name.

* Note: Parentheses *are* required around the new value when recoding into a character/factor variable.

* Note: We are recoding back into the same variable using `dplyr::mutate()`. However, we could have recoded into a new variable by changing the name of *gender*.

```{r}

d4 %>%
  dplyr::mutate(gender = dplyr::recode(gender, m = "male", f = "female", n = "nonbinary"))

```


**2\. Recode a character variable (`gender`) into a numeric variable**

Review the data (d4)

```{r, echo=FALSE}

source("data.R")
d4

```

Recode `gender`

* Note: Since we are recoding into a numeric variable, no quotes are necessary for the numeric new value.

* Note: We are recoding back into the same variable using `dplyr::mutate()`. However, we could have recoded into a new variable by changing the name of *gender*.

```{r}

d4 %>%
  dplyr::mutate(gender = dplyr::recode(gender, m = 1, f = 2, n = 3))

```


**3\. Recode just one value in a numeric variable (`Var2`)**

Review the data (d5)

```{r, echo=FALSE}

source("data.R")
d5

```

Recode 16 -> 4 in `Var2`

Note: We *did* have to put tick marks around the old values because `dplyr::recode()` replaces numeric values by either their name or their position. If a number with no quotes/backticks is given, it will assume it is a position.

Note: Notice here that we are only recoding one value in the variable. There is a *default* option that you can set all not recoded values to, and if no default is supplied and the replacement values are the same variable class as the original values (ex: numeric and numeric), then unmatched values are unchanged. *However*, if the replacement value is a new variable class (ex: numeric to character) then all other values will be recoded to NA.


```{r}

d5 %>%
  dplyr::mutate(Var2 = dplyr::recode(Var2, `16` = 4))

```


**4\. Recode just one value in a character variable (`Var`)**

Review the data (d5)

```{r, echo=FALSE}

source("data.R")
d14

```

In this case I want Var1 to be numeric. However, missing values were entered as character values ("MR" for missing response) so I need to recode these values before I can convert this variable to numeric.

* Note: Parentheses *are* required around the new value when recoding because the values in Var1 are still considered character even though they all appear as numeric after recoding.


```{r}

d14 <- d14 %>%
  dplyr::mutate(Var1 = dplyr::recode(Var1, MR = "-99"))

d14

```

Note: Notice again that we are only recoding one value in the variable. If no default replacement option is supplied and the replacement values are the same variable class as the original values (ex: character to character), then unmatched values are unchanged. 

I can then now convert my variable to numeric without having my character values converted to NAs. If I tried to convert my Var1 variable to numeric before recoding "MR" to -99, the `dplyr::recode()` function would convert "MR" to NA.

```{r}

d14 %>%
  dplyr::mutate(Var1 = base::as.numeric(Var1))

```


**5\. Recode one value in a numeric haven labelled variable (`Var3`)**

Review the data (d5)

```{r, echo=FALSE}

source("data.R")
d5

```

View the variable labels for `Var3`

* Note: You'll see in this example there are three values and two value labels. In this instance, 2s are supposed to also be yes, but they do not contain a value label.

```{r, echo=FALSE}

d5 <- d5 %>% 
  labelled::set_value_labels(Var3 = c(no = 0, yes = 1))

```


```{r}

labelled::val_labels(d5$Var3)

```

Recode 2 -> 1 in `Var3`

```{r}

d5 <- d5 %>% 
  dplyr::mutate(Var3 = dplyr::recode(Var3, `2`=1))

d5

```

View the variable labels for `Var3`

Note: The value labels are retained. If the `labelled` package is loaded, additional functionality is available for `dplyr::recode()`, such as additional arguments for value labels. For example, there is an argument, *.keep_value_labels*, that you can add to `dplyr::recode()` and the default is TRUE. However, if you do not want to keep the value labels you can set the option to FALSE.


```{r}

labelled::val_labels(d5$Var3)

```

**6\. Recode multiple values in a numeric haven labelled variable (`Var3`)**

Review the data (d13)

```{r, echo=FALSE}

source("data.R")
d13

```

View the variable labels for `Var3`


```{r, echo=FALSE}

d13 <- d13 %>% 
  labelled::set_value_labels(Var3 = c(no = 5, yes = 6))

```

```{r}

d13 %>%
  dplyr::select(Var3) %>%
  labelled::val_labels()

```

Recode 5 -> 0 and 6 -> 1 in `Var3`


```{r}


d13 <- d13 %>%
  dplyr::mutate(Var3 = dplyr::recode(Var3, `5` = 0, `6` = 1))

d13

```

View the variable labels for `Var3`

Note: We left the `labelled` and `dplyr::recode()` default of *.keep_value_labels* is TRUE so our value labels are still funky.

```{r}

d13 %>%
  dplyr::select(Var3) %>%
  labelled::val_labels()

```


In this instance I don't want to keep the original value labels. I don't necessarily need to set the *.keep_value_labels* to FALSE though. I can just add new value labels and they will overwrite the previous labels.

```{r}

d13 <- d13 %>% 
  labelled::set_value_labels(Var3 = c(no = 0, yes = 1))

d13 %>%
  dplyr::select(Var3) %>%
  labelled::val_labels()

```

<br>


---

#### Function: `case_when()`

---


**1\. Recode `Var3` based on `decision_var`**

Review the data (d21)

```{r, echo=FALSE}

source("data.R")
d21

```

My logic is

If `decision_var` != 4 then `Var3` should be -99,
else it should be it's current value

```{r}

d21 %>%
  dplyr::mutate(
    Var3 =
      dplyr::case_when(
        !decision_var %in% 4 ~ -99,
        TRUE ~ Var3
      )
  )

```

Notice that I use the `!` outside of `decision_var` and use `%in%` instead of `!=`. This is just one way to deal with the NA values that exist in `decision_var`. In this case I wanted values to be recoded to -99 even if `decision_var` is NA.

Let's see what would happen if I wrote this the other way.

```{r}

d21 %>%
  dplyr::mutate(
    Var3 =
      dplyr::case_when(
        decision_var != 4 ~ -99,
        TRUE ~ Var3
      )
  )

```

Using `!=` above, the NAs in `decision_var` are not evaluated and so the value of `Var3` remains the same value. I could solve this problem another way by adding an explicit call to recode NAs like this below.

```{r}

d21 %>%
  dplyr::mutate(
    Var3 =
      dplyr::case_when(
        decision_var != 4 ~ -99,
        is.na(decision_var) ~ -99,
        TRUE ~ Var3
      )
  )

```

And last, if I wanted NAs in the `decision_var` to be dealt with in a different way from the other values, I would **not** want to use %in%. See example below.

My new logic is

If `decision_var` !=4 then `Var3` should be -99,
If `decision_var` is NA then `Var3` should be NA,
else it should be it's current value

```{r}

d21 %>%
  dplyr::mutate(
    Var3 =
      dplyr::case_when(
        decision_var != 4 ~ -99,
        is.na(decision_var) ~ NA_real_,
        TRUE ~ Var3
      )
  )

```

**2\. Recode `Var3` into a dichotomous variable**

Review the data (d22)

```{r, echo=FALSE}

source("data.R")
d22

```

My logic is

If `Var3` !=4 then 0,
else 1

```{r}

d22 %>%
  dplyr::mutate(
    Var3 =
      dplyr::case_when(
        !Var3 %in% 4 ~ 0,
        TRUE ~ 1
      )
  )

```

Similar to above, using `!` and `%in%` ensures that the NA is accounted for. If however, I wanted NA to stay NA, I would want to write something like this.

```{r}

d22 %>%
  dplyr::mutate(
    Var3 =
      dplyr::case_when(
        Var3 !=4 ~ 0,
        is.na(Var3) ~ NA_real_,
        TRUE ~ 1
      )
  )

```



Return to [Recode](https://github.com/Cghlewis/data-wrangling-functions/wiki/Recode)