---
title: "Reverse code values"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: dplyr

---

#### Function: `recode()`

---

**Note:**  The `dplyr::recode()` formula is: old value=new value, this is opposite of `dplyr::rename()`

<br>

**1\. Reverse code multiple variables (Var2 and Var3) in the same way**

Review the data (d2)

```{r, echo=FALSE}

source("data.R")
d2

```

Check the values of Var2 and Var3

* Note: `base::lapply()` is used to iterate the `base::table()` function across all variables.

```{r}

lapply(d2, table, useNA="always")

```

Reverse code Var2 and Var3 and rename the variables with *_r* at the end by using the `dplyr::rename_at()` function as well as `base::paste0()`.

* Note: I wrote this all into a new dataset because this syntax, using `dplyr::mutate()` and `dplyr::across()`, writes over our existing Var2 and Var3 and in this scenario I wanted to keep the old variables.

* Note: You will get similar results using `dplyr::mutate_at()` rather than `dplyr::mutate()` in conjunction with `dplyr::across()`.

* Note:  The `dplyr::recode()` formula is: old value=new value, this is opposite of `dplyr::rename()`

* Note:  The `dplyr::recode()` function considers the old value as a *name* and therefore, in order to use a number as a name/character, you have to surround it with backticks (``)


```{r}

d2_new <- d2 %>% 
  dplyr::mutate(dplyr::across(c(Var2:Var3), ~dplyr::recode(.,`1`=5, `2`=4, `3`=3, `4`=2, `5`=1))) %>%
  dplyr::rename_at(vars(Var2:Var3), ~paste0(., "_r"))

```


Check that your reverse code worked correctly

```{r}

table(d2$Var2, d2_new$Var2_r)

table(d2$Var3, d2_new$Var3_r)

```

---

#### Function: `case_when()`

---

**1\. Reverse code a single variable**

Review the data (d2)

```{r, echo=FALSE}

source("data.R")
d2

```

Check the values of Var2

```{r}

d2 %>% janitor::tabyl(Var2)

```

Reverse code Var2 and name it with *_r* at the end. 

* Note: We are recoding into a new variable using `dplyr::mutate()` and naming the new variable a different name than the original. It keeps both the new and old versions of the variable. If we used `dplyr::transmute()` instead it would keep the new version and remove the old version.

```{r}

d2 <- d2 %>% dplyr::mutate(Var2_r = dplyr::case_when(
  Var2 == 1 ~ 5,
  Var2 == 2 ~ 4,
  Var2 == 3 ~ 3, 
  Var2 == 4 ~ 2,
  Var2 == 5 ~ 1,
  TRUE ~ NA_real_))

```

Check that your reverse code worked correctly

```{r}

d2 %>% janitor::tabyl(Var2, Var2_r)

```


**2\. Turn your reverse scale into a function to use for multiple variables (Data Science in Education Using R example)**

Review the data (d2)

```{r, echo=FALSE}

source("data.R")
d2

```

Check the values of Var2 and Var3

```{r}

d2 %>% janitor::tabyl(Var2)

d2 %>% janitor::tabyl(Var3)

```

Create your function

```{r}

reverse_scale <- function(Var) {
  x <- dplyr::case_when(
    Var == 1 ~ 5,
    Var == 2 ~ 4,
    Var == 3 ~ 3, 
    Var == 4 ~ 2,
    Var == 5 ~ 1,
    TRUE ~ NA_real_
  )
}

```


Reverse code Var2 and Var3 using your function and rename the new variables with *_r* at the end

* Note: We are recoding into new variables using `dplyr::mutate()` and naming the new variables different names than the original. It keeps both the new and old versions of the variables. 

```{r}

d2 <- d2 %>%
  dplyr::mutate(Var2_r = reverse_scale(Var2),
         Var3_r = reverse_scale(Var3))

```

Check that the recode worked

```{r}

d2 %>% janitor::tabyl(Var2, Var2_r)

d2 %>% janitor::tabyl(Var3, Var3_r)

```

You can also use your function this way

```{r, echo=FALSE}
source("data.R")

```

* Note: Again I wrote this all into a new dataset because this syntax, using `dplyr::mutate()`, writes over our existing Var2 and Var3 and in this scenario I wanted to keep the old variables.

* Note: We are using `dplyr::across()` to apply a transformation to multiple columns.

* Note: Renaming here is done using the `dplyr::rename_with()` function as well as `base::paste0()`.

```{r}

d2_new <- d2 %>% dplyr::mutate(across(c(Var2:Var3), reverse_scale)) %>%
  dplyr::rename_with(~paste0(., "_r"), .cols = c(Var2:Var3))

```

Check that the recode worked

```{r}

table(d2$Var2, d2_new$Var2_r)

table(d2$Var3, d2_new$Var3_r)

```

---

#### Function: `if_else()`

---

Review the data (d2)

```{r, echo=FALSE}

source("data.R")
d2

```


Check the values of Var2

```{r}

d2 %>% janitor::tabyl(Var2)

```

Reverse code Var2 and rename it with *_r* at the end

* Note: We are recoding into new variables using `dplyr::mutate()` and naming the new variable a different name than the original. It keeps both the new and old versions of the variable. 

```{r}

d2 <- d2 %>% dplyr::mutate(Var2_r = dplyr::if_else(Var2 == 1, 5,
                  dplyr::if_else(Var2==2, 4,
                          dplyr::if_else(Var2==4, 2,
                                  dplyr::if_else(Var2==5, 1, Var2)))))

```


Check that your reverse code worked correctly

```{r}

d2 %>% janitor::tabyl(Var2, Var2_r)

```
