---
title: "Reverse code values"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: dplyr

---

#### Function: `dplyr::recode()`

---

**1\. Reverse code multiple variables in the same way**

Review the data (d2)

```{r, echo=FALSE}

source("data.R")
d2

```

Check the values of Var2 and Var3

```{r}

lapply(d2, table, useNA="always")

```

Reverse code Var2 and Var3 and rename them with *_r* at the end

**Note:**  The dplyr::recode formula is: old value=new value, this is opposite of dplyr::rename

**Note:**  The dplyr::recode function considers the old value as a *name* and therefore, in order to use a number as a name, you have to surround it with backticks (``)


```{r}

d2_new <- d2 %>% 
  dplyr::mutate(across(c(Var2:Var3), ~dplyr::recode(.,`1`=5, `2`=4, `3`=3, `4`=2, `5`=1))) %>%
  dplyr::rename_at(vars(Var2:Var3), ~paste0(., "_r"))

```

Check that your reverse code worked correctly

```{r}

table(d2$Var2, d2_new$Var2_r)

table(d2$Var3, d2_new$Var3_r)

```

---

#### Function: `dplyr::case_when()`

---

**1\. Reverse code a single variable**

Review the data (d2)

```{r, echo=FALSE}

source("data.R")
d2

```

Check the values of Var2

```{r}

d2 %>% janitor::tabyl(Var2)

```

Reverse code Var2 rename it with *_r* at the end

```{r}

d2 <- d2 %>% dplyr::mutate(Var2_r = dplyr::case_when(
  Var2 == 1 ~ 5,
  Var2 == 2 ~ 4,
  Var2 == 3 ~ 3, 
  Var2 == 4 ~ 2,
  Var2 == 5 ~ 1,
  TRUE ~ NA_real_))

```

Check that your reverse code worked correctly

```{r}

d2 %>% janitor::tabyl(Var2, Var2_r)

```


**2\. Turn your reverse scale into a function to use for multiple variables (Data Science in Education Using R example)**

Create your function

```{r}

reverse_scale <- function(Var) {
  x <- dplyr::case_when(
    Var == 1 ~ 5,
    Var == 2 ~ 4,
    Var == 3 ~ 3, 
    Var == 4 ~ 2,
    Var == 5 ~ 1,
    TRUE ~ NA_real_
  )
  x
}

```

Review the data (d2)

```{r, echo=FALSE}

source("data.R")
d2

```

Check the values of Var2 and Var3

```{r}

d2 %>% janitor::tabyl(Var2)

d2 %>% janitor::tabyl(Var3)

```

Reverse code Var2 and Var3 using your function and rename it with *_r* at the end

```{r}

d2 <- d2 %>%
  dplyr::mutate(Var2_r = reverse_scale(Var2),
         Var3_r = reverse_scale(Var3))

```

Check that the recode worked

```{r}

d2 %>% janitor::tabyl(Var2, Var2_r)

d2 %>% janitor::tabyl(Var3, Var3_r)

```

You can also use your function this way

```{r, echo=FALSE}
source("data.R")
d2

```

```{r}

d2_new <- d2 %>% dplyr::mutate_at(vars(Var2:Var3), reverse_scale) %>%
  dplyr::rename_at(vars(Var2:Var3), ~paste0(., "_r"))

```

Check that the recode worked

```{r}

table(d2$Var2, d2_new$Var2_r)

table(d2$Var3, d2_new$Var3_r)

```

---

#### Function: `dplyr::if_else()`

---

Review the data (d2)

```{r, echo=FALSE}

source("data.R")
d2

```


Check the values of Var2

```{r}

d2 %>% janitor::tabyl(Var2)

```

Reverse code Var2 rename it with *_r* at the end

```{r}

d2 <- d2 %>% dplyr::mutate(Var2_r = dplyr::if_else(Var2 == 1, 5,
                  if_else(Var2==2, 4,
                          if_else(Var2==4, 2,
                                  if_else(Var2==5, 1, Var2)))))

```


Check that your reverse code worked correctly

```{r}

d2 %>% janitor::tabyl(Var2, Var2_r)

```
