---
title: "Reverse code values"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: dplyr

---

#### Function: `recode()`

---

**Note:**  The `dplyr::recode()` formula is: old value=new value, this is opposite of `dplyr::rename()`

<br>

**1\. Reverse code multiple variables (`Var2` and `Var3`) in the same way**

Review the data (d2)

```{r, echo=FALSE}

source("data.R")
d2

```

Reverse code `Var2` and `Var3` and rename the variables with *\_r* at the end.

* Note: We are modifying existing variables using `dplyr::mutate()`. 

* Note: We are adding `dplyr::across()` to select the variables we want to modify.

* Note:  The `dplyr::recode()` function considers the old value as a *name* and therefore, in order to use a number as a name/character, you have to surround it with backticks (``)

* Note: Last, in this scenario we are writing over the existing variables and then renaming them by using the `dplyr::rename_with()` function along with `base::paste0()`. 


```{r}

d2_new <- d2 %>%
  dplyr::mutate(dplyr::across(Var2:Var3, ~ dplyr::recode(
    .,
    `1` = 5,
    `2` = 4,
    `3` = 3,
    `4` = 2,
    `5` = 1
  ))) %>%
  dplyr::rename_with( ~ paste0(., "_r"), Var2:Var3)

d2_new

```

We could also do this using the formula (1 + highest possible scale value) - original value. In this case I know that the highest possible value for this scale is 5.

Note that I don't need to use the `dplyr::recode()` function here. I can simply apply a formula across my selected variables.

```{r}

d2_new <- d2 %>%
  dplyr::mutate(dplyr::across(Var2:Var3, ~ 6 - .x)) %>%
  dplyr::rename_with( ~ paste0(., "_r"), Var2:Var3)

d2_new

```


Check that your reverse code worked correctly

```{r}

base::table(d2$Var2, d2_new$Var2_r)

base::table(d2$Var3, d2_new$Var3_r)

```

If you want to keep *both* the old variables and the newly created variables, you can add the argument *.names* to the `dplyr::across()` function to rename your recoded variables as you create them, rather than writing over the old variables by adding the `dplyr::rename_with()` function as a second step. 

```{r}

d2 <- d2 %>%
  dplyr::mutate(dplyr::across(
    Var2:Var3,
    ~ dplyr::recode(
      .,
      `1` = 5,
      `2` = 4,
      `3` = 3,
      `4` = 2,
      `5` = 1
    ),
    .names = '{col}_r'
  ))

d2

```


---

#### Function: `case_when()`

---

**1\. Reverse code a single variable**

Review the data (d2)

```{r, echo=FALSE}

source("data.R")
d2

```

Check the values of `Var2`

```{r}

d2 %>% 
  janitor::tabyl(Var2)

```

Reverse code `Var2` and name it with *_r* at the end. 

* Note: We are recoding into a new variable using `dplyr::mutate()` and naming the new variable a different name than the original. It keeps both the new and old versions of the variable. If we used `dplyr::transmute()` instead it would keep the new version and remove the old version.

```{r}

d2 <- d2 %>% 
  dplyr::mutate(Var2_r = dplyr::case_when(
  Var2 == 1 ~ 5,
  Var2 == 2 ~ 4,
  Var2 == 3 ~ 3, 
  Var2 == 4 ~ 2,
  Var2 == 5 ~ 1,
  TRUE ~ NA_real_))

```

Check that your reverse code worked correctly

```{r}

d2 %>% 
  janitor::tabyl(Var2, Var2_r)

```


**2\. Turn your reverse scale into a function to use for multiple variables (Data Science in Education Using R example)**

Review the data (d2)

```{r, echo=FALSE}

source("data.R")
d2

```

Check the values of `Var2` and `Var3`

```{r}

d2 %>% 
  janitor::tabyl(Var2)

d2 %>% 
  janitor::tabyl(Var3)

```

Create your function

```{r}

reverse_scale <- function(Var) {
  x <- dplyr::case_when(
    Var == 1 ~ 5,
    Var == 2 ~ 4,
    Var == 3 ~ 3, 
    Var == 4 ~ 2,
    Var == 5 ~ 1,
    TRUE ~ NA_real_
  )
}

```


Reverse code `Var2` and `Var3` using your function and rename the new variables with *_r* at the end

* Note: We are recoding into new variables using `dplyr::mutate()` and naming the new variables different names than the original. It keeps both the new and old versions of the variables. 

```{r}

d2 <- d2 %>%
  dplyr::mutate(Var2_r = reverse_scale(Var2),
                Var3_r = reverse_scale(Var3))

```

Check that the recode worked

```{r}

d2 %>% 
  janitor::tabyl(Var2, Var2_r)

d2 %>% 
  janitor::tabyl(Var3, Var3_r)

```

You can also use your function this way

```{r, echo=FALSE}
source("data.R")

```

Reverse code `Var2` and `Var3` using your function and rename the new variables with *\_r* at the end

* Note: We are modifying existing variables using `dplyr::mutate()`

* Note: We are using `dplyr::across()` to apply a transformation to select columns and then adding the argument *names* to rename your recoded variables as you create them, rather than writing over the old variables.

```{r}

d2 <- d2 %>%
  dplyr::mutate(across(Var2:Var3, reverse_scale,
                       .names = '{col}_r'))

```

Check that the recode worked

```{r}

base::table(d2$Var2, d2$Var2_r)

base::table(d2$Var3, d2$Var3_r)

```


Return to [Recode](https://github.com/Cghlewis/data-wrangling-functions/wiki/Recode)