---
title: "Calculate column values"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: dplyr

---

#### Function: `summarise()`

---

**1\. Calculate a mean value for `item1` across all students**

Review the data (d2)

```{r, echo=FALSE}

source("data.R")
d2

```

Calculate a mean for `item1`

* Note: The default for `base::mean()` is to **not** calculate a sum if any NA value exists. If you want to still calculate a sum despite missing values, you can add the argument *na.rm = TRUE*.

```{r}

d2 %>%
  summarise(item1_mean = mean(item1, na.rm = TRUE))


```

 
**2\. Calculate a sum value for `n_students` across all schools**

Review the data (d9)

```{r, echo=FALSE}

source("data.R")
d9

```

Calculate the total number of students across all schools.

```{r}

d9 %>% 
  summarise(total_students = sum(n_students, na.rm = TRUE))

```


Return to [Calculate Sums and Means](https://github.com/Cghlewis/data-wrangling-functions/wiki/Calculate-Sums-and-Means)
