---
title: "Calculate percent adherence"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: dplyr

---

#### Function: `summarise()`

---

**1\. Calculate percent adherence for `item1_t1`**

Review the data (obs_data)

```{r, echo=FALSE}

source("data.R")
obs_data

```

Here we want to know the percentage of teachers that implemented `item1` in time 1. In this variable

1 = implemented  
0 = did not implement

* Note: The default for `sum()` is to **not** calculate a sum if any NA value exists. If you want to still calculate a sum despite missing values, you can add the argument *na.rm = TRUE*.

```{r}

obs_data %>%
  dplyr::summarise(item1_t1_per = sum(item1_t1, na.rm = TRUE)/n())

```

**2\. Calculate percent adherence for `item1` across all time periods**

Review the data (obs_data)

```{r, echo=FALSE}

source("data.R")
obs_data

```

Since our data is in wide format, this is not a simple column calculation. We first need to restructure our wide data into long format so that all item1 values are in the same column.

First restructure

```{r}

obs_data_long <- obs_data %>%
  tidyr::pivot_longer(
    cols = starts_with("item"),
    names_sep = ("_"),
    names_to = c(".value", "time"))

head(obs_data_long)

```

Then calculate percent adherence

```{r}

obs_data_long %>%
  dplyr::summarise(item1_per = sum(item1, na.rm = TRUE)/n())

```


Return to [Calculate Row Values](https://github.com/Cghlewis/data-wrangling-functions/wiki/Calculate-Values)
