---
title: "Calculate row values of NAs"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: base

---

#### Function: `rowSums()`

---

**1\. Calculate the sum of NA values across all items in the "toca" measure**

Review the data (d5)

```{r, echo=FALSE}

source("data.R")
d5

```

Calculate the sum of NAs

* Note: We are calculating a new variable using `dplyr::mutate()`

* Note: Adding `dplyr::across()` allows you to select the specific columns you want to calculate the `base::rowSums()` for. Otherwise rowSums will be applied across all columns.

* Note: We use the `tidyselect` selection helper `contains()` to refer to all variables that contain with the word "toca".

* Note: Adding `base::is.na()` returns a logical value which can be counted/summed.

```{r}

d5 %>% 
  dplyr::mutate(toca_na_sum = rowSums(is.na(dplyr::across(tidyselect::contains("toca")))))

```

We could also select the variables in other ways and it would work just as well.

```{r}

d5 %>% 
  dplyr::mutate(toca_na_sum = rowSums(is.na(dplyr::across(toca1:toca4))))

```

And last, we can also write this where, rather than wrapping the `is.na()` function around our `dplyr::across()` function, we call our variables first and then add `is.na()` in an anonymous function.

```{r}

d5 %>%
  dplyr::mutate(toca_na_sum = rowSums(dplyr::across(toca1:toca4, ~ is.na(.))))

```


Return to [Calculate Row Values](https://github.com/Cghlewis/data-wrangling-functions/wiki/Calculate Row Values)
