---
title: "Calculate row occurrences of strings"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: base

---

#### Function: `rowSums()`

---

**1\. Calculate all occurrences of "prefer not to answer" across a selection of variables (q1:q4)**

Review the data (d7)

```{r, echo=FALSE}

source("data.R")
d7

```

Here, using `dplyr::mutate()` I can create a new variable `prefer_sum`.

Then I can use `base::rowSums()` to calculate the occurrences of the string "prefer", detected using `stringr:str_detect()`, across my variables, selected using `dplyr::across()`.

I add the `rowSums()` argument *na.rm* = TRUE to calculate sums despite NA values in my variables.

```{r}

d7 %>%
  dplyr::mutate(prefer_sum = 
           rowSums(dplyr::across(q1:q4, ~ stringr::str_detect(., "prefer")), na.rm = TRUE))

```

I could also write this using the exact text.

```{r}

d7 %>%
  dplyr::mutate(prefer_sum = 
           rowSums(dplyr::across(q1:q4, ~ . == "prefer not to answer"), na.rm = TRUE))

```

**2\. Calculate a sum of occurences of kids between ages 1 and 17 (kid1:kid3)**

Review the data (d8)

```{r, echo=FALSE}

source("data.R")
d8

```

Similar to above I can use `stringr::str_detect()` in combination with regex to only grab values 1-17.

```{r}

d8 %>%
  dplyr::mutate(minor_kids = 
           rowSums(dplyr::across(kid1:kid3, ~ stringr::str_detect(., "^[1-9]$|^[1]{1}[0-7]$")), na.rm = TRUE))

```

But also, since the values are all numeric, I could simplify the code to this.

```{r}

d8 %>%
  dplyr::mutate(minor_kids = 
           rowSums(dplyr::across(kid1:kid3, ~ . <18), na.rm = TRUE))

```


Return to [Calculate Row Values](https://github.com/Cghlewis/data-wrangling-functions/wiki/Calculate Row Values)
