---
title: "Calculate percent agreement"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: base

---

#### Function: `rowMeans()`

---

**1\. Calculate the percent agreement across all variables for each teacher/pair of raters**

Review the data (rater_data)

```{r, echo=FALSE}

source("data.R")
rater_data

```

First get an agreement score for each variable.
0 = no agreement, 1 = agreement

We can do this using `base::diff()` and then recoding values to 0 or 1.

```{r}

rater_summary <- rater_data %>% 
  group_by(tch_id) %>%
  dplyr::summarise(dplyr::across(item1:item3, diff)) %>%
  dplyr::mutate(dplyr::across(item1:item3,
                       ~ dplyr::case_when(
                         . == 0 ~ 1,
                         TRUE ~ 0
                       )))

rater_summary

```

Next we can calculate a row mean (using `base::rowMeans()`) to get a percent agreement per teacher/rater pair.

```{r}

rater_summary %>%
  dplyr::mutate(percent_agree =
           round(rowMeans(dplyr::across(item1:item3), na.rm = TRUE),
                       digits = 2))

```


Return to [Calculate Values](https://github.com/Cghlewis/data-wrangling-functions/wiki/Calculate-Values)
