---
title: "Calculate row values"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: base

---

#### Function: `rowSums()`

---

**1\. Calculate a sum score for each student called `measure_sum`**

Review the data (d2)

```{r, echo=FALSE}

source("data.R")
d2

```

Calculate a sum score across `item1`, `item2` and `item`

* Note: We are calculating a new variable using `dplyr::mutate()`

* Note: Adding `dplyr::across()` allows you to select the specific columns you want to calculate the `rowSums()` for. Otherwise `rowSums` will be applied across all columns.
 
* Note: The default for `base::rowSums()` is to **not** calculate a sum if any NA value exists. If you want to still calculate a sum despite missing values, you can add the argument *na.rm = TRUE*.

```{r}

d2 %>% 
  mutate(measure_sum = rowSums(across(c(item1, item2, item3))))

```

**2\. Calculate a sum score for each student called `measure_sum` no matter what values are missing**

Review the data (d10)

```{r, echo=FALSE}

source("data.R")
d10

```

Calculate a sum score across `item1`, `item2` and `item` even if values are missing

```{r}

d10 %>% 
  mutate(measure_sum = rowSums(across(c(item1, item2, item3)), na.rm = TRUE))

```

NOTICE that we got a value of **0** when all columns have missing values. Most likely this is not what you want. So in order to return of value of NA when all columns have missing values, you need to add in logic.

Here I'm using `dplyr::case_when()` to first recode our new variable to `NA` when all items are `NA`, and then if all items are not `NA` we calculate `rowSums`.

```{r}

d10 %>%
  mutate(measure_sum =
           case_when(
             if_all(item1:item3, ~ is.na(.)) ~ NA,
             .default = rowSums(across(item1:item3), na.rm = TRUE)))

```


---

#### Function: `rowMeans()`

---


**1\. Calculate a mean score (`measure_mean`) for each student even with missing data**

Review the data (d2)

```{r, echo=FALSE}

source("data.R")
d2

```

Calculate the mean score of `item1`, `item2`, and `item3`
 
* Note: The default for `base::rowMeans()` is to **not** calculate a mean if any NA value exists. If you want to still calculate a mean despite missing values, you can set the argument to *na.rm = TRUE*.

* Note: I think it's **really important** to point out that the closing parentheses for `dplyr::across()` goes at the end of `item3`, rather than at the end of of the `rowMeans` function. This is different than how you add parentheses when using `dplyr::across()` for something like a `dplyr::case_when()` function. And if you do not put the parentheses at the end of `item3`, the `rowMeans` function will still run, but the *na.rm = TRUE* argument will not be evaluated!

```{r}

d2 %>% 
  mutate(measure_mean = rowMeans(across(item1:item3), na.rm = TRUE))

```

You can set the number of decimal places you want the new variable to have by wrapping the calculation in the function `base::round()` and setting the digits equal to the number you want.

* Note: It is very important to note how `base::round()` does rounding. See [Rounding](https://github.com/Cghlewis/data-wrangling-functions/wiki/Rounding) for more information.

```{r}

d2 %>% 
  mutate(measure_mean = round(rowMeans(across(item1:item3), na.rm=TRUE), digits= 1))

```

**2\. Calculate a mean score (`toca_mean`) for all the variables that begin with "toca", for each student even with missing data**

Review the data (d5)

```{r, echo=FALSE}

source("data.R")
d5

```

Calculate the mean toca score

* Note: Here we are using `tidyselect::starts_with()` to grab all variables that start with "toca"

```{r}

d5 %>% 
  mutate(
    toca_mean = rowMeans(
      across(starts_with("toca")), na.rm=TRUE))

```

If you wanted to exclude some of the "toca" items (say just `toca2`), you could exclude them this way, by wrapping your selected and excluded variables in `c()`. 

```{r}

d5 %>% 
  mutate(
    toca_mean = rowMeans(
      across(c(starts_with("toca"), -toca2)), na.rm=TRUE))

```

**3\. Calculate a mean score (`toca_mean`) for all the variables that contain the word "toca", for each student who is missing one or less values for an item**

Review the data (d5)

```{r, echo=FALSE}

source("data.R")
d5

```

Calculate the mean toca score

```{r}

d5 %>% 
  mutate(
    toca_mean = 
      case_when(
        rowSums(is.na(across(starts_with("toca")))) <= 1 ~ rowMeans(across(starts_with("toca")), na.rm=TRUE),
    .default = NA_real_))

```


**4\. Calculate a mean score (`toca_mean`) for all the variables that begin with "toca", when the toca variables contain _labelled na values_**

Review the data (d6)

```{r, echo=FALSE}

source("data.R")
d6

```

Let's review the labelled NA values. 

```{r, echo = FALSE}

library(labelled)

d6 <- d6 %>%
  mutate(across(contains("toca"), 
  ~(`na_values<-`(., c(-99)))))

```

```{r}

d6 %>%
  labelled::na_values()

```

We see that **-99** is labelled as an NA value for the toca variables. However, R does not recognize this as an NA value when calculating row scores. 

(1) So first we will need to recode these to NA before creating our mean score.

* Note: We use `base::ifelse()` to say if our current toca variables (denoted by .) are equal to -99, then make them NA, otherwise keep the value as is. Using `dplyr::if_else()` gives us errors in this process. I'm not sure why except that I read the function is more strict in checking that all alternatives are of the same variable type compared to using `base::ifelse()`. 

* Note: We use the `dplyr::across()` *.names* argument to rename these variable to have the suffix "_temp" because we do not want to alter our original toca variables.

(2) Second we calculate our mean score within `dplyr::mutate()` statement.

(3) Last we drop our temporary toca variables using `dplyr::select()`. We don't need them anymore.


```{r}

d6 %>%
  mutate(across(contains("toca"),
                ~ ifelse(. == -99 , NA, .), .names = "{col}_temp"),
         toca_mean = rowMeans(across(contains("temp")),
                              na.rm = TRUE)) %>%
  select(-contains("temp"))

```

We could also use `haven::zap_missing()` rather than `base::ifelse()`

```{r}

d6 %>%
  mutate(
    across(contains("toca"), haven::zap_missing, .names = "{col}_temp"),
    toca_mean = rowMeans(across(contains("temp")),
                         na.rm = TRUE)) %>%
  select(-contains("temp"))

```

<br>

### Package: dplyr

---

#### Function: `rowwise()`

---

**1\. Calculate a sum score for each student called `measure_sum`**

Review the data (d2)

```{r, echo=FALSE}

source("data.R")
d2

```

Calculate the sum score across `item1`, `item2`, and `item3`.

Here we use `dplyr::rowwise()` to calculate values across rows, in conjunction with `base::sum()` to calculate a sum score.

* Note: The default for `base::sum()` is to not calculate a sum if any NA value exists in your columns. If you want to still calculate a sum despite missing values, you can add the argument *na.rm = TRUE*.

```{r}

d2 %>%
  rowwise() %>%
  mutate(measure_sum = sum(c(item1, item2, item3)))

```

If you don't want to list out every single item in a measure, you can use `dplyr::c_across()` which uses tidy selection syntax to select variables. It is essentially a combination of `base::c()` and `dplyr::across()`. So instead of `mutate(measure_sum = dplyr::across(c(item1:item3)))`, you can simplify to the code below.

* Note: `dplyr::c_across()` cannot be used with `base::rowSums()` and `base::rowMeans()` because those functions take a different type of input compared to `dplyr::rowwise()`

```{r}

d2 %>%
  rowwise() %>%
  mutate(measure_sum = sum(c_across(item1:item3)))


```

**2\. Calculate a mean score (`measure_mean`) for each student even with missing data**

Review the data (d2)

```{r, echo=FALSE}

source("data.R")
d2

```

Calculate the mean score 

* Note: If all column values are NA, you will get a value of 0 returned. If you don't want this, you will need to add logic (like we saw in the above `rowSums` example, where you specifically code those values to NA).

```{r}

d2 %>%
  rowwise() %>%
  mutate(measure_mean = mean(c_across(item1:item3), na.rm=TRUE))

```


You can set the number of decimal places you want the new variable to have by wrapping the calculation in the function `base::round()` and setting the digits equal to the number you want.

```{r}

d2 %>%
  rowwise() %>%
  mutate(measure_mean = 
           round(mean(c_across(item1:item3), na.rm=TRUE), digits=2))

```


**3\. Calculate a mean score (`toca_mean`) for all the variables that contain the word "toca", for each student even with missing data**

Review the data (d4)

```{r, echo=FALSE}

source("data.R")
d4

```

Calculate the mean toca score

```{r}

d4 %>%
  rowwise() %>%
  mutate(toca_mean = 
           mean(c_across(contains("toca")), na.rm=TRUE))

```

If you wanted to exclude some of the "toca" items (say `toca2` and `toca4` this time), you could exclude them this way, by wrapping your selected and excluded variables in `c()`. 

```{r}

d4 %>%
  rowwise() %>%
  mutate(toca_mean = 
           mean(c_across(c(contains("toca"),-toca2,-toca4)), 
                na.rm = TRUE))

```


<br>

### Package: janitor

---

#### Function: `adorn_totals()`

---

**1\. Calculate a sum score for each student called `measure_sum`**

Review the data (d2)

```{r, echo=FALSE}

library(janitor)

source("data.R")
d2

```


Calculate the sum score across `item1`, `item2` and `item3`

* Note: We can add the argument *where = "col"* to denote we want rows summed across columns rather than the default which is column totals

* Note: We can add the argument *name = * to rename the new column what we want rather than the default "Total"

* Note: The argument default for the NA argument is *na.rm = TRUE* so in this case, if we don't want a sum when there is missing data, we need to change it to FALSE.

```{r}

d2 %>% 
  adorn_totals(where = "col", name = "measure_sum", na.rm = FALSE)

```

Return to [Calculate Sums and Means](https://github.com/Cghlewis/data-wrangling-functions/wiki/Calculate-Sums-and-Means)
