---
title: "Restructure data to long format"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: tidyr

---

#### Function: `pivot_longer()`

---

* Note: There are many arguments available in `tidyr::pivot_longer()`. Type *?pivot_longer* in your console to see all arguments. But the most basic arguments to be aware of include:  
*cols* = cols you want to pivot,   
*names_to* = name of your new categorical column/s typically,  
*values_to* = name of the new value column/s

<br>

**1\. Restructure our "select all" items into long format**

Review the data (d9)

```{r, echo=FALSE}

source("data.R")
d9

```

Oftentimes when we collect survey data that includes a "select all" question, the data exports with each item displayed in its own column. With the response option shown if selected and NA shown if not selected. 

If we want to **visualize** the options in this select all question, it can be very helpful to restructure these items into long format.

```{r}

d9 %>%
  pivot_longer(
    cols = contains("tool"),
    names_to = "tools",
    values_to = "response"
  )

```

You can then manipulate the "response" column to create your visualization, potentially filtering out any NA values.

Your "select all" data may also come in this way, with the columns named for the response options and the responses given as 1 or NA.

```{r, echo=FALSE}

source("data.R")
d10

```

This data can be restructured just as easily.

Except instead of working with the "response" column, you will work with the "tools" column, potentially filtering data based on NA values in the "response" column.

```{r}

d10 %>%
  pivot_longer(
    cols = google_meet:other,
    names_to = "tools",
    values_to = "response"
  )

```

**2\. Restructure our longitudinal data into long format**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

In education research we are often wanting to restructure longitudinal data for **analysis** purposes.

In this example, we want to restructure the data into long format, making a new "wave" variable

* Note: We use a `tidyselect` selection helper `matches()` in the *cols* argument to select the columns we wish to pivot long. In this case we want any column that has the prefix "w#_". Using `matches()` rather than a selection helper such as `contains()`, we can use regular expressions to select columns.

* Note: Here we use the *names_to* argument to state that we want two new columns, the first one being wave and the second column being whatever our remaining value is after wave (in this case "q1"). ".value" defines the column containing the cell values and overrides the *values_to* argument.

* Note: We use the *names_sep* argument to state how the wave and question variable should be separated (by "_"). Using *names_pattern* is another option if you have more complicated variable names.

* Note: We use the *names_prefix* argument to remove the "w" that would be attached to every number in the wave column since that is redundant information now. Be aware that the wave column will still be considered character since it contained "w" before we removed it. You may choose to convert that column type.


```{r}

d1 %>%
  pivot_longer(
    cols = tidyselect::matches("w[0-9]_"),
                      names_to = c("wave", ".value"),
                      names_sep = "_",
    names_prefix = "w")

```

**3\. Restructure our longitudinal data into long format**

Review the data (d8)

```{r, echo=FALSE}

source("data.R")
d8

```

Unlike the prior example, we now have underscores used in our variable name to separate other pieces of metadata besides just our time component. Because of this, we can no longer simply use the argument *names_sep*. Instead we can use the *names_pattern* argument to develop a regular expression to describe the variable name and year groups of the variable.

* Note: Instead of selecting the columns I want to pivot, I chose to select the columns I did NOT want to pivot in the *cols* argument.

```{r}

d8 %>%
  pivot_longer(
    cols = -c(tch_id:treatment),
    names_to = c(".value", "year"),
    names_pattern = "^(.*)_([0-9]{4})$")

```


**4\. Restructure our data into tidy format**

Review the data (d5)

```{r, echo=FALSE}

source("data.R")
d5

```

The wide format above is not conducive for **calculating summary statistics by group**. So we need to reformat the data to long/tidy format. We want to end up with 4 columns (school, grade level, attendance and enrollment).

* Note: Similar to above we use the argument *cols* to select the columns we wish to pivot. Here we select all columns associated with a grade level (6 or 7). 

* Note: We use the *names_to* argument again but this time we switch the order. Since the value we want to group by is second in the variable name, we want to put that second in the *names_to* argument.

```{r}

d5 %>%
  pivot_longer(
    cols = enroll_6:attend_7,
    names_to = c(".value", "grade"),
    names_sep = "_")

```

If we only had one column type to pivot longer, we could simplify the arguments. We would no longer need to use the *names_sep* argument and we would no longer need to use ".value" for the *names_to* argument.

For example, here we only have enrollment to pivot longer.

```{r, echo = FALSE}

d6

```

And we can now just name the column we want the grouping variable to be in, and we use the argument *values_to* rather than *names_sep* to state the name of the values column.

```{r}

d6 %>%
  pivot_longer(
    cols = -school,
    names_to = "grade",
    values_to = "enroll")

```

**5\. Restructure our wide repeating data into one row per student**

Review the data (d11)

```{r, echo=FALSE}

source("data.R")
d11

```

We should be able to pivot similar to how we did with our longitudinal data. However, in order to split the information into multiple columns (student identifier column and then question item columns), we need to have similar patterns across all items. 

So first we will add "_stuid" to the end of all of the student identifier variables so the pattern will match the question items.

```{r}

d11 <- d11 %>%
  dplyr::rename_with(~stringr::str_c(., "_stuid"), .cols = stu1:stu2)

```


Now we can pivot using our same arguments from example #2, except now we can change our first *names_to* value to NA (which discards the column), because we have no need for a column that tells us if the student was #1 or #2.

```{r}

d11 %>%
  pivot_longer(
    cols = stu1_stuid:stu2_q2,
                      names_to = c(NA, ".value"),
                      names_sep = "_")

```

**6\. Restructure to one row per organization**

Review the data (d12)

```{r, echo=FALSE}

source("data.R")
d12

```

In this example, a rater may rate up to 3 organizations in a row of data. We want each of these organizations to be on their own line.

We are unable to use *names_sep* because the underscores in our variable names are used in different places, so we will use *names_pattern* again. We will also use the *values_drop_na* argument to remove any empty row that is created in our pivot process.

```{r}

d12 %>%
  pivot_longer(
    cols = -rater,
    names_to = c("organization", ".value"),
    names_pattern = "^(org[0-9]{1})_(.*)$",
    values_drop_na = TRUE)

```

What if our data also contained additional variables to denote that a rater has another organization to rate (1 = yes, 0 = no), as in (d13)?

We can use the same code, but remove the *values_drop_na* argument because now empty rows are not created.

```{r, echo=FALSE}

source("data.R")
d13

```

```{r}

d13 %>%
  pivot_longer(
    cols = -rater,
    names_to = c("organization", ".value"),
    names_pattern = "^(org[0-9]{1})_(.*)$")

```




Return to [Restructure](https://github.com/Cghlewis/data-wrangling-functions/wiki/Restructure)