---
title: "Clean variable names"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: purrr

---

#### Function: `purrr::set_names()`

---

**1\. Change all variable names to upper case**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Change all variable names to upper by adding the `base::toupper()` function

```{r}

d1 %>% purrr::set_names(toupper)

```


### Package: dplyr

---

#### Function: `dplyr::rename_with()`

---

**1\. Change all variable names to lower case**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Change all variable names to lower by adding the `base::tolower()` function

```{r}

d1 %>% dplyr::rename_with(tolower)

```


### Package: janitor

---

#### Function: `janitor::clean_names()`

---

**1\. Change all variable names to snake_case (lower case and underscore to separate words)**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d3

```

Change the variable names to snake_case (the default for `janitor::clean_names()`)

```{r}

d3 %>% janitor::clean_names()

```

**2\. Change all variable names to CamelCase (no spaces, indicating word separation through capital letters)**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d4

```

Change the variable names to camelCase

**Note:** To see all of the case options type *?to_any_case* in the console.

```{r}

d4 %>% janitor::clean_names(., "big_camel")

```
