---
title: "Clean variable names"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: purrr

---

#### Function: `set_names()`

---

**1\. Change all variable names to upper case**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Change all variable names to upper by adding the `base::toupper()` function

* Note: You can also change all variable names to lower using `base::tolower()` function.

```{r}

d1 %>% purrr::set_names(toupper)

```

* Note: There several other functions that convert strings to upper or lower including `stringr::str_to_lower()` and `stringr::str_to_upper()`.

```{r}

d1 %>% purrr::set_names(stringr::str_to_upper)

```



### Package: dplyr

---

#### Function: `rename_with()`

---

**1\. Change all variable names to lower case**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Change all variable names to lower by adding the `base::tolower()` function

```{r}

d1 %>% dplyr::rename_with(tolower)

```


### Package: janitor

---

#### Function: `clean_names()`

---

**1\. Change all variable names to snake_case (lower case and underscore to separate words)**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d3

```

Change the variable names to snake_case (the default for `janitor::clean_names()`)

```{r}

d3 %>% janitor::clean_names()

```

**2\. Change all variable names to UpperCamelCase (no spaces, indicating word separation through capital letters)**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d4

```

Change the variable names to UpperCamelCase

* Note: Here we can simply add the argument *case = "upper_camel"*. To see all of the case options type *?to_any_case* in the console.

```{r}

d4 %>% janitor::clean_names(., case = "upper_camel")

```
