---
title: "Concatenate to variable names"
output: 
  html_document:
  anchor_sections: FALSE
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>
  
### Package: dplyr
  
---
  
#### Function: `rename_with()`
  
---
  
* Note: Unlike `dplyr::rename()` where you can rename individual columns, `dplyr::rename_with()` renames columns using a function.

<br>

**1\. Concatenate a prefix, "T1_", to all variable names**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Concatenate T1_ by adding the `base::paste0` function within the `dplyr::rename_with()` function.

* Note: Using `base::paste0` concatenates the string directly to the variable name. Using `base::paste()` or     `stringr::str_c()` concatenates with a string separator.

```{r}

d1 %>% 
  dplyr::rename_with(~ base::paste0("T1_", .))

```

To make a suffix _T1 rather than a prefix, rewrite the function like this.

```{r}

d1 %>% 
  dplyr::rename_with(~ base::paste0(., "_T1"))

```

You can also use the `base::paste()` function or the `stringr::str_c()` function (they are equivalent) instead and add your separator using the *sep* argument.

```{r}

d1 %>% 
  dplyr::rename_with(~ base::paste(., "T1", sep= "_"))

```

**2\. Concatenate a prefix, "T1_", to only numeric variable names**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Concatenate T1_ by adding the `base::paste0` function.

If we want to select specific variables, we can add them to the `dplyr::rename_with()` argument *.cols*. Here we are going to select columns using the selection helper `where()`. 

* Note: Other `tidyselect` selection helpers include `starts_with()`, `ends_with()`, `contains()`, `matches()`, and `num_range()`, and more.

* Note: You must wrap *is.numeric*, a predicate function (returns a true/false), in `where()`.

```{r}

d1 %>% 
  dplyr::rename_with(~ base::paste0("T1_", .), .cols = where(is.numeric))

```

**3\. Concatenate a prefix, "T1_", to only Var2 and Var3**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Concatenate T1_ by adding the `base::paste0` function.

```{r}

d1 %>% 
  dplyr::rename_with(~ base::paste0("T1_", .), .cols = Var2:Var3)

```

I could also do this by choose all variables *except* Var1.

```{r}

d1 %>% 
  dplyr::rename_with(~ base::paste0("T1_", .), .cols = -Var1)

```

**4\. Concatenate a prefix, "T1_", to Var1 to Var3 and Var5 **

Review the data (d8)

```{r, echo=FALSE}

source("data.R")
d8

```

Concatenate T1_ by adding the `base::paste0` function.

```{r}

d8 %>% 
  dplyr::rename_with(~ base::paste0("T1_", .), .cols = c(Var1:Var3, Var5))

```


### Package: purrr

---

#### Function: `set_names()`

---

* Note: You can use `purrr::set_names()` to concatenate to variable names as well, but it doesn't have the nice compatibility with `tidyselect` that allows you to modify only select variable names. But `purrr::set_names()` works well to modify all variable names.

<br>
  
  **1\. Concatenate a prefix, "T1_", to all variable names**
  
  Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Concatenate T1_ by adding the `stringr::str_c()` function.

* Note: Here we can add the argument *sep = "_"* to add an underscore after T1.


```{r}

d1 %>% 
  purrr::set_names(~ stringr::str_c("T1", sep = "_", .))

```

Return to [Name Variables](https://github.com/Cghlewis/data-wrangling-functions/wiki/Name Variables)
