---
title: "Concatenate to variable names"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: dplyr

---

#### Function: `rename_with()`

---

* Note: Unlike `dplyr::rename()` where you can rename individual columns, `dplyr::rename_with()` renames columns using a function.

* Note: You can either rename all variables or select variables with `dplyr::rename_with()`. The order is, *(function used to rename, columns selected)*. If you don't select columns, all variables will be renamed.

<br>

**1\. Concatenate a prefix, "T1_", to all variable names**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Concatenate T1_ by adding the `base::paste0` function.

* Note: Using `base::paste0` concatenates the string directly to the variable name. Using `base::paste()` concatenates with a string separator.

```{r}

d1 %>% dplyr::rename_with(~ base::paste0("T1_", .))

```

To make a suffix _T1 rather than a prefix, rewrite the function like this.

```{r}

d1 %>% dplyr::rename_with(~ base::paste0(., "_T1"))

```

You can also add your separator if you would rather use the `base::paste()` function

```{r}

d1 %>% dplyr::rename_with(~ base::paste(., "T1", sep= "_"))

```


**2\. Concatenate a prefix, "T1_", to only numeric variable names**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Concatenate T1_ by adding the `base::paste0` function and the selection helper `where()`.

* Note: Other `tidyselect` selection helpers include `starts_with()`, `ends_with()`, `contains()`, `matches()`, and `num_range()`, and more.

```{r}

d1 %>% dplyr::rename_with(~ base::paste0("T1_", .), where(is.numeric))

```

**3\. Concatenate a prefix, "T1_", to only Var2 and Var3**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Concatenate T1_ by adding the `base::paste0` function and the selection helper `where()`.

* Note: Other `tidyselect` selection helpers include `starts_with()`, `ends_with()`, `contains()`, `matches()`, and `num_range()`, and more.

```{r}

d1 %>% dplyr::rename_with(~ base::paste0("T1_", .), Var2:Var3)

```

