---
title: "Concatenate to variable names"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: dplyr

---

#### Function: `dplyr::rename_with()`

---

**1\. Concatenate "T1_" to all variable names**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Concatenate T1_ by adding the `base::paste0` function.

**Note:** Using `base::paste0` concatenates the string directly to the variable name. Using `base::paste()` concatenates with a string separator.

```{r}

d1 %>% dplyr::rename_with(~ base::paste0("T1_", .))

```


**2\. Concatenate "T1_" to only numeric variable names**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Concatenate T1_ by adding the `base::paste0` function and the selection helper `where()`.

**Note:**: Other `tidyselect` selection helpers include `ends_with()`, `contains`, `matches`, and `num_range`, `where` and more.

```{r}

d1 %>% dplyr::rename_with(~ base::paste0("T1_", .), where(is.numeric))

```

