---
title: "Modify variable names"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: dplyr

---

#### Function: `rename_with()`

---

* Note: Unlike `dplyr::rename()` where you can rename individual columns, `dplyr::rename_with()` renames columns using a function.

* Note: You can either rename all variables or select variables. The default argument is *.cols = everything()*.

<br>

**1\. Replace periods in all variable names with underscores**

Review the data (d2)

```{r, echo=FALSE}

source("data.R")
d2

```

Modify variable names by adding the `stringr::str_replace()` function.

* Note: In regular expressions (regex), "." has a special meaning. It is used to match any character. So therefore we will need to escape that functionality, telling the regex to match exactly, by adding "\\\\".

* Note: If there were more than one period in the variable names, you could also use `stringr::str_replace_all()` to remove all instances of periods, rather than just the first instance.

* Note: In this first example, we are creating an anonymous formula using `~` to add the `stringr::str_replace()` arguments of *pattern* and *replacement*.

```{r}

d2 %>% dplyr::rename_with(~ stringr::str_replace(., "\\.", "_"))

```

Instead of creating an anonymous function, we could also "pass the dots" through the *...* argument in `dplyr::rename_with()` and then we do not need to create an anonymous function.

```{r}

d2 %>% dplyr::rename_with(stringr::str_replace, pattern = "\\.", replacement = "_")

```

* Note: Last, if I only wanted to replace the periods for Var1 and Var2, I could add the argument *cols =*

```{r}

d2 %>% dplyr::rename_with(~ stringr::str_replace(., "\\.", "_"), .cols=c(Var.1, Var.2))

```


**2\. Remove 1:, 2: and 3: from Var1, Var2 and Var3**

Review the data (d6)

```{r, echo=FALSE}

source("data.R")
d6

```

Modify variable names by adding the `stringr::str_remove()` function.

```{r}

d6 %>% dplyr::rename_with(~ stringr::str_remove(., "[:digit:]:"))

```


**3\. For any variable that contains the word "var", remove the prefix "s_"**

Review the data (d5)

```{r, echo=FALSE}

source("data.R")
d5

```

Modify variable names by adding the `stringr::str_remove()` function.

* Note: We are using the *.cols* argument here and modifying the default which is *.cols = everything()* and selecting the columns we want by adding the `tidyselect` selection helper `contains()`.

* Note: Other `tidyselect` selection helpers include `startswith()`, `ends_with()`, `matches`, `num_range`, `where` and more.

```{r}

d5 %>% dplyr::rename_with(~ stringr::str_remove(., "s_"), .cols = tidyselect::contains("var"))

```

<br>

### Package: purrr

---

#### Function: `set_names()`

---

* Note: You can use `purrr::set_names()` to modify names as well, but it doesn't have the nice compatibility with `tidyselect` that allows you to modify only select variable names. But `purrr::set_names()` works well to modify all variable names.


<br>

**1\. Remove s_ in all variable names**

Review the data (d5)

```{r, echo=FALSE}

source("data.R")
d5

```

Modify variable names by adding the `stringr::str_remove()` function.

* Note: If there were more than one period in the variable names, you could also use `stringr::str_replace_all()` to remove all instances of s_ in a variable name, rather than just the first instance.

```{r}

d5 %>% purrr::set_names(~ stringr::str_remove(., "s_"))

```


Return to [Name Variables](https://github.com/Cghlewis/data-wrangling-functions/wiki/Name Variables)