---
title: "Modify variable names"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: dplyr

---

#### Function: `dplyr::rename_with()`

---

**1\. Remove all periods in variable names and replace with underscores**

Review the data (d2)

```{r, echo=FALSE}

source("data.R")
d2

```

Modify variable names by adding the `stringr::str_replace()` function.

```{r}

d2 %>% dplyr::rename_with(~ stringr::str_replace(., "\\.", "_"))

```


**2\. For any variable that starts with "s_", remove the prefix**

Review the data (d5)

```{r, echo=FALSE}

source("data.R")
d5

```

Modify variable names by adding the `stringr::str_remove()` function and the selection helper `starts_with()`.

**Note:**: Other `tidyselect` selection helpers include `ends_with()`, `contains`, `matches`, and `num_range`, `where` and more.

```{r}

d5 %>% dplyr::rename_with(~ stringr::str_remove(., "s_"), starts_with("s_"))

```

