---
title: "Set variable names using a data dictionary"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: dplyr

---

#### Function: `rename()`

---

* Note: The `dplyr::rename()` formula is *new value=old value*, this is opposite of `dplyr::recode()`.

<br>

**1\. Set all column names using an existing file (for example a data dictionary).**

Review the data (d1)

```{r echo = FALSE}

source("data.R")
d1

```

Read in and review the data dictionary

* Note: The order of your data dictionary data frame must be new_names, old_names. If your variables are not already in that order, you can use `dplyr::relocate()` to reorder them.


```{r, eval = FALSE}

dict <- readxl::read_excel("dictionary.xlsx")

```

```{r, echo = FALSE}

source("data.R")
dict

```

In this case we can use the existing file to rename the variables for us, rather than us hand entering "stu_id = Var1, read_score = Var2, math_score = Var3". You can see how this would save us time if we have many variables to rename.


The first thing we need to do is use the function `tibble::deframe()` to convert our dictionary data frame to a two-column names vector.


```{r}

dict_names <- tibble::deframe(dict)

dict_names

```

Now we can rename our variables using this vector.

* Note: We need to add `tidyselect::all_of()` to remove ambiguity between columns and external variables. See this link for more details https://tidyselect.r-lib.org/reference/faq-external-vector.html

```{r}

d1 %>%
 dplyr::rename(tidyselect::all_of(dict_names))

```

There are many other ways to rename variables using another column of names. However I like this way the best because of versatility.

Even if the data dictionary rows are not in the same order as the variables in our data frame, the renaming will still work.

For example, here is the dictionary where the rows are in a different order.

```{r}

dict2

```

And we can still use this to rename our variables.


```{r}

dict_names <- tibble::deframe(dict2)

dict_names

```


```{r}

d1 %>%
 dplyr::rename(tidyselect::all_of(dict_names))

```

We can also use this dictionary, even if it doesn't have new names for all of our variables. Take this dictionary that can only relabel 2 of our variables.

```{r}

dict3

```

We can still rename.


```{r}

dict_names <- tibble::deframe(dict3)

dict_names

```


```{r}

d1 %>%
 dplyr::rename(tidyselect::all_of(dict_names))

```

Return to [Name Variables](https://github.com/Cghlewis/data-wrangling-functions/wiki/name-variables)