---
title: "Rename variables"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: dplyr

---

#### Function: `rename()`

---

* Note: The `dplyr::rename()` formula is *new value=old value*, this is opposite of `dplyr::recode()`.

<br>

**1\. Change only the second variable name (Var2)**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Change the second variable name (Var2)

* Note: Quotes are not required around variable names.

```{r}

d1 %>% 
  dplyr::rename(grade = Var2)

```

If you wanted to change names of more than one variable you can just keep adding more variables.

```{r}

d1 %>% 
  dplyr::rename(grade = Var2, score = Var3)

```

<br>

### Package: dplyr

---

#### Function: `select()`

---

**1\. Rename variables when selecting them**

Review the data (d9)

```{r, echo=FALSE}

source("data.R")
d9

```


Select "school name" and "Var1" and rename them in the select statement.

* Note: The formula is the same as `dplyr::rename()`, *new value=old value*

```{r}

d9 %>%
  dplyr::select(school_name = `school name`, enrollment = `Var1`)

```

Return to [Name Variables](https://github.com/Cghlewis/data-wrangling-functions/wiki/Name Variables)