---
title: "Create a single race variable from dummy variables"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: dplyr

---

#### Function: `mutate()`

---

**1\. Create a single race variable from a series of dummy coded variables**

Review the data (d11)

```{r, echo = FALSE}

source("data.R")
d11

```

We want a single race variable where each value is represented if it is the only one selected, but if more than one race is selected, the value is set to 7 (to represent multi-racial).

We will use `dplyr::case_when()` to draft the logic, along with `dplyr::if_all()` to check values across multiple variables.

```{r}

d11 %>%
  dplyr::mutate(
    race = 
      dplyr::case_when(
        race1 == 1 & dplyr::if_all(race2:race6, ~. == 0) ~ 1,
        race2 == 1 & dplyr::if_all(c(race1, race3:race6), ~. == 0) ~ 2,
        race3 == 1 & dplyr::if_all(c(race1:race2, race4:race6), ~. == 0) ~ 3,
        race4 == 1 & dplyr::if_all(c(race1:race3, race5:race6), ~. == 0) ~ 4,
        race5 == 1 & dplyr::if_all(c(race1:race4, race6), ~. == 0) ~ 5,
        race6 == 1 & dplyr::if_all(race1:race5, ~. == 0) ~ 6,
        rowSums(across(race1:race6)) > 1 ~ 7))

```

While the code above is definitely the more explicit way to assign values, you could also simplify the above code by putting the logic for the multi-racial value at the top because `dplyr::case_when()` reviews logic in order and does not replace previously assigned values.


```{r}

d11 %>%
  dplyr::mutate(
    race = 
      dplyr::case_when(
        rowSums(across(race1:race6)) > 1 ~ 7,
        race1 == 1 ~ 1,
        race2 == 1 ~ 2,
        race3 == 1 ~ 3,
        race4 == 1 ~ 4,
        race5 == 1 ~ 5,
        race6 == 1 ~ 6,
))

```

**Note: You could of course use this same logic for other variables, not just race. For example, dummy coded grade levels that teachers work with (assigning a new value when teachers work with multiple grade levels).

Return to [Create New Variables](https://github.com/Cghlewis/data-wrangling-functions/wiki/Create New Variables)
