---
title: "Add dummy coded columns"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: dplyr

---

#### Function: `mutate()`

---


**1\. Create new dummy coded variables from a categorical column with multiple items selected per cell (`cat`)**

Review the data (d2)

```{r, echo = FALSE}

source("data.R")
d2

```

We see here that our `cat` variable was a select all that apply question (where options are separated by a comma). We want these options to separated out into their own dummy coded variables that are 1 if the selection was chosen and 0 if the selection was not chosen.

We can create our new variables using `dplyr::mutate()`, adding all variables in our one `mutate` function.

cat1 = 1  
cat2 = 2  
cat4 = 4  
cat5 = 5  
cat10 = 10  
cat11 = 11  

* Note: I used `stringr::str_count()` to count the number of instances for each categorical variable (there should only be one per row). We used regular expressions to pull out the values of interest.

```{r}

d2 %>%
  dplyr::mutate(cat1 = stringr::str_count(cat, " 1,|^1,|, 1$"),
                cat2 = stringr::str_count(cat, "2"),
                cat4 = stringr::str_count(cat, "4"),
                cat5 = stringr::str_count(cat, "5"),
                cat10 = stringr::str_count(cat, "^10| 10,|, 10$"),
                cat11 = stringr::str_count(cat, "^11| 11,|, 11$"))

```

We could have also used `stringr::str_detect()` if we were worried a category could duplicate in each row. 

* Note: With `dplyr::case_when()` we need to specifically recode NA values back to NA, otherwise they will be coded to 0.

```{r}

d2 %>%
  dplyr::mutate(cat1 = 
                  dplyr::case_when(
                    stringr::str_detect(cat, " 1,|^1,|, 1$") ~ 1,
                    is.na(cat) ~ NA_real_,
                    TRUE ~ 0))

```

Or `dplyr::if_else()`

```{r}

d2 %>%
  dplyr::mutate(cat1 = dplyr::if_else(stringr::str_detect(cat, " 1,|^1,|, 1$"), 1, 0))

```


**2\. Create new dummy coded race variables from a race variable with multiple items selected per cell**

Review the data (d12)

```{r, echo = FALSE}

source("data.R")
d12

```

Here we want to create one dummy variable per race option BUT when "99" is selected (prefer not to answer), all values for race dummy variables should be 99.

While I could write out logic to create all 6 variables, I could instead create a function which would remove some of the repetitiveness.

```{r}

# Create function

race_vars <- function(Var, Num) {
  x <- case_when(
    is.na(Var) ~ NA,
    Var == 99 ~ 99,
    str_detect(Var, Num) ~ 1, 
    .default = 0
  )
}

# Use function to create race variables

d12 %>% 
  mutate(race1 = 
           race_vars(race, "1"),
         race2 =
           race_vars(race, "2"),
         race3 = 
           race_vars(race, "3"),
         race4 =
           race_vars(race, "4"),
         race5 =
           race_vars(race, "5"),
         race6 = 
           race_vars(race, "6"))


```

<br>

### Package: tidyr

---

#### Function: `pivot_wider()`

---

**1\. Create new dummy coded variables from a string categorical column (`cert`)**

Review the data (d3)

```{r, echo = FALSE}

source("data.R")
d3

```

In this situation, we only have one word per cell selected.

Using `tidyr::pivot_wider()` we can pivot our categorical variable into separate dummy variables where 1 = yes the category was selected and 0 = the category was not selected.

Before doing this, we will need to add one new count variable using the `dplyr::mutate()` function. This count variable will be used in our *values_from* argument in for `tidyr::pivot_wider()`.

* Note: We also add the argument *values_fill* to fill in any cell where the response option was not selected. 

* Note: I added `janitor::clean_names()` at the end to replace the space within variable names with an underscore.

```{r}

d3 %>%
  dplyr::mutate(count = 1) %>%
  tidyr::pivot_wider(names_from = cert,
              values_from = count,
              values_fill = 0) %>%
  janitor::clean_names()

```

There are many other ways to do this using `dplyr::if_else()` or `dplyr::case_when()`. Here is an example below. However, I think using `tidyr::pivot_wider()` is a quicker and more efficient option.

```{r}

d3 %>%
  dplyr::mutate(elem_ed = dplyr::if_else(cert == "elem ed", 1, 0))

```

**2\. Create new dummy coded variables from a categorical column with multiple items selected per cell (`cert`)**

Review the data (d7)

```{r, echo = FALSE}

source("data.R")
d7

```

Now we have a select all question where multiple responses may exist in the cert variable. This is okay, we can still use `dplyr::pivot_wider()` here.

First we need to separate our variable into multiple rows. We can do this using `tidyr::separate_rows()` like we did in the [separate](https://cghlewis.github.io/data-wrangling-functions/separate/split-rows.html) section. When we do the separation, I save the output over the original "cert" variable. I could have also created a new variable if I wanted to.

Once we do that our cases will repeat with each response to the `cert` variable appearing on a new line. This may be scary at first but we will clean it up in just a second.

Next we can create a count variable like we did above using `dplyr::mutate()`. We can call this variable whatever we want. Here I am calling it `count`.

Now we can pivot. After we pivot, we notice that our data has been regrouped to one line per case again. No more duplicate rows. And our columns such as age are still in our dataset, which is what we wanted.

```{r}

d7 %>% 
  tidyr::separate_rows(cert, sep = ", ") %>%
  dplyr::mutate(count = 1) %>%
  tidyr::pivot_wider(
    names_from = cert,
    values_from = count,
    values_fill = 0
  ) %>%
  janitor::clean_names()

```

If we wanted to more explicitly keep our age and id variables, we could name them in the *id_cols* argument to make sure they carry over.

```{r}

d7 %>% 
  tidyr::separate_rows(cert, sep = ", ") %>%
  dplyr::mutate(count = 1) %>%
  tidyr::pivot_wider(
    id_cols = c(id, age),
    names_from = cert,
    values_from = count,
    values_fill = 0
  ) %>%
  janitor::clean_names()

```

**3\. Create new dummy coded variables from a categorical column with multiple items selected per cell (cat)**

Review the data (d2)

```{r, echo = FALSE}

source("data.R")
d2

```

This is the same example from #1 at the top of this page. However, now we can create our dummy coded variables using `tidy::pivot_wider()` saving us so much time compared to the `stringr::str_count()` or `stringr::str_detect()` method.

The only thing I am adding here, compared to the example directly above, is the *names_prefix* argument.

I am also adding a `dplyr::select()` statement at the end to reorder the variables and drop the "catNA" variable that is created due to id 14 having a value of NA for "cat".

```{r}

d2 %>% 
  tidyr::separate_rows(cat, sep = ", ") %>%
  dplyr::mutate(count = 1) %>%
  tidyr::pivot_wider(
    names_from = cat,
    names_prefix = "cat",
    values_from = count, 
    values_fill = 0
  ) %>%
  dplyr::select(id, cat1, cat2, cat4, cat5, cat10, cat11)

```

Last, you'll notice above that id 14 now has zeros for all of the categories. Whereas before they had NAs. If the NAs actually mean that their data is missing, then we don't want id 14 to have zeros for all of our dummy categories. We want them to have NAs.

We could fix this by keeping the "catNA" variable and then using this variable as an indicator for when to recode our other categorical variables back to NA. 

```{r}

d2 %>% 
  tidyr::separate_rows(cat, sep = ", ") %>%
  dplyr::mutate(count = 1) %>%
  tidyr::pivot_wider(
    names_from = cat,
    names_prefix = "cat",
    values_from = count, 
    values_fill = 0
  ) %>%
  dplyr::select(id, catNA, cat1, cat2, cat4, cat5, cat10, cat11) %>%
  dplyr::mutate(dplyr::across(cat1:cat11,
                       ~ dplyr::case_when(
                         catNA == 1 ~ NA_real_,
                         TRUE ~ .
                       ))) %>%
  dplyr::select(-catNA)

```


Return to [Create New Variables](https://github.com/Cghlewis/data-wrangling-functions/wiki/Create New Variables)
