---
title: "Add hand inputted columns"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: tibble

---

#### Function: `add_column()`

---


**1\. Add a new column (`year`) to the data**

Review the data (d1)

```{r, echo = FALSE}

source("data.R")
d1

```

Add participation year equal to "2018-19" for all IDs.

```{r}

d1 %>%
  tibble::add_column(year = "2018-19")

```

I can also change the location of the new column by adding the argument *.before* or *.after*.

```{r}

d1 %>%
  tibble::add_column(year = "2018-19", .after = "id")

```

<br>

### Package: dplyr

---

#### Function: `mutate()`

---

**1\. Add a new column (`year`) to the data**

Review the data (d1)

```{r, echo = FALSE}

source("data.R")
d1

```

I can use `dplyr::mutate()` to create our new variable "year" and assign it a value that is applied to every row. In this case I want to assign a value of 1, for the first year of the study. I can also add the argument *.before* or *.after* to change the location of the new variable.

```{r}

d1 %>%
  dplyr::mutate(year = 1, .after = id)

```



Return to [Create New Variables](https://github.com/Cghlewis/data-wrangling-functions/wiki/Create New Variables)
