---
title: "Add a consecutive ID variable"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: dplyr

---

#### Function: `consecutive_id()`

---


**1\. Assign an ID for each combination of two variables**

Review the data (d10)

```{r, echo = FALSE}

source("data.R")
d10

```

I want to create a new variable `video_id` that assigns an ID value for each unique date within each `tch_id` group. I also want those numbers to be order chronologically.

In order to do this, 

1. I first group by `tch_id` (using `dplyr::group_by()`)  
2. Then order dates chronologically (using `dplyr::arrange()`)  
3. Then create our new variable using `dplyr::mutate()` and use `dplyr::consecutive_id()` to create a unique identifier that increments every time a variable changes  
4. Last I make sure to `dplyr::ungroup()` our data

```{r}

d10 %>%
  group_by(tch_id) %>%
  arrange(tch_id, date_observed) %>%
  mutate(video_id = consecutive_id(date_observed)) %>%
  ungroup()

```

You could also do this without `dplyr::consecutive_id()` by using `base::as.factor()` instead.

```{r eval = FALSE}

d10 %>%
  group_by(tch_id) %>%
  arrange(tch_id, date_observed) %>%
  mutate(video_id = as.numeric(as.factor(date_observed))) %>%
  ungroup()

```


Return to [Create New Variables](https://github.com/Cghlewis/data-wrangling-functions/wiki/Create New Variables)
