---
title: "Add recoded score columns"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: dplyr

---

#### Function: `mutate()`

---

**1\. Create new scored variables for all of the "item" variables**

Review the data (d1)

```{r, echo = FALSE}

source("data.R")
d1

```

`item1` through `item4` were all multiple choice variables with one correct answer. We now want to score these variables to be correct (1) or incorrect (0).

Item1: correct answer = 3  
Item2: correct answer = 5  
Item3: correct answer = 3  
Item4: correct answer = 5  

Since `item1` and `item3` have the same correct answers, and `item2` and `item4` have the same correct answers, we can use `dplyr::across()` within `dplyr::mutate()` to score these variables at the same time. We use `.` to denote we want the criteria to apply to the variables we have selected.

* Note: We use the `dplyr::across()` argument *.names* to rename our newly scored variables. You can use {col} to stand for the selected column name.

* Note: We have NA values in our variables, so since `dplyr::case_when()` recodes in order of how statements are provided, we will first want to say we want NA values to stay as NA values (NA_real_).

Note: In the `dplyr::case_when()` function, TRUE means if a value was not evaluated in my arguments above, replace with the value I give. In this case, I am saying replace remaining values with 0.

* Note: We are adding the `dplyr::across()` argument *.names* to rename our newly created variables with the suffix "_scored".

```{r}


d1 %>%
  dplyr::mutate(
    dplyr::across(
      c(item1, item3),
      ~ dplyr::case_when(is.na(.) ~ NA_real_,
                         . == 3 ~ 1,
                         TRUE ~ 0),
      .names = "{col}_scored"
    ),
    dplyr::across(
      c(item2, item4),
      ~ dplyr::case_when(is.na(.) ~ NA_real_,
                         . == 5 ~ 1,
                         TRUE ~ 0),
      .names = "{col}_scored"
    )
  ) %>%
  dplyr::select(id, tidyselect::contains("scored"))

```

We could also write the formula this way, with explicitly listing the values to be recoded to 0 and recoding NA at the end, but this is a bit longer.

```{r}

d1 %>%
  dplyr::mutate(
    dplyr::across(
      c(item1, item3),
      ~ dplyr::case_when(. == 3 ~ 1,
                         . %in% c(1, 2, 4, 5) ~ 0,
                         TRUE ~ .),
      .names = "{col}_scored"
    ),
    dplyr::across(
      c(item2, item4),
      ~ dplyr::case_when(. == 5 ~ 1,
                         . %in% c(1, 2, 3, 4) ~ 0,
                         TRUE ~ .),
      .names = "{col}_scored"
    )
  ) %>%
  dplyr::select(id, tidyselect::contains("scored"))

```


Return to [Create New Variables](https://github.com/Cghlewis/data-wrangling-functions/wiki/Create New Variables)
