---
title: "Add dummy coded columns"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: dplyr

---

#### Function: `mutate()`

---


**1\. Create new dummy coded variables from a categorical column with multiple items selected per cell (cat)**

Review the data (d2)

```{r, echo = FALSE}

source("data.R")
d2

```

We see here that our "cat" variable was a select all that apply question (where options are separated by a comma). We want these options to separated out into their own dummy coded variables that are 1 if the selection was chosen and 0 if the selection was not chosen.

We can create our new variables using `dplyr::mutate()`, adding all variables in our one `mutate` function.

cat1 = 1  
cat2 = 2  
cat4 = 4  
cat5 = 5  
cat10 = 10  
cat11 = 11  

* Note: I used `stringr::str_count()` to count the number of instances for each categorical variable (there should only be one per row). We used regular expressions to pull out the values of interest.

```{r}

d2 %>%
  dplyr::mutate(cat1 = stringr::str_count(cat, " 1,|^1,|, 1$"),
                cat2 = stringr::str_count(cat, "2"),
                cat4 = stringr::str_count(cat, "4"),
                cat5 = stringr::str_count(cat, "5"),
                cat10 = stringr::str_count(cat, "^10| 10,|, 10$"),
                cat11 = stringr::str_count(cat, "^11| 11,|, 11$"))

```

We could have also used `stringr::str_detect()` if we were worried a category could duplicate in each row. 

* Note: With `dplyr::case_when()` we need to specifically recode NA values back to NA, otherwise they will be coded to 0.

```{r}

d2 %>%
  dplyr::mutate(cat1 = 
                  dplyr::case_when(
                    stringr::str_detect(cat, " 1,|^1,|, 1$") ~ 1,
                    is.na(cat) ~ NA_real_,
                    TRUE ~ 0))

```

Or `dplyr::if_else()`

```{r}

d2 %>%
  dplyr::mutate(cat1 = dplyr::if_else(stringr::str_detect(cat, " 1,|^1,|, 1$"), 1, 0))

```

<br>

### Package: tidyr

---

#### Function: `pivot_wider()`

---

**1\. Create new dummy coded variables from a string categorical column (cert)**

Review the data (d3)

```{r, echo = FALSE}

source("data.R")
d3

```

In this situation, we only have one word per cell selected.

Using `tidyr::pivot_wider()` we can pivot our categorical variable into separate dummy variables where 1 = yes the category was selected and 0 = the category was not selected.

Before doing this, we will need to add one new count variable using the `dplyr::mutate()` function. This count variable will be used in our *values_from* argument in for `tidyr::pivot_wider()`.

* Note: We also add the argument *values_fill* to fill in any cell where the response option was not selected. 

```{r}

d3 %>%
  dplyr::mutate(count = 1) %>%
  tidyr::pivot_wider(names_from = cert,
              values_from = count,
              values_fill = 0) %>%
  janitor::clean_names()

```

There are many other ways to do this using `dplyr::if_else()` or `dplyr::case_when()`. Here is an example below. However, I think using `tidyr::pivot_wider()` is a quicker and more efficient option.

```{r}

d3 %>%
  dplyr::mutate(elem_ed = dplyr::if_else(cert == "elem ed", 1, 0))

```

Return to [Create New Variables](https://github.com/Cghlewis/data-wrangling-functions/wiki/Create New Variables)
