---
title: "Add recoded columns"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: dplyr

---

#### Function: `mutate()`

---

**1\. Create a new scored variable for item1**

Review the data (d1)

```{r, echo = FALSE}

source("data.R")
d1

```

item1 through item4 were all multiple choice variables with one correct answer. We now want to score these variables to be correct (1) or incorrect (0).

Item1: correct answer = 3  
Item2: correct answer = 5  
Item3: correct answer = 3  
Item4: correct answer = 5  

* Note: Here we use `dplyr::mutate()` to create a new variable, named item1_scored. We could have also chosen to name it item1 and write over the original variable.

* Note: As we learned from the [Recode](https://github.com/Cghlewis/data-wrangling-functions/wiki/Recode) section, we can recode this variable in many ways including `dplyr::recode()`, `dplyr::if_else()` or `dplyr::case_when()`. In this case we use `dplyr::case_when()`.

```{r}

d1 %>%
  dplyr::mutate(item1_scored = 
                  dplyr::case_when(
                    item1 == 3 ~ 1,
                    TRUE ~ 0))

```


**2\. Create new scored variables for _item_ variables**

Review the data (d1)

```{r, echo = FALSE}

source("data.R")
d1

```

Since Item 1 and 3 and Items 2 and 4 have the same correct answers, we can use `dplyr::across()` within `dplyr::mutate()` to score these variables at the same time. We use `.` to denote we want the criteria to apply to the variables we have selected.

* Note: We use the `dplyr::across()` argument *.names* to rename our newly scored variables. You can use {.col} to stand for the selected column name.

* Note: We have NA values in our variables, so since `dplyr::case_when()` recodes in order of how statements are provided, we will first want to say we want NA values to stay as NA values (NA_real_).

```{r}

d1 %>%
  dplyr::mutate(dplyr::across(c(item1, item3),
                ~dplyr::case_when(
                  is.na(.) ~ NA_real_,
                  . == 3 ~ 1,
                  TRUE ~ 0),
                .names = "{col}_scored"),
                dplyr::across(c(item2, item4),
                ~dplyr::case_when(
                  is.na(.) ~ NA_real_,
                  . == 5 ~ 1,
                  TRUE ~ 0),
                .names = "{col}_scored")) %>%
  dplyr::select(id, tidyselect::contains("scored"))

```

We could also write the formula this way, with explicitly listing the values to be recoded to 0 and recoding NA at the end, but this is much longer.

```{r}

d1 %>%
  dplyr::mutate(dplyr::across(c(item1, item3),
                ~dplyr::case_when(
                  . == 3 ~ 1,
                  . %in% c(1,2,4,5) ~ 0,
                  TRUE ~ .),
                .names = "{col}_scored"),
                dplyr::across(c(item2, item4),
                ~dplyr::case_when(
                  . == 5 ~ 1,
                  . %in% c(1,2,3,4) ~ 0,
                  TRUE ~ .),
                .names = "{col}_scored")) %>%
  dplyr::select(id, tidyselect::contains("scored"))

```

**3\. Create new dummy coded variables from a categorical column (cat)**

Review the data (d2)

```{r, echo = FALSE}

source("data.R")
d2

```

We see here that our "cat" variable was a select all that apply question (where options are separated by a comma). We want these options to separated out into their own dummy coded variables that are 1 if the selection was chosen and 0 if the selection was not chosen.

We can create our new variables using `dplyr::mutate()`, adding all variables in our one `mutate` function.

cat1 = 1
cat2 = 2
cat4 = 4
cat5 = 5
cat10 = 10
cat11 = 11

* Note: I used `stringr::str_count()` to count the number of instances for each categorical variable (there should only be one per row). We used regular expressions to pull out the values of interest.

```{r}

d2 %>%
  dplyr::mutate(cat1 = stringr::str_count(cat, " 1,|^1,|, 1$"),
                cat2 = stringr::str_count(cat, "2"),
                cat4 = stringr::str_count(cat, "4"),
                cat5 = stringr::str_count(cat, "5"),
                cat10 = stringr::str_count(cat, "^10| 10,|, 10$"),
                cat11 = stringr::str_count(cat, "^11| 11,|, 11$"))

```

We could have also used `stringr::str_detect()` if we were worried a category could duplicate in each row. 

* Note: With `dplyr::case_when()` we need to specifically recode NA values back to NA, otherwise they will be coded to 0.

```{r}

d2 %>%
  dplyr::mutate(cat1 = 
                  case_when(
                    stringr::str_detect(cat, " 1,|^1,|, 1$") ~ 1,
                    is.na(cat) ~ NA_real_,
                    TRUE ~ 0))

```

Or `dplyr::if_else()`

```{r}

d2 %>%
  dplyr::mutate(cat1 = dplyr::if_else(stringr::str_detect(cat, " 1,|^1,|, 1$"), 1, 0))

```


Return to [Create New Variables](https://github.com/Cghlewis/data-wrangling-functions/wiki/Create New Variables)
