---
title: "Add recoded score column"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: dplyr

---

#### Function: `mutate()`

---

**1\. Create a new scored variable for item1**

Review the data (d1)

```{r, echo = FALSE}

source("data.R")
d1

```

item1 through item4 were all multiple choice variables with one correct answer. We now want to score these variables to be correct (1) or incorrect (0).

Item1: correct answer = 3  
Item2: correct answer = 5  
Item3: correct answer = 3  
Item4: correct answer = 5  

* Note: Here we use `dplyr::mutate()` to create a new variable, named item1_scored. We could have also chosen to name it item1 and write over the original variable.

* Note: As we learned from the [Recode](https://github.com/Cghlewis/data-wrangling-functions/wiki/Recode) section, we can recode this variable in many ways including `dplyr::recode()`, `dplyr::if_else()` or `dplyr::case_when()`. In this case we use `dplyr::case_when()`.

```{r}

d1 %>%
  dplyr::mutate(item1_scored = 
                  dplyr::case_when(
                    item1 == 3 ~ 1,
                    TRUE ~ 0))

```


Return to [Create New Variables](https://github.com/Cghlewis/data-wrangling-functions/wiki/Create New Variables)
