---
title: "Add a recoded column"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: dplyr

---

#### Function: `mutate()`

---

**1\. Create a new scored variable for item1**

Review the data (d1)

```{r, echo = FALSE}

source("data.R")
d1

```

Item1 through Item4 were all multiple choice variables with one correct answer. We now want to score these variables to be correct or incorrect.

Item1: correct answer = 3
Item2: correct answer = 5
Item3: correct answer = 3
Item4: correct answer = 5

* Note: Here we use `dplyr::mutate()` to create a new variable, named item1_scored. We could have also chosen to name it item1 and write over the original variable.

* Note: As we learned from the [Recode](https://github.com/Cghlewis/data-wrangling-functions/wiki/Recode) section, we can recode this variable in many ways including `dplyr::recode()`, `dplyr::if_else()` or `dplyr::case_when()`. In this case we use `dplyr::if_else()`.

```{r}

d1 %>%
  dplyr::mutate(item1_scored = 
                  dplyr::case_when(
                    item1 == 3 ~ 1,
                    TRUE ~ item1))

```


**2\. Create new scored variables for _item_ variables**

Review the data (d1)

```{r, echo = FALSE}

source("data.R")
d1

```

Since Item 1 and 3 and Items 2 and 4 have the same correct answers, we can use `dplyr::across()` to score these variables at the same time. We use `.` to denote we want the criteria to apply to the variables we have selected.

* Note: We use the `dplyr::across()` argument *.names* to rename our newly scored variables. You can use {.col} to stand for the selected column name.

```{r}

d1 %>%
  dplyr::mutate(dplyr::across(c(item1, item3),
                ~dplyr::case_when(
                  . == 3 ~ 1,
                  TRUE ~ .),
                .names = "{col}_scored"),
                dplyr::across(c(item2, item4),
                ~dplyr::case_when(
                  . == 5 ~ 1,
                  TRUE ~ .),
                .names = "{col}_scored"))

```

Return to [Create New Variables](https://github.com/Cghlewis/data-wrangling-functions/wiki/Create New Variables)

Return to [XXXX](https://github.com/Cghlewis/data-wrangling-functions/wiki/Complete)