---
title: "Add new indicator column"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: dplyr

---

#### Function: `mutate()`

---

**1\. Create an indicator column (`scale1_risk`)**

```{r, echo = FALSE}

source("data.R")
d9

```

I have a sum score and a cut score for a scale. The cut scores are different depending on grade level. 

I want to create an indicator of risk level:

sum score < cut score = 0
score score >= cut score = 1 

* Note: I use `dplyr::case_when()` to recode my sum variable into a dichotomous variable.

My logic says:  

- If either `grade` or my selected sum score variable is NA, then my new variable should be coded to NA.  
- Then if my selected sum score variable is less than my selected cut score variable, code my new variable to 0.  
- Else (`TRUE`) code to 1.

```{r}

d9 %>%
  dplyr::mutate(scale1_risk =
                  dplyr::case_when(
    is.na(grade) | is.na(scale1_sum) ~ NA_real_,
    scale1_sum < cut_scale1 ~ 0,
    TRUE ~ 1
  ))

```


**2\. Create multiple `_risk` indicator columns**

Review the data (d6)

```{r, echo = FALSE}

source("data.R")
d6

```

This time I have sum scores for 3 different scales. I also have 3 different cut scores for each of those scales. The cut scores are different depending on grade level. 

I want to create an indicator of risk level:

sum score < cut score = 0
score score >= cut score = 1 

I could write this calculation out one piece at a time, but I would prefer to create a function that I can use that will minimize duplication.

First I create a cutoff function that I can use.

In this function I call my data frame and I use `dplyr::mutate()` to create my new `at_risk` variable which will be named dynamically. 

* Note: In order to name this variable dynamically I have to use `:=`. I also have to surround all data frame variables with `{{}}` to denote that I am referring to data frame variables, not environment variables.


```{r}

cutoff <- function(df, new_name, cut_score, sum_score) {
  df %>%
    dplyr::mutate({{new_name}} := dplyr::case_when(
      is.na(grade) | is.na({{sum_score}}) ~ NA_real_,
      {{sum_score}} < {{cut_score}} ~ 0,
      TRUE ~ 1))
}

```

Now I can use my new function to create my new `_risk` variables.

```{r}


d6 <- cutoff(d6, scale1_risk, cut_scale1, scale1_sum)

d6 <- cutoff(d6, scale2_risk, cut_scale2, scale2_sum)

d6 <- cutoff(d6, scale3_risk, cut_scale3, scale3_sum)

d6 %>%
  dplyr::select(tidyselect::contains("risk"))

```



Return to [Create New Variables](https://github.com/Cghlewis/data-wrangling-functions/wiki/Create New Variables)
