---
title: "Combine strings in long format"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: stringr

---

#### Function: `str_c()`

---

**1\. Combine subtests based on id and status**

Review the data (d8)

```{r, echo=FALSE}

source("data.R")

d8

```

To combine subtests in long format, I first need to group the data by both "id" and "status" using `dplyr::group_by()`.

Then I can create a new variable "subtest" which will be my combined subtests. I can use my `stringr::str_c()` function to combine or concatenate the subtests. I want subtests to be separated by a "," so I will add the argument *collapse = ","*

```{r}

d8_new <- d8 %>%
  dplyr::group_by(id, status) %>%
  dplyr::summarize(subtest = stringr::str_c(subtest, collapse=","))

d8_new

```

If I wanted all of a user's information on the same row, I could take this one step further and use `tidyr::pivot_wider()` to pivot the data to wide format.

```{r}

d8_new %>%
  tidyr::pivot_wider(
  names_from = status,
  values_from = subtest
  )

```


Return to [Strings](https://github.com/Cghlewis/data-wrangling-functions/wiki/Strings)
