---
title: "Combine strings in wide format"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: stringr

---

#### Function: `str_c()`

---

**1\. Combine two variables (`school` and `district`) with a separator (`_`)**

Review the data (d9)

```{r, echo=FALSE}

source("data.R")

d9

```

I will first use `dplyr::mutate()` to create my new variable which I am calling `sch_dist`.

I can then use my `stringr::str_c()` function to combine or concatenate the two variables which I want to separate by a "," so I will add the argument *collapse = ","*

```{r}

d9 %>%
  dplyr::mutate(sch_dist = stringr::str_c(school, district, sep = ","))

```

**2\. Combine two variables (`item1` and `measure`) with no separator**

Review the data (d10)

```{r, echo=FALSE}

source("data.R")

d10

```

Again, I will use `dplyr::mutate()` first to create my new variable `item1_measure`.

Then I can use my `stringr` function to combine values.

In this case I don't want any separator between my two values so I will write my separator argument as *sep = ""*.

```{r}

d10 %>%
  dplyr::mutate(item1_measure = stringr::str_c(item1, measure, sep = ""))

```

* Note: This is also equivalent to using `base::paste0()` which unlike `base::paste()` has no separator.

```{r}

d10 %>%
  dplyr::mutate(item1_measure = base::paste0(item1, measure))

```

<br>

### Package: tidyr

---

#### Function: `unite()`

---

**1\. Combine multiple dummy variables into one variable**

Review the data (d16)

```{r, echo=FALSE}

source("data.R")

d16

```

Combine responses to the question "Which pizza toppings do you like?"

The problem with `stringr::str_c()` and `base::paste0()` is that they don't handle NA values well. Since this was a select all question, answers were provided when selected, and NA otherwise.

```{r}

d16 %>%
  mutate(toppings = str_c(cheese, pepperoni, mushrooms, sep = ","))

```

```{r}

d16 %>%
  mutate(toppings = paste0(cheese, pepperoni, mushrooms, collapse = ","))

```

A great alternative is to use `tidyr::unite()`. Here we can use the argument *na.rm = TRUE*

*Note: `dplyr::mutate()` is not needed here

```{r}

d16 %>%
  unite(col = toppings, cheese, pepperoni, mushrooms, sep = ",", na.rm = TRUE)

```


Return to [Strings](https://github.com/Cghlewis/data-wrangling-functions/wiki/Strings)
