---
title: "Count values in strings"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: stringr

---

#### Function: `str_length()`

---

**1\. Count the length of item1**

Review the data (d11)

```{r, echo=FALSE}

source("data.R")

d11

```

In this data, "item1" represents student responses of a word. Each letter represents a correct phonetic sound. We want to count the number of correct sounds for each student.

We first create a new variable "item1_score" using `dplyr::mutate()`. We can then use `stringr::str_length()` to count the length of our strings.

```{r}

d11 %>%
  dplyr::mutate(item1_score = stringr::str_length(item1))

```

<br>

---

#### Function: `str_count()`

---

**1\. Count the length of item1**

Review the data (d12)

```{r, echo=FALSE}

source("data.R")

d12

```

Same as above, "item1" represents student responses of a word. Each letter represents a correct phonetic sound. We want to count the number of correct sounds for each student. However, this time, our data has commas between letters so we cannot use `stringr::str_length()` because it will count the commas, giving us incorrect scores.

Instead we use `stringr::str_count()` and add the argument *pattern =* and we add the regex pattern "[a-z]" which says to count any lowercase letter a to z.

```{r}

d12 %>%
  dplyr::mutate(item1_score = stringr::str_count(item1, pattern = "[a-z]"))

```



Return to [Strings](https://github.com/Cghlewis/data-wrangling-functions/wiki/Strings)
