---
title: "Extract values and combine"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: stringr

---

#### Function: `str_extract()`

---

**1\. Extract specified phrases from a list of tags in a variable**

Review the data (d17)

```{r, echo=FALSE}

source("data.R")

d17

```

We have a variable with several tags included but we want to create a new variable which only includes tags from a specified list. First we create a vector of the tags of interest.

```{r}

tags_of_interest <- c("long tag 1", "other long tag 2", "long tag 6")

```

Now we want to create our new variable, `special_tags` that pulls out any tag from that list in our current   `tag` variable and places it in our new variable.

* Note: Because we are adding `dplyr::rowwise()` we will need to `ungroup()` at the end so that that it doesn't slow down our future operations.


```{r}

d17 %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    special_tags =
      stringr::str_extract(tags, tags_of_interest) %>%
      na.omit() %>%
      paste0(collapse = "; ")
  ) %>%
  dplyr::ungroup()

```


Return to [Strings](https://github.com/Cghlewis/data-wrangling-functions/wiki/Strings)
