---
title: "Extract values based on pattern"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: stringr

---

#### Function: `str_extract()`

---

**1\. Extract all numeric values from `item1`**

Review the data (d3)

```{r, echo=FALSE}

source("data.R")

d3

```

We first need to create our new variable `item1` which will write over our existing `item1` variable since I named it the same name. We do this using `dplyr::mutate()`.

Next we use the argument *pattern =* in our `stringr` function to denote what pattern we want to extract. If it wasn't for the 2.5 value, this pattern would be fairly simple. We could use something like "\\\\d\*" to say zero or more digits.

But since we have a "2.5" value, the pattern is a little more complicated. We say the pattern starts with zero or more digits ("^\\\\d\*") and then an optional period ("\\\\.?") and then zero or more digits ("\\\\d\*").


```{r}

d3 %>%
  dplyr::mutate(item1 = stringr::str_extract(item1, pattern = "^\\d*\\.?\\d*"))

```

* Note: It is important to note above that item1 continues to be a character variable. If we want to work with this as a numeric variable, we would need to further transform this variable into numeric.

<br>

### Package: tidyr

---

#### Function: `extract()`

---

**1\. Extract all numeric values from `item1`**

Review the data (d3)

```{r , echo=FALSE}

source("data.R")

d3

```

Using `tidyr::extract()` you no longer need to use `dplyr::mutate()` because you have the argument *into* where you name the new extracted variable/s.

The other thing to note is that in the *regex* argument, each pattern associated with an "into" variable needs to be surrounded by parentheses.

```{r}

d3 %>%
  tidyr::extract(col =item1, into = "item1", regex = "(^\\d*\\.?\\d*)")

```

Note again that the variable is still character and would need to be converted to numeric as a next step.

<br>

### Package: readr

---

#### Function: `parse_number()`

---

**1\. Extract all numeric values from `item1`**

Review the data (d3)

```{r, echo=FALSE}

source("data.R")

d3

```

There is another function `readr::parse_number()` that pulls out all numeric values from a string variable and converts the variable to numeric for us, which is pretty cool.


```{r}

d3 %>%
  dplyr::mutate(item1 = readr::parse_number(item1))

```



Return to [Strings](https://github.com/Cghlewis/data-wrangling-functions/wiki/Strings)
