---
title: "Extract values based on pattern"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: stringr

---

#### Function: `str_extract()`

---

**1\. Extract all numeric values from item1**

Review the data (d3)

```{r, echo=FALSE}

source("data.R")

d3

```

We first need to create our new variable "item1" which will write over our existing "item1" variable since I named it the same name. We do this using `dplyr::mutate()`.

Next we use the argument *pattern =* in our `stringr` function to denote what pattern we want to extract. If it wasn't for the 2.5 value, this pattern would be fairly simple. We could use something like "\\\\d\*" to say zero or more digits.

But since we have a "2.5" value, the pattern is a little more complicated. We say the pattern starts with zero or more digits ("^\\\\d\*") and then an optional period ("\\\\.?") and then zero or more digits ("\\\\d\*").


```{r}

d3 %>%
  dplyr::mutate(item1 = stringr::str_extract(item1, pattern = "^\\d*\\.?\\d*"))

```



Return to [Strings](https://github.com/Cghlewis/data-wrangling-functions/wiki/Strings)
