---
title: "Remove spaces"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: stringr

---

#### Function: `str_trim()`

---

**1\. Remove spaces from `item1`**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")

d1

```

I can see here that my variable `item1` has some values with added spaces. This is going to cause issues for me when I try to do things like filter or group using this variable.

For example if I want to get a table of values for item1, we can see that the yes values are not grouping together.

```{r}

d1 %>%
  janitor::tabyl(item1)

```

So I want to remove all white space on the left side of my variable.

* Note: I am using `dplyr::mutate()` to create a new variable "item1" that replaces the prior "item1" variable.

* Note: I used the argument *side = "left"* here but I could also say "right" or "both" (which is the default argument).

```{r}

d1 %>%
  dplyr::mutate(item1 = stringr::str_trim(item1, side = "left"))

```

**2\. Remove spaces from `item1` and `item2`**

Review the data (d2)

```{r, echo=FALSE}

source("data.R")

d2

```

We see here that we have spacing issues across multiple variables ("item1" and "item2") so I want to remove spaces from both variables. I can use `dplyr::across()` to efficiently take care of this.

```{r}

d2 %>%
  dplyr::mutate(dplyr::across(item1:item2, stringr::str_trim, side = "both"))

```

### Package: stringr

---

#### Function: `str_squish()`

---

**1\. Remove all excessive leading, trailing and middle spaces from `item1`**

Review the data (d7)

```{r, echo=FALSE}

source("data.R")

d7

```

Remove all spaces from `item1`

```{r}

d7 %>%
  dplyr::mutate(item1 = stringr::str_squish(item1))

```

Return to [Strings](https://github.com/Cghlewis/data-wrangling-functions/wiki/Strings)
