---
title: "Remove values"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: stringr

---

#### Function: `str_remove()`

---

**1\. Remove all "cm" or "CM" from item1**

Review the data (d3)

```{r, echo=FALSE}

source("data.R")

d3

```

In this case I only want the number, not the measurement type, so I need to remove all versions of "cm".

* Note: I am using `dplyr::mutate()` to create a new variable "item1" that will replace the existing "item1".

* Note: I am using `|` to say remove "cm" AND/OR "CM".

```{r}

d3 %>%
  dplyr::mutate(item1 = stringr::str_remove(item1, "cm|CM"))

```

If I did not want to list both the upper and lower case versions of "cm" I could also use another `stringr` function `str_to_lower()` first.

```{r}

d3 %>%
  dplyr::mutate(item1 = stringr::str_remove(stringr::str_to_lower(item1), "cm"))

```


**2\. Remove all instances of the word "and" from item1**

Review the data (d4)

```{r, echo=FALSE}

source("data.R")

d4

```

Notice here, if we simply use `stringr::str_remove()` we will only remove the first instance of "and".

```{r}

d4 %>%
  dplyr::mutate(item1 = stringr::str_remove(item1, "and"))

```

So therefore, if we want to remove all instances, we need to use `stringr::str_remove_all()`

```{r}

d4 %>%
  dplyr::mutate(item1 = stringr::str_remove_all(item1, "and"))

```


Return to [Strings](https://github.com/Cghlewis/data-wrangling-functions/wiki/Strings)
