---
title: "Replace values"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: stringr

---

#### Function: `str_replace()`

---

**1\. Replace values in `school`**

Review the data (d13)

```{r, echo=FALSE}

source("data.R")

d13

```

Replace "sr." and "(senior)" with "senior"

We first need to create a new variable "school" using `dplyr::mutate()` which will overwrite the existing `school` variable. We can then use `stringr::str_replace()` to replace our pattern using the *pattern* argument and denote our replacement using the *replacement* argument. 

* Note: The *pattern* argument uses regex arguments. Both periods and parentheses have special meaning but you can escape the special meaning using a double backslash or a bracket (as used in this case).

```{r}

d13 %>%
  dplyr::mutate(school = stringr::str_replace(school, pattern = "sr[.]|[(]senior[)]", replacement = "senior"))

```

**2\. Replace values in `item1`**

Review the data (d4)

```{r, echo=FALSE}

source("data.R")

d4

```

Replace all instances of the word "and" from `item1` with "or"

Notice here, if we simply use `stringr::str_replace()` we will only replace the first instance of "and".

```{r}

d4 %>%
  dplyr::mutate(item1 = stringr::str_replace(item1, "and", "or"))

```

So therefore, if we want to replace all instances, we need to use `stringr::str_replace_all()`

```{r}

d4 %>%
  dplyr::mutate(item1 = stringr::str_replace_all(item1, "and", "or"))

```

Return to [Strings](https://github.com/Cghlewis/data-wrangling-functions/wiki/Strings)
