---
title: "Truncate strings"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: stringr

---

#### Function: `str_trunc()`

---

**1\. Truncate all `_text` variables to a max of 10 characters**

Review the data (d14)

```{r, echo=FALSE}

source("data.R")

d14

```

In this case, we need to import this data into a program but the program only allows character columns to have a max of 10 characters.

We can see what our max count is for each variable right now.

* Note: Some variables contain NA. If we don't add any additional arguments to our `base::nchar` function, the value returned for those columns will be NA.


```{r}

d14 %>%
  dplyr::select(contains("text")) %>%
  lapply(., \(x) max(nchar(x)))

```

That is probably not what we want. So we can add the argument *keepNA = FALSE* to return a value when the column contains NA. If the column is all NAs, it will return a value of 2, the number of printing characters used when strings are written to output.

```{r}

d14 %>%
  dplyr::select(contains("text")) %>%
  lapply(., \(x) max(nchar(x, keepNA = FALSE)))

```


Now we can truncate our text variables.

```{r}

d14 <- d14 %>%
  dplyr::mutate(dplyr::across(contains("text"),
                ~stringr::str_trunc(., 10, "right")))

```

Let's see what the data looks like now

```{r}

d14

```

And let's see what the new max values are

```{r}

d14 %>%
  dplyr::select(contains("text")) %>%
  lapply(., \(x) max(nchar(x, keepNA = FALSE)))

```


Return to [Strings](https://github.com/Cghlewis/data-wrangling-functions/wiki/Strings)
