---
title: "Create new folders"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)

library(fs)
library(here)
library(tidyverse)
```

<br>

### Package: fs

---

#### Function: `dir_create()`

---

**1\. Create a new observation folder with today's date**

* Note: I used the `here::here()` function which implicitly sets my directory to the top level (root directory) of my current project. Then any subsequent folders can be listed in descending order. In this case I want all my observation folders to be in a folder called "directories" which is located in the top level of my project directory. You could also just list your entire file path.

Let's see what folders currently exist in our "directories" folder

* Note: A reminder that `base::basename()` removes all of the path up to and including the last path separator so we can simply view the folder names here

```{r, eval = FALSE}

base::basename(fs::dir_ls(here::here("directories")))

```

```{r, echo = FALSE}

base::basename(fs::dir_ls(here::here("directories"))) %>%
  stringr::str_subset("obs_")

```

Let's make another folder called "obs_2022-03-16" (which is today's date at the time of creating this file).

* Note: I use the `lubridate::today()` function to grab today's date in the form of YYYY-MM-DD

```{r, eval = FALSE}

fs::dir_create(here::here("directories", paste0("obs_",lubridate::today())))

```

```{r, echo = FALSE}

fs::dir_create(here::here("directories", paste0("obs_", "2022-03-16")))

```


Now let's see what folders currently exist in our "directories" folder

```{r, eval = FALSE}

base::basename(fs::dir_ls(here::here("directories")))

```

```{r, echo = FALSE}

base::basename(fs::dir_ls(here::here("directories"))) %>%
  stringr::str_subset("obs_")

```

