---
title: "Data frame comparison"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: diffdf

---

#### Function: `diffdf`

---

**1\. Compare two data frames that should be identical to see if there are any differences in their values**

Read in the data

* Note: Here we read in csv files but of course you can read in any type of data (.xlsx, .sav, etc)

```{r, eval=FALSE}

entry1 <- readr::read_csv("project-a_forms_entry1.csv")
entry2 <- readr::read_csv("project-b_forms_entry2.csv")

```


Review entry1

```{r, echo=FALSE}

source("data.R")
head(entry1)

```

Review entry2

```{r, echo=FALSE}

source("data.R")
head(entry2)

```

Now check if there are any differences in the two data frames.  

* Note: For the *key* argument I added "stu_id". This is because "stu_id" is what makes each row unique. If you needed multiple variables to make a row unique, you can simply add them all using combine. For example, *keys = c("stu_id", "tch_id")*.

* Note: I've included the base arguments you need for the function to run. Run *?diffdf* in the console to see additional arguments that can be added.

```{r}

diffdf::diffdf(entry1,entry2, keys="stu_id")

```

You can see our output tells us that **there are differences** between the two data frames. 

1. The first difference is found in row *stu_id = 1234*. The *base* data frame (entry1) entered 2 for "q1" and the *compare* data frame (entry2) entered 1. 

2. The second difference found is in row *stu_id = 1236*. The *base* data frame entered 12 for "q2" and the *compare* data frame entered 1.

You could use this information to return to your original forms to see what the correct answer should be and then rectify the errors in whichever data frame has the incorrect response. You would then read your revised data back into R and then run this code again to see if all errors have been corrected.

If all errors have been corrected, you will see this message.

```{r, echo=FALSE}

diffdf::diffdf(entry1,entry1, keys="stu_id")

```

