---
title: "Data frame column type comparison"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>


### Package: janitor

---

#### Function: `compare_df_cols`

---

**1\. Compare two data frames that should have identical column types to see if there are any differences**

Read in the data

* Note: Here we read in csv files but of course you can read in any type of data (.xlsx, .sav, etc)

```{r, eval=FALSE}

entry1 <- readr::read_csv("project-a_forms_entry1.csv")
entry3 <- readr::read_csv("project-b_forms_entry3.csv")

```


Review entry1

```{r, echo=FALSE}

source("data.R")
head(entry1)

```

Review entry3

```{r, echo=FALSE}

source("data.R")
head(entry3)

```

Now check if there are any differences in the column classes between the two data frames.
The function `janitor::compare_df_cols()` will indicate if data frames will successfully bind together by rows.

```{r}

janitor::compare_df_cols(entry1, entry3)

```

If we only wanted to see the differences, we could add the argument *return = "mismatch"*. We can easily see here that grade is a different variable type across files and we will need to transform the variable in entry1 or entry 3 in if we want to bind the two data frames.

```{r}

janitor::compare_df_cols(entry1, entry3, return = "mismatch")

```


Return to [Compare data frames](https://github.com/Cghlewis/data-wrangling-functions/wiki/Compare data frames)
