---
title: "Conditional Filtering"
output: 
  html_document:
    anchor_sections: FALSE
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```


### Package: dplyr

---

#### Function: `filter()`

---

**1\.Conditionally filter grade levels based on school name. Filter to only grade=1 for school a and grade=9 for school b**

Review the data (d11)

```{r, echo=FALSE}

source("data.R")
d11

```

Conditionally filter

* Note: Note that we put each condition within () and add the `and/or` operator between the two conditions

```{r}

d11 %>%
  filter((sch == "a" & grade == 1) | (sch == "b" & grade == 9))

```

You can use this same conditional filtering for more than just getting exact matches. Here is an example where we use the same logic but get all grades >= 9 for school b.

```{r}

d11 %>%
  filter((sch == "a" & grade == 1) | (sch == "b" & grade >= 9))

```


**2\.Conditionally filter grade levels based on school name. Filter to grades between 1 & 2 for school a and grade=9 for school b**

Review the data (d11)

```{r, echo=FALSE}

source("data.R")
d11

```

Conditionally filter

```{r}

d11 %>%
  filter((sch == "a" & between(grade, 1, 2)) | (sch == "b" & grade==9))

```

I could have also used *and/or* rather than *between* in this situation.

```{r}

d11 %>%
  filter((sch == "a" & (grade == 1 | grade ==2)) | (sch == "b" & grade == 9))

```

I also could have used `%in%

```{r}

d11 %>%
  filter((sch == "a" & (grade %in% 1:2)) | (sch == "b" & grade == 9))
```

**3\. Conditionally filter based on school name. Filter to only grade=1 for school a and grade=9 for school b, but also keep all other schools, no matter the grade level**

Review the data (d15)

Notice this dataset also has a school c

```{r, echo = FALSE}

d15

```

Conditionally filter

```{r}

d15 %>%
  filter((sch == "a" & grade == 1) | (sch == "b" & grade == 9) | (sch == "c"))

```

If I did not add a condition to retain school c, only a and b would remain

```{r}

d15 %>%
  filter((sch == "a" & grade == 1) | (sch == "b" & grade == 9))

```

Return to [Filter](https://github.com/Cghlewis/data-wrangling-functions/wiki/Filter)
