---
title: "Conditional Filtering"
output: 
  html_document:
    anchor_sections: FALSE
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```


### Package: dplyr

---

#### Function: `filter()`

---

**1\.Conditionally filter grade levels based on school name. Filter to only grade=1 for school a and grade=9 for school b**

Review the data (d11)

```{r, echo=FALSE}

source("data.R")
d11

```

Conditionally filter

* Note: Note that we put each condition within () and add the `and/or` operator between the two conditions

```{r}

d11 %>%
  filter((sch == "a" & grade==1) | (sch == "b" & grade==9))

```


**2\.Conditionally filter grade levels based on school name. Filter to grades between 1 & 2 for school a and grade=9 for school b**

Review the data (d11)

```{r, echo=FALSE}

source("data.R")
d11

```

Conditionally filter

```{r}

d11 %>%
  filter((sch == "a" & between(grade, 1, 2)) | (sch == "b" & grade==9))

```

I could have also used *and/or* rather than *between* in this situation.

```{r}
d11 %>%
  filter((sch == "a" & (grade == 1 | grade ==2)) | (sch == "b" & grade==9))

```



#### Function: `semi_join()`


In the filtering join `semi_join(data_table, condition_table)`, a data table is
filtered using a condition table. Only data rows that match a row in the
condition table are kept. (`anti_join()` is a similar function but it filters *out*
rows.)

**1\.Conditionally filter grade levels based on school name. Filter to only grade=1 for school a and grade=9 for school b**

```{r}
condition1 <- tibble(
  sch = c("a", "b"),
  grade = c(1, 9)
)

d11 %>%
  semi_join(condition1)
```

**2\.Conditionally filter grade levels based on school name. Filter to grades between 1 & 2 for school a and grade=9 for school b**



```{r}
condition2 <- tibble(
  sch = c("a", "a", "b"),
  grade = c(1, 2, 9)
)

d11 %>%
  semi_join(condition2)
```

Note that this approach does not handle more open-ended criteria (`grade > 1`)
very well.