---
title: "Filter rows based on one criteria for one variable"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```


### Package: dplyr

---

#### Function: `filter()`

Examples using one criteria for one variable (character)

---

**1\. Keep any row that has the exact match of "harris" for `tch_name`.**

Review the data (d8)

```{r, echo=FALSE}

source("data.R")
d8

```

Keep any row that has "harris" for `tch_name`. 

* Note: Use the logical operator *==* to denote exactly equal to.

```{r}

d8 %>% 
  dplyr::filter(tch_name == "harris")

```

You could get the same result adding the `stringr::str_detect()` function.

```{r}

d8 %>% 
  dplyr::filter(stringr::str_detect(tch_name, 'har') )

```

**2\. Remove any row that has the exact match of "harris" for `tch_name`.**

Review the data (d8)

```{r, echo=FALSE}

source("data.R")
d8

```


Filter out any row that has "harris" for `tch_name`. 

* Note: Use the logical operator *!=* to denote not equal to.

```{r}

d8 %>% 
  dplyr::filter(tch_name != "harris")

```

Again, you could get the same result adding the `stringr::str_detect()` function and adding the *!* operator.

```{r}

d8 %>% 
  dplyr::filter(!stringr::str_detect(tch_name, 'har'))

```

You could also remove the *!* and add the `stringr::str_detect` argument *negate=TRUE* which returns non-matching elements.

```{r}

d8 %>% 
  dplyr::filter(stringr::str_detect(tch_name, 'har', negate=TRUE))

```

---

#### Function: `filter()`

Examples using one criteria for one variable (date)

---

**3\. Keep any row where `test_date` occurred in the year 2021.**

Review the data (d12)

```{r, echo=FALSE}

source("data.R")
d12

```

First lets check to see what type of variable `test_date` is.

We find that it is a character. In order to filter as a date, we either need to convert the variable type to date now, or within our filter.

```{r}

class(d12$test_date)

```

Keep any row where the test date occurred in the year 2021.

* Note: We added lubridate::as_date in order to change the variable class to date

* Note: Use the logical operator *>=* to denote greater than or equal to.


```{r}

d12 %>% 
  dplyr::filter(lubridate::as_date(test_date) >= "2021-01-01")

```

* Note: I knew that the highest year in the data was 2021 so I didn't need to add a less than for this function. However if there were years greater than 2021, I could write the function like this.

```{r}

d12 %>% 
  dplyr::filter(lubridate::as_date(test_date) >= "2021-01-01" &
                  lubridate::as_date(test_date) < "2022-01-01")

```

I could also simplify this by not adding the day and month and just looking for the year by using the `lubridate::year()` function.

```{r}

d12 %>% 
  dplyr::filter(lubridate::year(lubridate::as_date(test_date)) ==
                  "2021")

```


---

#### Function: `filter()`

Examples using one criteria for one variable (numeric)

---


**1\. Keep any row whose value is greater than or equal to 0 for `extra2`.**

Review the data (d8)

```{r, echo=FALSE}

source("data.R")
d8

```

Keep any row that has a value greater than or equal to zero for `extra2`. 

* Note: Use the logical operator *>=* to denote greater than or equal to.

```{r}

d8 %>% 
  dplyr::filter(extra2 >= 0)

```


Return to [Filter](https://github.com/Cghlewis/data-wrangling-functions/wiki/Filter)
