---
title: "Filter rows using if_all or if_any"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```


### Package: dplyr

---

#### Function: `filter()`

---

* Note: Using `dplyr::across()` in `dplyr::filter()` is deprecated. `dplyr::if_any()` and `dplyr::if_all()` are predicate functions used to select columns within `dplyr::filter()`. This function is available in version 1.0.5 of `dplyr`. `dplyr::if_any()` returns a true when the statement is true for **any** of the variables. `dplyr::if_all()` returns a true when the statement is true for **all** of the variables.


<br>

**1\. Keep rows where respondent has missing responses for AT LEAST ONE variable (var1:var4). Missing data is denoted by "MR"**

Review the data (d20)

```{r, echo=FALSE}

source("data.R")
d20

```

You can read this as "If a row has a value of \"MR" for ANY variables (var1:var4), return the row"


```{r}

d20 %>%
  dplyr::filter(dplyr::if_any(var1:var4, ~ . == "MR"))

```

**2\. Keep rows where respondent has missing responses for ALL variables (var1:var4). Missing data is denoted by "MR"**

Review the data (d20)

```{r, echo=FALSE}

source("data.R")
d20

```

You can read this as "If a row has a value of \"MR" for ALL variables (var1:var4), return the row"

```{r}

d20 %>%
  dplyr::filter(dplyr::if_all(var1:var4, ~ . == "MR"))

```

**3\. Remove rows where respondent has missing responses for ALL variables (var1:var4). Missing data is denoted by "MR"**

Review the data (d20)

```{r, echo=FALSE}

source("data.R")
d20

```

This example can be solved in one of two ways. Use whichever makes the most sense to you.

The first and probably least confusing way to me is using `dplyr::if_all()`.

Here we use the same formula that we used in #2, where we first use `dplyr::if_all()` to pull rows that have values of \"MR" for ALL variables. 

But then, we add the `!` operator outside of `dplyr::if_all()` to say, after we grab those rows, we want to drop them.

```{r}

d20 %>%
  dplyr::filter(!dplyr::if_all(var1:var4, ~ . == "MR"))

```

The other way we can solve this is reframing our question to say, **"Keep rows where respondent has responses that are NOT missing for ANY variable (var1:var4)".**

Reframing our question this way allows us to see how we can also you `dplyr::if_any()` to get the rows we want.

You can read the code below as "If a row has a value that is **NOT** \"MR" for ANY variable (var1:var4), return the row"

Because if a row has a value that is NOT \"MR" for any variable (var1:var4) then they are NOT missing all responses and should be returned.

```{r}

d20 %>%
  dplyr::filter(dplyr::if_any(var1:var4, ~ . != "MR"))

```


**4\. Remove rows where respondent has missing responses for AT LEAST ONE variable (var1:var4). Missing data is denoted by "MR"**

Review the data (d20)

```{r, echo=FALSE}

source("data.R")
d20

```

Again, this example can be solved in one of two ways. Use whichever makes the most sense to you.

The first and probably least confusing way to me is using `dplyr::if_any()`.

Here we use the same formula that we used in #1, where we first use `dplyr::if_any()` to pull rows that have values of \"MR" for ANY variables. 

But then, we add the `!` operator outside of `dplyr::if_any()` to say, after we grab those rows, we want to drop them.

```{r}

d20 %>%
  dplyr::filter(!dplyr::if_any(var1:var4, ~ . == "MR"))

```

The other way we can solve this is reframing our question to say, **"Keep rows where respondent has NO missing responses across ALL variables (var1:var4)".**

Reframing our question this way allows us to see how we can also you `dplyr::if_all()` to get the rows we want.

You can read the code below as “If values for ALL variables (var1:var4) in a row are NOT \"MR”, return the row”


```{r}

d20 %>%
  dplyr::filter(dplyr::if_all(var1:var4, ~ . != "MR"))

```



Return to [Filter](https://github.com/Cghlewis/data-wrangling-functions/wiki/Filter)
