---
title: "Filter based on row sums or means"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```


### Package: dplyr

---

#### Function: `filter()`

---

**1\. Keep rows that have two or less values of "MR" across var1:var4.**

Review the data (d23)

```{r, echo=FALSE}

source("data.R")
d23

```

Keep rows that have two or less values of "MR" across var1:var4

* Note: Here I am not using `if_any()` or `if_all()` like we typically use in a filter statement. I am using `dplyr::across()` because we are selecting variables within `base::rowSums()`.

* Note: I am using an anonymous function within our `dplyr::across()` statement to check if the values of our variables are equal to "MR".

* Note: I add the `base::rowSums()` argument, *na.rm = TRUE* because there are NA values in my data and I still want my "MR" values summed for rows where there are NA values.

```{r}

d23 %>%
  dplyr::filter(rowSums(dplyr::across(var1:var4, ~ . == "MR"), na.rm = TRUE) < 2)

```

If the pattern we were trying to count was not always exactly the same across variables, we could use something like `stringr::str_detect()` to capture values instead.

```{r}

d23 %>%
  dplyr::filter(rowSums(dplyr::across(var1:var4, ~ stringr::str_detect(., "MR")), na.rm = TRUE) < 2)

```

**2\. Keep rows where the mean of `var1` and `var2` is greater than 2.**

Review the data (d24)

```{r, echo=FALSE}

source("data.R")
d24

```

Here we don't need to add an anonymous function within `dplyr::across()`.

```{r}

d24 %>%
  dplyr::filter(rowMeans(dplyr::across(var1:var2), na.rm = TRUE) > 2)

```


Return to [Filter](https://github.com/Cghlewis/data-wrangling-functions/wiki/Filter)
