---
title: "Filter rows using if_all or if_any"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```


### Package: dplyr

---

#### Function: `filter()`

---

* Note: Using `dplyr::across()` in `dplyr::filter()` is deprecated. `dplyr::if_any()` and `dplyr::if_all()` are predicate functions used to select columns within `dplyr::filter()`. This function is available in version 1.0.5 of `dplyr`. `dplyr::if_any()` returns a true when the statement is true for **any** of the variables. `dplyr::if_all()` returns a true when the statement is true for **all** of the variables.


<br>

**1\. Keep rows where respondent has responses that are _not_ missing for ANY variable (var1:var4). Missing data is denoted by "MR"**

Review the data (d20)

```{r, echo=FALSE}

source("data.R")
d20

```

You can read this as "If a row has a value that is **NOT** \"MR" for ANY variable (var1:var4), return the row"

```{r}

d20 %>%
  dplyr::filter(dplyr::if_any(var1:var4, ~ . != "MR"))

```

**2\. Keep rows where respondent has missing reponses for ALL variables (var1:var4). Missing data is denoted by "MR"**

Review the data (d20)

```{r, echo=FALSE}

source("data.R")
d20

```

You can read this as "If a row has a value of \"MR" for ALL variables (var1:var4), return the row"

```{r}

d20 %>%
  dplyr::filter(dplyr::if_all(var1:var4, ~ . == "MR"))

```


**3\. Keep rows where respondent has missing responses for ANY variable (var1:var4). Missing data is denoted by "MR"**

Review the data (d20)

```{r, echo=FALSE}

source("data.R")
d20

```

You can read this as "If a row has a value of \"MR" for ANY variables (var1:var4), return the row"


```{r}

d20 %>%
  dplyr::filter(dplyr::if_any(var1:var4, ~ . == "MR"))

```

**4\. Keep rows where respondent has _no_ missing responses for ALL variables (var1:var4). Missing data is denoted by "MR"**

Review the data (d20)

```{r, echo=FALSE}

source("data.R")
d20

```

You can read this as "If values for ALL variables (var1:var4) in a row are **NOT** \"MR", return the row"


```{r}

d20 %>%
  dplyr::filter(dplyr::if_all(var1:var4, ~ . != "MR"))

```

**5\. Remove rows where respondent has missing responses for ALL variables (var1:var4). Missing data is denoted by "MR"**

Review the data (d20)

```{r, echo=FALSE}

source("data.R")
d20

```

Removing rows is where things get complicated. 

So above (in #2) we returned every row that has "MR" for ALL our variables (var1:var4). And it seems like to get the opposite, we would just use `!=` instead of `==`, but as we see above (in #4) that gives us a different result. That gives us rows that do not contain "MR" as a value across ALL of our variables.

So if we want to **remove** the rows that we grabbed in #2, we actually need to move our `!` to a new location, to before our `dplyr::if_all()` statement. This way we still grab the rows from #2, but we are saying to "drop" them this time.

```{r}

d20 %>%
  dplyr::filter(!dplyr::if_all(var1:var4, ~ . == "MR"))

```




Return to [Filter](https://github.com/Cghlewis/data-wrangling-functions/wiki/Filter)
