---
title: "Filter rows based on one criteria that apply to all variables"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```


### Package: dplyr

---

#### Function: `filter()`

Examples using criteria that applies to all variables

---

* Note: Using `dplyr::across()` in `dplyr::filter()` is deprecated. `dplyr::if_any()` and `dplyr::if_all()` are predicate functions used to select columns within `dplyr::filter()`. This function is available in version 1.0.5 of `dplyr`.

**1\. Remove any row that has -999 for *AT LEAST ONE* variable.**

Review the data (d19)

```{r, echo=FALSE}

source("data.R")
d19

```


Filter out any row that has -999 for at least one variable

* Note: You can read `dplyr::if_any()` here as, keep rows where ANY variable is NOT -999 (hence the ! on the outside of the `dplyr::if_any()` function)

* Note: We use the `tidyselect` selection helper `everything()` to refer to all variables.


```{r}

d19 %>% 
  dplyr::filter(!dplyr::if_any(everything(), ~ . %in% -999))

```


**2\. Keep any row that has -999 for *ALL* variables.**

Review the data (d19)

```{r, echo=FALSE}

source("data.R")
d19

```

Keep any row that has -999 for all variables

* Note: You can read `dplyr::if_all()` here as, keep rows where ALL variables are -999

```{r}

d19 %>% 
  dplyr::filter(dplyr::if_all(everything(), ~ . == -999))

```


Return to [Filter](https://github.com/Cghlewis/data-wrangling-functions/wiki/Filter)
