---
title: "Filter empty rows"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```


### Package: dplyr

---

#### Function: `filter()`

---

* Note: Using `dplyr::across()` in `dplyr::filter()` is deprecated. `dplyr::if_any()` and `dplyr::if_all()` are predicate functions used to select columns within `dplyr::filter()`. This function is available in version 1.0.5 of `dplyr`. `dplyr::if_any()` returns a true when the statement is true for **any** of the variables. `dplyr::if_all()` returns a true when the statement is true for **all** of the variables. See [Filter using if_all or if_any](https://cghlewis.github.io/data-wrangling-functions/filter-data/filter-rows-if.html) for further explanation


<br>

**1\. Remove empty rows**

Review the data (d9)

```{r, echo=FALSE}

source("data.R")
d9

```

Filter out any completely empty row

* Note: Note the use of `tidyselect::everything()` to select all variables


```{r}

d9 %>% 
  dplyr::filter(!dplyr::if_all(tidyselect::everything(), ~ is.na(.)))


```



### Package: expss

---

#### Function: `drop_empty_rows()`

---

**1\. Filter out empty rows**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```


Filter out any completely empty row

```{r}

d1 %>% 
  expss::drop_empty_rows()

```


### Package: janitor

---

#### Function: `remove_empty()`

---


**1\. Remove empty rows**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Filter out any completely empty row

```{r}

d1 %>% 
  janitor::remove_empty("rows")

```


Return to [Filter](https://github.com/Cghlewis/data-wrangling-functions/wiki/Filter)
