---
title: "Filter rows based on matching variable values"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```


### Package: dplyr

---

#### Function: `filter()`

---

**1\. Keep any row where q1 equals q2.**

Review the data (d13)

```{r, echo=FALSE}

source("data.R")
d13

```

Filter rows

```{r}

d13 %>% 
  dplyr::filter(q1 == q2)

```


**2\. Remove any row where q1 does not equal q2, while also keeping rows that have NA for q2.**

Review the data (d13)

```{r, echo=FALSE}

source("data.R")
d13

```

Filter out any row where q1 is not equal to q2

* Note: Use the logical operator *!=* to denote not equal to.

* Note: I make an explicit call to **keep** rows with NA values for q2. If I did not do this, filter would drop the last row of data.

```{r}

d13 %>% 
  dplyr::filter(q1 != q2 | is.na(q2))

```


There are still other ways to keep NAs as well. As we saw in [NA] there are other ways to keep NA values using the *%in%* operator. 

However, in this scenario where we are comparing two vectors, you cannot use the *%in%* operator the same way we that we used it in [NA values]. Here you need to add the `dplyr::rowwise()` function. This is also a better method when both q1 and q2 might have NA values and you want to keep rows when either column has NA, but not when both columns have NA.


```{r}

d13 %>% 
  dplyr::rowwise() %>%
  dplyr::filter(!q1 %in% q2)

```


And yet, another way you may want to keep rows in your data is using the `dplyr::case_when()` function. 

```{r}

d13 %>%
  dplyr::filter(
    dplyr::case_when(
      q1 == q2 ~ FALSE,
      TRUE ~ TRUE))

```


Return to [Filter](https://github.com/Cghlewis/data-wrangling-functions/wiki/Filter)
