---
title: "Filter rows based on one criteria for multiple variables"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```


### Package: dplyr

---

#### Function: `filter()`

Examples selecting specific variables

---

**1\. Keep rows that have -999 for extra1 *AND/OR* -999 for extra2.**

Review the data (d8)

```{r, echo=FALSE}

source("data.R")
d8

```

Keep rows that have -999 for extra2 and/or extra3. 

* Note: Use of the *|* operator to request and/or.

```{r}

d8 %>% 
  dplyr::filter(extra2 == -999 | extra3 == -999)


```

**2\. Keep rows that have -999 for extra1 *OR* -999 for extra2 (not for both).**

Review the data (d8)

```{r, echo=FALSE}

source("data.R")
d8

```

Keep rows that have -999 for extra2 OR -999 for extra3 (but not -999 for both). 

* Note: Use the *xor* operator to request OR (not both).

```{r}

d8 %>% 
  dplyr::filter(xor(extra2 == -999 , extra3 == -999))


```

**3\. Keep rows that have *BOTH* -999 for extra1 *AND* -999 for extra2.**

Review the data (d8)

```{r, echo=FALSE}

source("data.R")
d8

```

Keep rows that have -999 for extra2 and also have -999 extra3. 

* Note: Use the *&* operator to request and.

```{r}

d8 %>% 
  dplyr::filter(extra2 == -999 & extra3 == -999)

```

**4\. Remove any row that has -999 for extra2 *AND/OR* has -999 for extra3.**

Review the data (d8)

```{r, echo=FALSE}

source("data.R")
d8

```

Filter out any row that has -999 for extra2 and/or has -999 extra3. 

* Note: Notice that I put the operator *!* outside of the parentheses. While there are other ways to write this equation, I find this the least confusing way to remove rows based on removal criteria from multiple variables.

```{r}

d8 %>% 
  dplyr::filter(!(extra2 ==-999 | extra3 ==-999))

```


**5\. Remove any row that has -999 for *BOTH* extra2 *AND* for extra3.**

Review the data (d8)

```{r, echo=FALSE}

source("data.R")
d8

```

Filter out any row that has -999 for extra2 and has -999 extra3. 

* Note: Notice that I put the operator *!* outside of the parentheses. While there are other ways to write this equation, I find this the least confusing way to remove rows based on removal criteria from multiple variables.


```{r}

d8 %>% 
  dplyr::filter(!(extra2 ==-999 & extra3 ==-999))


```

---

#### Function: `filter()`

Examples selecting across variables

---

* Note: Using `dplyr::across()` in `dplyr::filter()` is deprecated. `dplyr::if_any()` and `dplyr::if_all()` are predicate functions used to select columns within `dplyr::filter()`. This function is available in version 1.0.5 of `dplyr`.

<br>

**1\. Keep rows where any variable that starts with the word "item" contains the value "MR".**

Review the data (d16)

```{r, echo=FALSE}
source("data.R")
d16
```

Keep rows where any of the item variables contains the value "MR"

* Note: You can read `dplyr::if_any()` here as, keep rows where ANY variable that starts with "item" is equal to "MR"

* Note: We use the `tidyverse` selection helper `starts_with()` to select variables.

```{r}
d16 %>%
  dplyr::filter(dplyr::if_any(tidyselect::starts_with("item"), ~ . == "MR"))
```

**2\. Keep rows where any variable that starts with the word "item" contain letters.**

Review the data (d16)

```{r, echo = FALSE}
d16
```

We want to review any rows where either item1 OR item2 contain letters.

* Note: You can read `dplyr::if_any()` here as, keep rows where ANY of the variables we denoted (item1 or item2), contain letters.

* Note: We use `stringr::str_detect()` to detect a specific pattern, assigned in the *pattern* argument. Here we use a regular expression to denote that we would like to identify zero or more letters from the A-Z alphabet, lower or upper case.

```{r}
d16 %>% 
  dplyr::filter(dplyr::if_any(item1:item2, 
                ~ stringr::str_detect(., pattern = "[a-zA-Z]")))
```

Return to [Filter](https://github.com/Cghlewis/data-wrangling-functions/wiki/Filter)
