---
title: "Filter rows based on variable type"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```


### Package: dplyr

---

#### Function: `filter()`

Examples using criteria that applies to a type of variable

---

*Note: Using `dplyr::across()` in `dplyr::filter()` is deprecated. `dplyr::if_any()` and `dplyr::if_all()` can be used for this functionality in `dplyr::filter()`.

**1\.Remove rows based on values from a certain type of variable (in this case numeric variables). Filter out any row that has NA for *ALL* numeric variables.**

Review the data (d9)

```{r, echo=FALSE}

source("data.R")
d9

```

Filter out any row that has NA for **ALL** numeric variables

* Note: Note the use of `dplyr::if_all()` predicate function to select columns. This function is available in version 1.0.5 of `dplyr`. You can read this as, keep rows where ALL numeric variables are NOT NA (hence the ! on the outside of the `dplyr::if_all()` function)


```{r}

d9 %>% 
  dplyr::filter(!(dplyr::if_all(is.numeric,~ is.na(.))))


```


**2\. Remove rows based on values from a certain type of variable (in this case numeric). Filter out any row that has -999 for *AT LEAST ONE* numeric variable.**

Review the data (d9)

```{r, echo=FALSE}

source("data.R")
d9

```

Filter out any row that has -999 for **AT LEAST ONE** numeric variable

* Note: Note the use of `dplyr::if_any()` predicate function to select columns. This function is available in version 1.0.5 of `dplyr`. You can read this as, keep rows where ANY numeric variables are NOT -999 (hence the ! on the outside of the `dplyr::if_any()` function)

* Note: You'll notice here that I used `%in%` rather than `==` to ensure I kept the NA values. 

```{r}

d9 %>% 
  dplyr::filter(!(dplyr::if_any(is.numeric, ~ . %in% -999)))

```

Return to [Filter](https://github.com/Cghlewis/data-wrangling-functions/wiki/Filter)
