---
title: "Filter rows based on criteria that apply to all or a selection of variables"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```


### Package: dplyr

---

#### Function: `filter()`

Examples using criteria that applies to all variables

---

* Note: Using `dplyr::across()` in `dplyr::filter()` is deprecated. `dplyr::if_any()` and `dplyr::if_all()` can be used for this functionality in `dplyr::filter()`.

**1\.Filter out rows based on values from all variables. Filter out any row that has -999 for *AT LEAST ONE* variable.**

Review the data (d10)

```{r, echo=FALSE}

source("data.R")
d10

```


Filter out any row that has -999 for at least one variable

* Note: Note the use of `dplyr::if_any()` predicate function to select columns. This function is available in version 1.0.5 of `dplyr`. You can read this as, keep rows where ANY variable is NOT NA (hence the ! on the outside of the `dplyr::if_any()` function)

* Note: We use the `tidyselect` selection helper `everything()` to refer to all variables.

* Note: You'll notice here that I used `%in%` rather than `==` to ensure I kept the NA values. 

```{r}

d10 %>% 
  dplyr::filter(!(dplyr::if_any(everything(), ~ .x %in% -999)))

```


**2\.Filter out rows based on values from all variables. Filter out any row that has -999 for *ALL* variables.**

Review the data (d10)

```{r, echo=FALSE}

source("data.R")
d10

```

Filter out any row that has -999 for all variables

* Note: Note the use of `dplyr::if_all()` predicate function to select columns. This function is available in version 1.0.5 of `dplyr`. You can read this as, keep rows where ALL variables are NOT -999 (hence the ! on the outside of the `dplyr::if_all()` function)

* Note: We use the `tidyverse` selection helper `everything()` to refer to all variables.

* Note: You'll notice here that I used `%in%` rather than `==` to ensure I kept the NA values. 

```{r}

d10 %>% 
  dplyr::filter(!(dplyr::if_all(everything(), ~ .x %in% -999)))

```

**3\. Keep rows based on a value from an selection of variables. Keep rows where any variable that starts with the word "item" contains the value "MR".**

Review the data (d16)

```{r, echo=FALSE}

source("data.R")
d16

```

Keep rows where any of the item variables contains the value "MR"

* Note: Note the use of `dplyr::if_any()` predicate function to select columns. This function is available in version 1.0.5 of `dplyr`. You can read this as, keep rows where ANY variable that starts with "item" is equal to "MR"

* Note: We use the `tidyverse` selection helper `starts_with()` to select variables.

```{r}

d16 %>%
  dplyr::filter(dplyr::if_any(tidyselect::starts_with("item"), ~ .x == "MR"))

```


Return to [Filter](https://github.com/Cghlewis/data-wrangling-functions/wiki/Filter)
