---
title: "Filter rows based on multiple criteria"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```


### Package: dplyr

---

#### Function: `filter()`

Examples using multiple criteria same variables

---

**1\. Keep rows based on multiple criteria from one variable (in this case for a numeric variable). Keep any row that has -999 *OR* 0 for extra3.**

Review the data (d8).

```{r, echo=FALSE}

source("data.R")
d8

```

Keep any row that has -999 or 0 for extra2. 

* Note: Use of the operator *|* to denote and/or (*and* doesn't really apply since it's the same variable).

```{r}

d8 %>% 
  dplyr::filter(extra3 == -999 | extra3 ==0)

```

You would get the same result using the *xor* operator.

```{r}

d8 %>% 
  dplyr::filter(xor(extra3 == -999 , extra3 ==0))

```


**2\. Keep rows based on multiple criteria from one variable (in this case for a numeric variable). Keep any row whose value for extra3 is between 0 and 10.**

Review the data (d8).

```{r, echo=FALSE}

source("data.R")
d8

```


Keep any row that has a value between 0 and 10 for extra3. 

* Note: Use the operator *&* to denote that both criteria must be met.

```{r}

d81 <- d8 %>% 
  dplyr::filter(extra3 >= 0 & extra3 <=10)

d81

```

You would get the same result using the `dplyr::between()` function. It includes both 0 and 10.

```{r}

d8 %>% 
  dplyr::filter(dplyr::between(extra3, 0, 10))

```


**3\. Keep rows based on multiple criteria from one variable (in this case for a character variable). Keep any row whose tch_name is either "harris" *OR* "lewis".**

Review the data (d8).

```{r, echo=FALSE}

source("data.R")
d8

```

Keep any row where tch_name is "harris" or "lewis".

* Note: Use of the operator *|* to denote and/or (*and* doesn't really apply since it's the same variable).

```{r}

d8 %>% 
  dplyr::filter(tch_name == "harris" | tch_name == "lewis")

```

You would get the same result using the *%in%* operator.

```{r}

d8 %>% 
  dplyr::filter(tch_name %in% c("harris", "lewis"))

```

Or the same result adding the `stringr::str_detect()` function.

```{r}

d8 %>% 
  dplyr::filter(stringr::str_detect(tch_name, 'harris|lewis') )

```

---

#### Function: `filter()`

Examples using multiple criteria different variables, keep rows

---


**1. Keep rows based on multiple criteria (in this case numeric variables). Keep rows that have -999 for extra1 *AND/OR* -999 for extra2.**

Review the data (d8)

```{r, echo=FALSE}

source("data.R")
d8

```

Keep rows that have -999 for extra2 and/or extra3. 

* Note: Use of the *|* operator to request and/or.

```{r}

d8 %>% 
  dplyr::filter(extra2 == -999 | extra3 == -999)


```

**2. Keep rows based on multiple criteria (in this case numeric variables). Keep rows that have -999 for extra1 *OR* -999 for extra2 (not for both).**

Review the data (d8)

```{r, echo=FALSE}

source("data.R")
d8

```

Keep rows that have -999 for extra2 OR -999 for extra3 (but not -999 for both). 

* Note: Use the *xor* operator to request OR (not both).

```{r}

d8 %>% 
  dplyr::filter(xor(extra2 == -999 , extra3 == -999))


```

**3. Keep rows based on multiple criteria (in this case numeric variables). Keep rows that have *BOTH* -999 for extra1 *AND* -999 for extra2.**

Review the data (d8)

```{r, echo=FALSE}

source("data.R")
d8

```

Keep rows that have -999 for extra2 and also have -999 extra3. 

* Note: Use the *&* operator to request and.

```{r}

d8 %>% 
  dplyr::filter(extra2 == -999 & extra3 == -999)

```

---

#### Function: `filter()`

Examples using multiple criteria from different variables, remove rows

---


**1\. Remove rows based on multiple criteria from multiple variables (in this case for two numeric variables). Filter out any row that has -999 for extra2 *AND/OR* has -999 for extra3.**

Review the data (d8)

```{r, echo=FALSE}

source("data.R")
d8

```

Filter out any row that has -999 for extra2 and/or has -999 extra3. 

* Note: Notice that I put the operator *!* outside of the parentheses. While there are other ways to write this equation, I find this the least confusing way to remove rows based on removal criteria from multiple variables.

```{r}

d8 %>% 
  dplyr::filter(!(extra2 ==-999 | extra3 ==-999))

```


**2\. Remove rows based on multiple criteria from multiple variables (in this case for two numeric variables). Filter out any row that has -999 for *BOTH* extra2 *AND* for extra3.**

Review the data (d8)

```{r, echo=FALSE}

source("data.R")
d8

```

Filter out any row that has -999 for extra2 and has -999 extra3. 

* Note: Notice that I put the operator *!* outside of the parentheses. While there are other ways to write this equation, I find this the least confusing way to remove rows based on removal criteria from multiple variables.


```{r}

d8 %>% 
  dplyr::filter(!(extra2 ==-999 & extra3 ==-999))


```


**3\. Remove rows based on multiple criteria (in this case for two numeric variables). Filter out any row that has -999 for extra2 *AND/OR* has -999 for extra3. While also keeping your NA values.**

Review the data (d5)

```{r, echo=FALSE}

source("data.R")
d5

```

Filter out any row that has -999 for extra2 and/or -999 for extra3 (retaining any row that has NA values for those variables). 

* Note: Again, you'll notice that I make an explicit call to **keep** rows with NA values for extra2 and extra3 in addition to removing rows with the value of -999 for extra2 and extra 3, using the *|* operator to denote and/or. Remember, a drawback of the `filter` function is that if you filter on a certain variable it will not only filter out the values you request but it will also drop any row that has NA for those variables. And we may not necessarily want that. In this case, I want to keep the last row of data that has NA for extra2.

* Note: Notice that for the first part of the equation I put the operator *!* outside of the parentheses. While there are other ways to write this equation, I find this the least confusing way to remove rows based on removal criteria from multiple variables.

```{r}

d5 %>% 
  dplyr::filter(!(extra2==-999 | extra3 ==-999) | (is.na(extra2) | is.na(extra3)))

```

Again, to get around the issue of dropping NAs we can use the operator *%in%* rather than *==* to make sure we keep the NA values for extra2 and extra3.

```{r}

d5 %>% 
  dplyr::filter(!(extra2 %in% -999 | extra3 %in% -999))

```


Return to [Filter](https://github.com/Cghlewis/data-wrangling-functions/wiki/Filter)
