---
title: "Filter empty rows"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```


### Package: dplyr

---

#### Function: `filter()`

---

**1\. Remove empty rows**

Review the data (d9)

```{r, echo=FALSE}

source("data.R")
d9

```

Filter out any completely empty row

* Note: Note the use of `dplyr::if_all()` predicate function to select columns. This function is available in version 1.0.5 of `dplyr`. You can read this as, keep rows where ALL variables are NOT NA (hence the ! on the outside of the `dplyr::if_all()` function)

```{r}

d9 %>% 
  dplyr::filter(!(dplyr::if_all(everything(), ~ is.na(.))))


```



### Package: expss

---

#### Function: `drop_empty_rows()`

---

**1\. Filter out empty rows**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```


Filter out any completely empty row

```{r}

d1 %>% 
  expss::drop_empty_rows()

```


### Package: janitor

---

#### Function: `remove_empty()`

---


**1\. Remove empty rows**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Filter out any completely empty row

```{r}

d1 %>% 
  janitor::remove_empty("rows")

```


Return to [Filter](https://github.com/Cghlewis/data-wrangling-functions/wiki/Filter)
