---
title: "Filter rows based on NA values"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

### Package: dplyr

---

#### Function: `filter()`

Examples using one criteria, remove NA

---


**1\. Remove rows based on one criteria. Remove any row whose value is not NA for extra2.**

Review the data (d7).

```{r, echo=FALSE}

source("data.R")
d7

```

Keep any row that does not have NA for extra2.

* Note: Use the logical operator *!* to denote NOT.

```{r}

d7 %>% dplyr::filter(!is.na(extra2))

```

If you wanted to apply this criteria to multiple variables (ex: extra 2 and extra 3), you could add the *OR* operator `|`.

* Note: Notice that I put the operator ! outside of the parentheses. While there are other ways to write this equation, I find this the least confusing way to remove rows based on removal criteria from multiple variables.

```{r}

d7 %>% dplyr::filter(!(is.na(extra2) | is.na(extra3)))

```

**2\. Filter out any row that has NA for *ALL* numeric variables.**

Review the data (d9)

```{r, echo=FALSE}

source("data.R")
d9

```

Filter out any row that has NA for **ALL** numeric variables

* Note: Note the use of `dplyr::if_all()` predicate function to select columns. This function is available in version 1.0.5 of `dplyr`. You can read this as, keep rows where ALL numeric variables are NOT NA (hence the ! on the outside of the `dplyr::if_all()` function)


```{r}

d9 %>% 
  dplyr::filter(!(dplyr::if_all(is.numeric,~ is.na(.))))


```

### Package: tidyr

---

#### Function: `tidyr::drop_na()`

---

**1\.Filter out rows based on a select variable (stu_id) having NA values**

Review the data (d7)

```{r, echo=FALSE}

source("data.R")
d7

```


Filter out any row that has NA for stu_id.

```{r}

d7 %>% 
  tidyr::drop_na(stu_id)


```


**2\.Filter out rows based on any variable having NA values**

Review the data (d7)

```{r, echo=FALSE}

source("data.R")
d7

```

Filter out any row that has NA for any variable.

```{r}

d7 %>% 
  tidyr::drop_na()


```

Return to [Filter](https://github.com/Cghlewis/data-wrangling-functions/wiki/Filter)
