---
title: "Filter rows based on variable type"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```


### Package: dplyr

---

#### Function: `filter()`

---

* Note: Using `dplyr::across()` in `dplyr::filter()` is deprecated. `dplyr::if_any()` and `dplyr::if_all()` are predicate functions used to select columns within `dplyr::filter()`. This function is available in version 1.0.5 of `dplyr`. `dplyr::if_any()` returns a true when the statement is true for **any** of the variables. `dplyr::if_all()` returns a true when the statement is true for **all** of the variables. See [Filter using if_all or if_any](https://cghlewis.github.io/data-wrangling-functions/filter-data/filter-rows-if.html) for further explanation

<br>

**1\. Remove any row that has -999 for *ALL* numeric variables.**

Review the data (d19)

```{r, echo=FALSE}

source("data.R")
d19

```

Filter out any row that has -999 for **ALL** numeric variables

Note: You need to wrap `is.numeric`, a predicate function (returns a true/false), in the `tidyselect` selection helper `where()`.

```{r}

d9 %>% 
  dplyr::filter(!dplyr::if_all(where(is.numeric),~ . == -999))


```


**2\. Remove any row that has -999 for *AT LEAST ONE* numeric variable.**

Review the data (d19)

```{r, echo=FALSE}

source("data.R")
d19

```

Filter out any row that has -999 for **AT LEAST ONE** numeric variable


```{r}

d19 %>% 
  dplyr::filter(!dplyr::if_any(where(is.numeric), ~ . == -999))

```

Return to [Filter](https://github.com/Cghlewis/data-wrangling-functions/wiki/Filter)
