---
title: "Filter columns"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```


### Package: dplyr

---

#### Function: `select()`

---

**1\. Drop columns (extra1, extra2, extra3)**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Drop extra1, extra2, extra3

```{r}

d1 %>% dplyr::select(-c(extra1:extra3))

```

You can also drop columns based on shared column attributes using `tidyselect` selection helpers such as `starts_with()`.

* Note: Other `tidyselect` selection helpers include `contains()`, `ends_with()`, `where()`, `matches()`, and more.

```{r}

d1 %>% dplyr::select(-(tidyselect::starts_with("extra")))

```


**2\. Keep selected columns (id and test_score)**

```{r, echo=FALSE}

source("data.R")
d2

```

Keep id and test_score

```{r}

d2 %>% dplyr::select(id:test_score2)

```

You can also keep columns based on shared column attributes using `tidyselect` selection helpers such as `contains()`.

* Note: Other `tidyselect` selection helpers include `starts_with()`, `ends_with()`, `where()`, `matches()`, and more.

```{r}

d2 %>% dplyr::select(id, tidyselect::contains("test"))

```

