---
title: "Filter rows based on NA values"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

### Package: dplyr

---

#### Function: `filter()`

Examples using one criteria, remove NA

---


**1\. Remove rows based on one criteria. Remove any row whose value is not NA for extra2.**

Review the data (d7).

```{r, echo=FALSE}

source("data.R")
d7

```

Keep any row that does not have NA for extra2.

* Note: Use the logical operator *!* to denote NOT.

```{r}

d7 %>% dplyr::filter(!is.na(extra2))

```


### Package: tidyr

---

#### Function: `tidyr::drop_na()`

---

**1\.Filter out rows based on a select variable (stu_id) having NA values**

Review the data (d7)

```{r, echo=FALSE}

source("data.R")
d7

```


Filter out any row that has NA for stu_id.

```{r}

d7 %>% 
  tidyr::drop_na(stu_id)


```


**2\.Filter out rows based on any variable having NA values**

Review the data (d7)

```{r, echo=FALSE}

source("data.R")
d7

```

Filter out any row that has NA for any variable.

```{r}

d7 %>% 
  tidyr::drop_na()


```

Return to [Filter](https://github.com/Cghlewis/data-wrangling-functions/wiki/Filter)
