---
title: "Filter rows based on multiple criteria for one variable"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```


### Package: dplyr

---

#### Function: `filter()`

Examples using multiple criteria same variable (numeric)

---


**1\. Keep any row that has -999 *OR* 0 for `extra3`.**

Review the data (d8).

```{r, echo=FALSE}

source("data.R")
d8

```

Keep any row that has -999 or 0 for `extra3`. 

* Note: Use of the operator *|* to denote and/or (*and* doesn't really apply since it's the same variable).

```{r}

d8 %>% 
  dplyr::filter(extra3 == -999 | extra3 ==0)

```

You would get the same result using the *xor* operator.

```{r}

d8 %>% 
  dplyr::filter(xor(extra3 == -999 , extra3 ==0))

```

Last, you can use the *%in%* operator as well.

```{r}

d8 %>% 
  dplyr::filter(extra3 %in% c(-999,0))

```


**2\. Keep any row whose value for `extra3` is between 0 and 10.**

Review the data (d8).

```{r, echo=FALSE}

source("data.R")
d8

```


Keep any row that has a value between 0 and 10 for `extra3`. 

* Note: Use the operator *&* to denote that both criteria must be met.

```{r}

d81 <- d8 %>% 
  dplyr::filter(extra3 >= 0 & extra3 <=10)

d81

```

You would get the same result using the `dplyr::between()` function. It includes both 0 and 10.

```{r}

d8 %>% 
  dplyr::filter(dplyr::between(extra3, 0, 10))

```

---

#### Function: `filter()`

Examples using multiple criteria same variable (character)

---

**1\. Keep any row whose `tch_name` is either "harris" *OR* "lewis".**

Review the data (d8).

```{r, echo=FALSE}

source("data.R")
d8

```

Keep any row where `tch_name` is "harris" or "lewis".

* Note: Use of the operator *|* to denote and/or (*and* doesn't really apply since it's the same variable).

```{r}

d8 %>% 
  dplyr::filter(tch_name == "harris" | tch_name == "lewis")

```

You would get the same result using the *%in%* operator.

```{r}

d8 %>% 
  dplyr::filter(tch_name %in% c("harris", "lewis"))

```

Or the same result adding the `stringr::str_detect()` function.

```{r}

d8 %>% 
  dplyr::filter(stringr::str_detect(tch_name, 'harris|lewis') )

```


Return to [Filter](https://github.com/Cghlewis/data-wrangling-functions/wiki/Filter)
