---
title: "Filter empty rows"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```


### Package: dplyr

---

#### Function: `filter()`

---

**1\.Filter out empty rows**

Review the data (d9)

```{r, echo=FALSE}

source("data.R")
d9

```

Count the number of cases in the current data using the function `dplyr::count()`

* Note: We could also use `base::nrow()` to get this number.

```{r}

dplyr::count(d9)

```


Filter out any completely empty row

```{r}

d9 <- d9 %>% dplyr::filter(!is.na(.))

d9

```


Count the number of cases after you filter

```{r}

dplyr::count(d9)

```



### Package: expss

---

#### Function: `drop_empty_rows()`

---

**1\.Filter out empty rows**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Count the number of cases in the current data using the function `dplyr::count()`

```{r}

dplyr::count(d1)

```


Filter out any completely empty row

```{r}

d1 <- d1 %>% expss::drop_empty_rows()

d1

```

Count the number of cases after you filter

```{r}

dplyr::count(d1)

```



### Package: janitor

---

#### Function: `remove_empty()`

---


**1\.Filter out empty rows**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Count the number of cases in the current data using the function `dplyr::count()`

```{r}

dplyr::count(d1)

```


Filter out any completely empty row

```{r}

d1 <- d1 %>% janitor::remove_empty("rows")

d1

```

Count the number of cases after you filter

```{r}

dplyr::count(d1)

```


