---
title: "Filter rows based on multiple criteria for multiple variables"
output: 
  html_document:
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```


### Package: dplyr


---

#### Function: `filter()`

---

* Note: Using `dplyr::across()` in `dplyr::filter()` is deprecated. `dplyr::if_any()` and `dplyr::if_all()` are predicate functions used to select columns within `dplyr::filter()`. This function is available in version 1.0.5 of `dplyr`. `dplyr::if_any()` returns a true when the statement is true for **any** of the variables. `dplyr::if_all()` returns a true when the statement is true for **all** of the variables. See [Filter using if_all or if_any](https://cghlewis.github.io/data-wrangling-functions/filter-data/filter-rows-if.html) for further explanation

<br>

**1\. Return rows where `form` is empty, but ANY values for variables (var1:var4) are NOT missing. Missing is denoted by "MR" as well as NA.**

Review the data (d17)

```{r, echo=FALSE}

source("data.R")
d17

```

We can see here that form is a character variable and missing values are denoted by "", so we will need to account for that in our formula.

* Note: You'll notice here that I did NOT use `%in%` too keep NA values. Remember from [Filter rows with NA values](https://cghlewis.github.io/data-wrangling-functions/filter-data/filter-rows-na.html), NA values are not evaluated and are dropped automatically when you use `!=`. In this case, since both "MR" and NA are missing values, I am okay with NA values being dropped by default.


```{r}

d17 %>%
  filter(form == "" & dplyr::if_any(var1:var4, ~ . != "MR"))

```

I could get the same outcome by being more explicit and using this formula instead.

```{r}

d17 %>%
  filter(form == "" & dplyr::if_any(var1:var4, ~ !. %in% c("MR", NA)))

```

Or using something like this.

```{r}

d17 %>%
  filter(form == "" & dplyr::if_any(var1:var4, ~ !(. == "MR"|is.na(.))))

```

Or I could also get the same result this way, to say "return the rows where form is missing but ALL values for variables (var1:var4) are not missing."

```{r}

d17 %>%
  filter(form == "" & !dplyr::if_all(var1:var4, ~ . %in% c("MR", NA)))

```


**2\. Return rows where `form` is _not_ empty, and ALL variables (var1:var4) are empty.**

Review the data (d17)

```{r, echo=FALSE}

source("data.R")
d17

```


```{r}

d17 %>%
  filter(form != "" & dplyr::if_all(var1:var4, ~ . %in% c("MR", NA)))

```


Return to [Filter](https://github.com/Cghlewis/data-wrangling-functions/wiki/Filter)
