---
title: "Separate rows"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: tidyr

---

#### Function: `separate_rows()`

---

* Note: Unlike `tidyr::separate()` which splits columns into new columns, `tidyr::separate_rows()` splits a column with multiple delimited values into new rows.

**1\. Split a variable (toppings) into multiple rows**

Review the data (d5)

```{r, echo=FALSE}

source("data.R")
d5

```

Split toppings into new rows

* Note: The original variable will be removed unless you add the argument *remove=FALSE*
* Note: If no separator is indicated, the function will split values wherever a non-alphanumeric character appears. If you want to be clear where to separate, add the argument *sep=","* as I did below otherwise the words *red* and *pepper* will be split into separate rows.

```{r}

d5 %>% tidyr::separate_rows(toppings, sep=",", convert=TRUE)

```
