---
title: "Separate rows"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: tidyr

---

#### Function: `separate_rows()`

---

* Note: Unlike `tidyr::separate()` which splits columns into new columns, `tidyr::separate_rows()` splits a column with multiple delimited values into new rows.

**1\. Split a variable (`toppings`) into multiple rows**

Review the data (d5)

```{r, echo=FALSE}

source("data.R")
d5

```

Split `toppings` into new rows

* Note: The original variable will be removed unless you add the argument *remove=FALSE*
* Note: If no separator is indicated, the function will split values wherever a non-alphanumeric character appears. If you want to be clear where to separate, add the argument *sep=","* as I did below otherwise the words *red* and *pepper* will be split into separate rows.

```{r}

d5 %>% 
  tidyr::separate_rows(toppings, sep=",")

```

* Note: While I think `tidyr::separate_rows()` is a very clean way of accomplishing the goal above, I think it is nice to see that there are other methods to accomplishing this result. One of those other ways is using `stringr::str_split()`. 

Notice here though that you have to use `dplyr::mutate()` to create a new column (or in this case replace the existing column). 

```{r}

d5 %>%
  mutate(toppings = stringr::str_split(toppings, pattern = ",")) 

```

You'll also notice that you do not get a normal character vector for `toppings`. You get a nested list-column that now requires us to `tidyr::unnest()` in order to see the information.

So while this method does work, you can see what `tidyr::separte_rows()` is much simpler for this use case.

```{r}

d5 %>%
  mutate(toppings = stringr::str_split(toppings, pattern = ",")) %>%
  tidyr::unnest()

```


Return to [Separate](https://github.com/Cghlewis/data-wrangling-functions/wiki/Separate)