---
title: "Separate columns"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: tidyr

---

#### Function: `separate()`

---

**1\. Split a variable (name) into first name and last name**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Split name into two variables

* Note: The original variable will be removed unless you add the argument *remove=FALSE*
* Note: If no separator is indicated, the function will split values wherever a non-alphanumeric character appears. If you want to be clear where to separate, add the argument *sep=* and insert a regular expression such as *sep=" "*. If working with numeric columns rather than character columns, you can separate based on character position.

```{r}

d1 %>% tidyr::separate(name, into=c("first_name", "last_name"))

```

* Note: The number of new variables must equal the number of pieces you are separating. So for example if some names had a middle initial, we would only receive the first name and middle initial for those names

```{r, echo=FALSE}
d2

```

```{r}

d2 %>% tidyr::separate(name, into=c("first_name", "last_name"))

```

You can make decisions on how to handle these anomalies by adding the argument *extra=* and add the option "warn", "drop" or "merge", with the default being "warn". To learn more type *?separate* in the console.

Here is an example using *extra="merge"*

```{r}

d2 %>% tidyr::separate(name, into=c("first_name", "last_name"), extra="merge")

```

* Note: Similarly since the number of new variables must equal the number of pieces you are separating, if some names only have a first or last name, you will have missing data in your new variables.

```{r, echo=FALSE}

d3

```

```{r}

d3 %>% tidyr::separate(name, into=c("first_name", "last_name"))

```

You can make decisions on how to handle these anomalies by adding the argument *fill=* and add the option "warn", "right" or "left", with the default being "warn". To learn more type *?separate* in the console.

Here is an example using *fill="left"*

```{r}

d3 %>% tidyr::separate(name, into=c("first_name", "last_name"), fill="left")

```

