---
title: "Separate columns"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: tidyr

---

#### Function: `separate()`

---

* Note: Since writing this, `tidyr::separate()` has been superseded by `separate_wider_position()` and `separate_wider_delim()`. See this [link](https://tidyr.tidyverse.org/reference/separate.html) for more info.

<br>

**1\. Split a character variable (`name`) into first name and last name**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Split `name` into two variables

* Note: The values for the *into* argument need to be provided as a character vector.

* Note: The original variable will be removed unless you add the argument *remove=FALSE*

* Note: If no separator is indicated, the function will split values wherever a non-alphanumeric character appears (for example, a space). If you want to be clear where to separate, add the argument *sep=* and insert a regular expression such as *sep=" "*. 

```{r}

d1 %>% 
  tidyr::separate(name, into=c("first_name", "last_name"))

```

* Note: The number of new variables must equal the number of pieces you are separating. So for example if some names had a middle initial, we would only receive the first name and middle initial for those names

```{r, echo=FALSE}
d2

```

```{r}

d2 %>% 
  tidyr::separate(name, into = c("first_name", "last_name"))

```

You can make decisions on how to handle these anomalies by adding the argument *extra=* and add the option "warn", "drop" or "merge", with the default being "warn". To learn more type *?separate* in the console.

Here is an example using *extra="merge"*

```{r}

d2 %>% 
  tidyr::separate(name, into = c("first_name", "last_name"), extra="merge")

```

* Note: Similarly since the number of new variables must equal the number of pieces you are separating, if some names only have a first or last name, you will have missing data in your new variables.

```{r, echo=FALSE}

d3

```

```{r}

d3 %>% 
  tidyr::separate(name, into = c("first_name", "last_name"))

```

You can make decisions on how to handle these anomalies by adding the argument *fill=* and add the option "warn", "right" or "left", with the default being "warn". To learn more type *?separate* in the console.

Here is an example using *fill="left"*

```{r}

d3 %>% 
  tidyr::separate(name, into = c("first_name", "last_name"), fill="left")

```

**2\. Split only some of the cells in a character column (`school_name`) and keep only one column of the split**

Review the data (d6)

```{r, echo=FALSE}

source("data.R")
d6

```

In this example, I want to clean up the school names by removing any first name and middle initials from school names. Most schools only go by the last name. So I want to remove David H. and Muriel W.

With only a few values, I could easily do a recode. However, if there were hundreds of schools with similar first name and middle initial patterns, that would take too long. So here we use `tidyr::separate()`.

* Note: We use regular expressions (regex) here to denote we are separating on any capitalized letter "[A-Z]" followed by a period "\\." and last followed by a white space "\\s". Notice the use of "\\" as an escape to tell the regex to match the period exactly, rather than using it for its special behavior which is to match *any* character. You can also put the period in [] as an escape used to match characters within the brackets, similar to the "[A-Z]" above.

* Note: I added the argument *fill="left"* to bring any value that did not separate into our new `school_name` column from the left side to the right column "school_name".

```{r}

d6 %>% 
  tidyr::separate(school_name, sep = "[A-Z]\\.\\s", into = c("name1", "school_name"), fill="left")

```

Notice now that we have an extra column `name1` that I don't want. I could either remove this using a `dplyr::select()` statement OR I could name this "name1" column NA in my function above and the column would drop automatically.

```{r}

d6 %>% 
  tidyr::separate(school_name, sep = "[A-Z]\\.\\s", into = c(NA, "school_name"), fill="left")

```



**3\. Split a numeric variable (`stid_schid`) into `state_id` and `school_id`**

Review the data (d7)

```{r, echo=FALSE}

source("data.R")
d7

```

Split stid_schid into 2 variables

* Note: If working with numeric columns you can separate based on character position.

```{r}

d7 %>% 
  tidyr::separate(stid_schid, sep = 3, into=c("state_id", "school_id"))

```

Return to [Separate](https://github.com/Cghlewis/data-wrangling-functions/wiki/Separate)