---
title: "Separate rows and columns"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: tidyr

---

#### Function: `separate_longer_delim()`, `separate_wider_delim()`

---

<br>

**1\. Restructure a data dictionary into long format**

Review the dictionary

```{r, echo=FALSE}

source("data.R")
dictionary

```

Here I want to convert my wide formatted data dictionary into long format. I want all of my variables to repeat a row for each response option and I want each row to only show one response option.

I also want the response option to be split out into a value column and a label column.

So here I will use `separate_longer_delim()` to first put each response option on their own row. Then I use `separate_wider_delim()` to split the response option value and label into their own columns. If there are variables that only have one option (like the ID and age variables), I added the argument *too_few* to keep the values in the `value` column

```{r}

dictionary %>%
  separate_longer_delim(var_values, delim = "| ") %>%
  separate_wider_delim(var_values, delim = " = ", names = c("value", "value_label"), too_few = "align_start")

```



Return to [Separate](https://github.com/Cghlewis/data-wrangling-functions/wiki/Separate)