---
title: "Complete long data"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: tidyr

---

#### Function: `fill()`

---

**1\. Fill NA values for both school and year with existing values**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Fill values using the default argument *direction = "down"*.

```{r}

d1 %>% tidyr::fill(school, year)

```

**2\. Fill NA values for cohort with existing values (grouped by the School Name)**

Review the data (d4)

```{r, echo=FALSE}

source("data.R")
d4

```

Fill values using the default argument *direction = "down"*.

You have two options. You can either use `dplyr::group_by()` to group by school.

```{r}

d4 %>% dplyr::group_by(school) %>% tidyr::fill(cohort) %>% dplyr::ungroup()

```

Or you can arrange your data by school using `dplyr::arrange()` and then fill values.

```{r}

d4 %>% dplyr::arrange(school) %>% tidyr::fill(cohort)

```


---

#### Function: `complete()`

---

**1\. Create rows for any missing time periods in the data**

Review the data (d2)

```{r, echo=FALSE}

source("data.R")
d2

```

Complete the data by creating rows for any student that is missing a time period and filling the test score with NA.

```{r}

d2 %>% tidyr::complete(s_id, time_period)

```

* Note: If you want to fill with a value other than NA, add the argument *fill = list()* and add the value you want to replace NA with. 

Here is an example, using *fill = list(test_score=0)*

```{r}

d2 %>% tidyr::complete(s_id, time_period, fill = list(test_score=0))

```

Return to [Complete](https://github.com/Cghlewis/data-wrangling-functions/wiki/Complete)