---
title: "Complete long data using existing values"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: tidyr

---

#### Function: `fill()`

---

**1\. Fill NA values for both `school` and `year` with existing values**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Fill values using the default argument *direction = "down"*.

```{r}

d1 %>% 
  tidyr::fill(school, year)

```

**2\. Fill NA values for `cohort` with existing values (grouped by the `school`)**

Review the data (d4)

```{r, echo=FALSE}

source("data.R")
d4

```

Fill values using the default argument *direction = "down"*.

You have two options. You can either use `dplyr::group_by()` to group by `school`.

```{r}

d4 %>% 
  dplyr::group_by(school) %>% 
  tidyr::fill(cohort) %>% 
  dplyr::ungroup()

```

Or you can arrange your data by `school` using `dplyr::arrange()` and then fill values. This option is only possible because there are consistently 2 rows for each school, and data is always missing from the later years. If the data was more inconsistent, you would want to stick with the `dplyr::group_by()` option.

```{r}

d4 %>% 
  dplyr::arrange(school) %>% 
  tidyr::fill(cohort)

```


**3\. Fill NA values for `cohort` with existing values (grouped by the `school`)**

Review the data (d7)

```{r, echo=FALSE}

source("data.R")
d7

```

In this scenario, the data is inconsistently organized. Sometimes we only have values in later years and need to fill earlier years, sometimes data doesn't exist at all for certain years. In this case, we will need to make sure to use `dplyr::group_by()` as well as the argument *direction = "downup"* to make sure our values are correctly filled in.

```{r}

d7 %>% 
  dplyr::group_by(school) %>% 
  tidyr::fill(cohort, .direction = "downup") %>% 
  dplyr::ungroup()

```

---

#### Function: `complete()`

---

**1\. Create rows for any missing time periods in the data**

Review the data (d2)

```{r, echo=FALSE}

source("data.R")
d2

```

Complete the data by creating rows for any student that is missing a time period and filling the test score with NA.

```{r}

d2 %>% 
  tidyr::complete(s_id, time_period)

```

* Note: If you want to fill with a value other than NA, add the argument *fill = list()* and add the value you want to replace NA with. 

Here is an example, using *fill = list(test_score=0)*

```{r}

d2 %>% 
  tidyr::complete(s_id, time_period, fill = list(test_score=0))

```

Return to [Complete](https://github.com/Cghlewis/data-wrangling-functions/wiki/Complete)