---
title: "Count distinct number of groups"
output: 
  html_document:
    toc: false
    anchor_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

<br>

### Package: dplyr

---

#### Function: `count()`

---

**1\. Count how many distinct schools we have**

Review the data (d1)

```{r, echo=FALSE}

source("data.R")
d1

```

Count the distinct number of schools

Because schools inherently occur more than once in a dataset, you need to consider how to remove duplicates to get distinct values. We can use `dplyr::distinct()` to get our distinct school names and then do a count.

```{r}

d1 %>%
  dplyr::distinct(school) %>%
  dplyr::count()

```

We could also use another function `dplyr::n_groups()` which requires you to first group by your variable of interest (in this case `school`) and then use the function.

Note: It is important to note that the output between using `dplyr::count()` and `dplyr::n_groups()` is different. The former produces a tibble and the latter produces a numeric vector.

```{r}

d1 %>%
  dplyr::group_by(school) %>%
  dplyr::n_groups()

```


Return to [Count](https://github.com/Cghlewis/data-wrangling-functions/wiki/Count)
